---
title: "EIC compounds individually"
subtitle: "LC-MS standards"
author: "Mar Garcia-Aloy"
html_document:
    toc: true
    number_sections: true
    toc_float: true
---

```{r, include = TRUE}
knitr::opts_chunk$set(echo=TRUE)

startpoint <- Sys.time()
```

In this document I'm going to print the EIC plots of all compounds found in a specific MIX. 
For that I'll use the second injeciton of MIX "B", which concentration was 25 ppm.

# Preliminaries
I define the specific parameters for the current study, I load all 
required libraries, and I set up the parallel processing.

## Parameters
```{r settings}
polarity <- "NEG" # specify "POS" or "NEG"
mzd <- 0.01
my.mix <- 9
MZML_PATH <- "C:/Users/mgarciaaloy/Documents/mzML/"
```

## Libraries
```{r libraries, message=FALSE, warning=FALSE}
library(doParallel)
library(xcms)
library(RColorBrewer)
library(Rdisop)
library(CompoundDb)
```

## Parallel processing
```{r parallel-proc}
ncores <- detectCores()-1
registerDoParallel(ncores)
register(DoparParam(), default = TRUE)
```

# Data import

## Standard library
I load below the most likely adduct and the expected retention time for the 
standards defined in 
[our inhouse repository](http://github.com/EuracBiomedicalResearch/lcms-standards).

```{r std-library}
std_info <- read.table(
  #"https://raw.githubusercontent.com/EuracBiomedicalResearch/lcms-standards/master/data/standards_dilution.txt",
  "C:/Users/mgarciaaloy/Documents/GitHub/lcms-standards/data/standards_dilution.txt",
  sep = "\t", header = TRUE, as.is = TRUE)

std_info <- std_info[std_info$mix == my.mix, ]

std_info$adduct <- std_info[, grep(polarity, colnames(std_info))]

std_info$POS[is.na(std_info$POS)] <- "[M+H]+"
std_info$NEG[is.na(std_info$NEG)] <- "[M-H]-"

std_info$mzneut = NA
std_info$mz_ion = NA
for (i in seq(nrow(std_info))) {
  if (grepl("C", std_info$formula[i])){
    std_info$mzneut[i] <- getMolecule(
      as.character(std_info$formula[i]))$exactmass
  } else {
    std_info$mzneut[i] = as.numeric(std_info$formula[i])
  }
  #' Calculate also the m/z
  std_info$mz_ion[i] <- unlist(
    mass2mz(std_info$mzneut[i],
            adduct = as.character(
              std_info[i, grep(polarity, colnames(std_info))])))
}

std_info <- std_info[order(std_info$RT),]
```

# mzML file
```{r mzML}
dda_data <- 
  readMSData(paste0(MZML_PATH, "2016/2016_05/MIX ", my.mix, "B_", 
                    polarity, "_b.mzML"), mode = "onDisk")
```

# Output
```{r EIC, fig.height=nrow(std_info), fig.width=10}
getPalette = colorRampPalette(brewer.pal(9, "Set1"))
mycols = getPalette(nrow(std_info))  

par(mfrow = c(nrow(std_info), 1),
    mar = c(0, 0, 0, 0)) # bottom = 5, left = 4, top = 4, right = 2
for(i in seq(nrow(std_info))){
  mymz <- std_info$mz_ion[i]
  if(!is.na(mymz)){
    chr <- chromatogram(dda_data, mz = c(mymz - mzd, mymz + mzd))[[1]]
    chr <- data.frame(rt = rtime(chr),
                      i = intensity(chr))
    chr[is.na(chr)] <- 0
    chr$i100 <- (chr$i / max(chr$i, na.rm = TRUE))*100
    plot(chr$rt, chr$i100, type = "l", xaxt='n', col = mycols[i])
    if(!is.na(std_info[i, "adduct"])){
      text(chr$rt[which.max(chr$i)]+5, 100, round(chr$rt[which.max(chr$i)]))
    }
    text(400, 90, std_info$name[i], font = 2)
    text(400, 75, paste0(std_info[i, grep(polarity, colnames(std_info))], ": ",
                         round(mymz - mzd, 5), " - ", round(mymz + mzd, 5)))
    if(!is.na(std_info[i, "adduct"])){
      text(400, 60, 
           paste0("max. intensity = ", 
                  formatC(max(chr$i, na.rm = TRUE), format = "e", digits = 2), 
                  " counts"))
    }
  }
}
```

# Session information
```{r session-info}
Sys.time()-startpoint

devtools::session_info()
```
