---
title: "Define the retention times"
author: "Mar Garcia-Aloy"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false
    toc_float: true
---

```{r settings, echo = FALSE, results = "hide", message = FALSE}
study <- "standards_dilution" # specify "internal_standards" OR 
                              #         "standards_dilution"
mixnum <- 17 # specify which MIX
polarity <- "POS" # specify "POS" or "NEG"
da <- 0.01

#' Define the path where we can find the mzML files:
MZML_PATH <- "C:/Users/mgarciaaloy/Documents/mzML/"
#if (!file.exists(MZML_PATH))
#    stop("Can not find the directory with the mzML files: ", MZML_PATH)
```

# Preliminaries 
```{r, message=FALSE}
library(xcms)
library(magrittr)
library(CompoundDb)
library(Rdisop)
```

# Data import
In the file `XXXX_files.txt` there is the information regarding 
the injection sequence.
```{r}
injections <- read.table(paste0("data/", study, "_files.txt"), #import the file
                         sep = "\t", header = TRUE, as.is = TRUE)
myfiles <- injections$mzML # get file names
myfiles <- myfiles[grep(polarity, myfiles)]# select files names of our polarity
if(study == "standards_dilution"){
  myfiles <- myfiles[grep(paste0("MIX ", mixnum, "K"), myfiles)]
  }
myfiles
```

Import the information regarding the standards that are in the samples.  
```{r}
std_info <- read.table(paste0("data/", study, ".txt"),
                      sep = "\t", header = TRUE, as.is = TRUE)
std_info$name <- c(substring(std_info$name, 1, 33))
if(study == "standards_dilution"){
  std_info <- subset(std_info, mix == mixnum)
}
std_info$mzneut = NA
for(i in seq(nrow(std_info))){
  if(grepl("C", std_info$formula[i])){std_info$mzneut[i] = 
    getMolecule(as.character(std_info$formula[i]))$exactmass}else{
      std_info$mzneut[i] = paste(std_info$formula[i])}
}
write.csv(std_info, "x.csv", row.names = FALSE)
std_info = read.csv("x.csv")
std_info
std_info <- std_info[!is.na(std_info[, grep(polarity, colnames(std_info))]),]
mzvalues <- c()
for(i in seq(nrow(std_info))){
  mzvalues <- c(mzvalues, 
                unlist(mass2mz(std_info$mzneut[i], 
                               adduct = 
                                 as.character(
                                   std_info[i, 
                                            grep(polarity, colnames(std_info))]
                                   ))))
}
```

# Inspect data
Steps:  
  
* Import the file `j`.  
* Get the EIC for the compound `i`.  
* Annotate the RT of the highest peak in the dataframe `myrt`.  
* Get the spectrum of that RT.  
* Annotate the mz corresponding to compound `i` in the dataframe `mymz`.   
* Annotate the relative intensity of the mz within the spectrum.  

```{r}
mymz <- data.frame(matrix(ncol=nrow(std_info), nrow=length(myfiles)))
colnames(mymz) <- std_info$name

myrt <- data.frame(matrix(ncol=nrow(std_info), nrow=length(myfiles)))
colnames(myrt) <- std_info$name

myint <- data.frame(matrix(ncol=nrow(std_info), nrow=length(myfiles)))
colnames(myint) <- std_info$name

if(study == "internal_standards"){
  colnames(mymz) <- gsub(" .*","",colnames(mymz))
  colnames(mymz) <- gsub("L-", "", colnames(mymz))
  
  colnames(myrt) <- gsub(" .*","",colnames(myrt))
  colnames(myrt) <- gsub("L-", "", colnames(myrt))
  
  colnames(myint) <- gsub(" .*","",colnames(myint))
  colnames(myint) <- gsub("L-", "", colnames(myint))
}
```


```{r}
for(j in seq(length(myfiles))){
  # Import the file "j":
  raw_data <- readMSData(paste0(MZML_PATH, myfiles[j]), 
                         mode = "onDisk")
  for(i in seq(nrow(std_info))){
    # Get the EIC for the compound "i":
    chr <- chromatogram(raw_data, aggregationFun = "max",
                        mz=c(mzvalues[i] - da,
                             mzvalues[i] + da))
    
    # Annotate the RT of the highest peak:
    myrt[j,i] <- chr@.Data[[1]]@rtime[which.max(chr@.Data[[1]]@intensity)]
    
    # Get the spectrum of that RT:
    sps <- raw_data %>%
      filterRt(rt = c(chr@.Data[[1]]@rtime[which.max(
        chr@.Data[[1]]@intensity)] - 0.5,
        chr@.Data[[1]]@rtime[which.max(
          chr@.Data[[1]]@intensity)] + 0.5)) %>%
      spectra 
    sps.df=as.data.frame(sps[[2]])
    
    if(length(sps.df$mz[which(
      sps.df$mz > (mzvalues[i] - da) & sps.df$mz < (mzvalues[i] + da))]) > 0){
      
      # Annotate the mz corresponding to compound "i":
      mymz[j,i] <- sps.df$mz[which(
        sps.df$mz > (mzvalues[i] - da) & sps.df$mz < (mzvalues[i] + da))]
      
      # Annotate the relative intensity of the mz within the spectrum:
      sps.df$irel <- (sps.df$i*100)/max(sps.df)
      myint[j,i] <- sps.df$irel[which(
        sps.df$mz > (mzvalues[i] - da) & sps.df$mz < (mzvalues[i] + da))]
      }
    
    
  }
}
```

# Output
## RT
```{r}
t(myrt)
cbind(colnames(myrt), 
      matrix(apply(myrt, 2, function(x) round(median(x, na.rm = TRUE)))))
if(study == "internal_standards"){
  t(apply(myrt, 2, function(x) round(range(x, na.rm = TRUE))))
}
cbind(colnames(myrt), 
      matrix(apply(myrt, 2, function(x) round(diff(range(x, na.rm = TRUE))))))
```

## mz
```{r}
(mymz.t <- data.frame(t(mymz)))
cbind(colnames(mymz), 
      matrix(apply(mymz, 2, function(x) round(median(x, na.rm = TRUE), 5))))
t(apply(mymz, 2, function(x) round(range(x, na.rm = TRUE), 4)))
round(abs((((mymz.t) - mzvalues) / mzvalues) * 1e6)) # error in ppm
round(abs((((mymz.t) - mzvalues)) * 1e3)) # error in mDa
```

## Intensity
```{r}
t(round(myint))
cbind(colnames(myint), 
      matrix(apply(myint, 2, function(x) round(median(x, na.rm = TRUE)))))
if(study == "internal_standards"){
  t(apply(myint, 2, function(x) round(range(x, na.rm = TRUE))))
}
```

## All
```{r}
cbind(t(round(myrt)), mymz.t, t(round(myint)))
```

