---
title: "Evaluate internal standards"
author: "Mar Garcia-Aloy"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false
    toc_float: true
---

```{r settings, echo = FALSE, results = "hide", message = FALSE}
polarity <- "NEG" # specify "POS" or "NEG"

#' Define the path where we can find the mzML files:
MZML_PATH <- "C:/Users/mgarciaaloy/Documents/mzML/"
#if (!file.exists(MZML_PATH))
#    stop("Can not find the directory with the mzML files: ", MZML_PATH)
```

# Preliminaries
```{r, message=FALSE}
source("R/listpeaks.R")

library(xcms)
library(reshape2)
library(ggplot2)
```

# Data import
In the file `internal_standards_files.txt` there is the information regarding 
the injection sequence.
```{r}
injections <- read.table("data/internal_standards_files.txt", # import the file
                         sep = "\t", header = TRUE, as.is = TRUE)
myfiles <- injections$mzML # get file names
myfiles <- myfiles[grep(polarity, myfiles)]# select files names of our polarity
```

In the file `internal_standards.txt` there is the information regarding 
the standards that are in the samples.
```{r}
is_info <- read.table("data/internal_standards.txt",
                      sep = "\t", header = TRUE, as.is = TRUE)
colnames(is_info) <- c("name", "molecular_weight", "POS", "NEG", "RT")
is_info <- is_info[,1:5]
is_info <- is_info[!is.na(is_info[, grep(polarity, colnames(is_info))]),]
head(is_info)
```

# Peak detection
Apply the `listpeaks()` function in order to detect the peaks of ISs:
```{r, warning=FALSE}
mzvalues <- is_info[, grep(polarity, colnames(is_info))]
rtvalues <- is_info$RT
peaktb <- listpeaks(file = paste0(MZML_PATH, myfiles),
                    mz = mzvalues,
                    param = CentWaveParam(peakwidth = c(2, 10), 
                                          ppm = 30))
peaktb$name <- is_info$name[peaktb$row]

rtvalues <- rtvalues[peaktb$row]
peaktb <- peaktb[(peaktb$rtmin <= (rtvalues)) & (peaktb$rtmax >= (rtvalues)),]

peaktb_wide <- dcast(peaktb, peaktb$name ~ peaktb$colum, value.var = "into")
peaktb_wide <- setNames(data.frame(t(peaktb_wide[,-1])), peaktb_wide[,1])
peaktb_wide[is.na(peaktb_wide)] <- 0

```
  
# Plots  
## Absolute intensities  
```{r}
myplots = list()
for(i in seq(ncol(peaktb_wide))){
  tb <- data.frame(sample=seq(length(myfiles)), intensity=peaktb_wide[,i])
  myplots[[i]] <- 
    ggplot(tb, aes(sample, intensity)) + 
    geom_point() + geom_line() + theme_classic() + 
    ggtitle(colnames(peaktb_wide)[i]) +
    xlab("") + ylab("Intensity")
}
ggsave(paste0("internal_standard_evaluation_absolute_", polarity, ".pdf"), 
       gridExtra::marrangeGrob(grobs = myplots, top="", 
                               layout_matrix = 
                                 rbind(c(1,2,3,4), c(5,6,7,8), c(9,10,11,12), 
                                       c(13,14,15,16), c(17,18,19,20))), 
       width=21, height=15)
```
  
## Relative intensities  
```{r}
myplots = list()
for(i in seq(ncol(peaktb_wide))){
  tb <- data.frame(sample = seq(length(myfiles)), 
                   intensity = peaktb_wide[,i] / 
                     mean(peaktb_wide[peaktb_wide[,i] != 0, i]))
  myplots[[i]] <- 
    ggplot(tb, aes(sample, intensity)) + 
    geom_point() + geom_line() + theme_classic() + 
    ggtitle(colnames(peaktb_wide)[i]) +
    xlab("") + ylab("Intensity")
}
ggsave(paste0("internal_standard_evaluation_relative_", polarity, ".pdf"), 
       gridExtra::marrangeGrob(grobs = myplots, top="", 
                               layout_matrix = 
                                 rbind(c(1,2,3,4), c(5,6,7,8), c(9,10,11,12), 
                                       c(13,14,15,16), c(17,18,19,20))), 
       width=21, height=15)
```

