---
title: "Counting the number of C, H, N, O, P, S for compounds in HMDB"
author: "Andrea Vicini, Johannes Rainer"
output:
  rmarkdown::html_document:
    highlight: pygments
    toc: true
    toc_depth: 3
    fig_width: 5
---

```{r style, echo = FALSE, results = 'asis', message = FALSE}
library(BiocStyle)
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

**Last modified:** `r file.info("isotope-intensity-estimation.Rmd")$mtime`<br />
**Compiled**: `r date()`

# Introduction

In this document we evaluate possibilities to estimate abundances for isotopes
without knowing the chemical formula for the compound. To this end we analyze
the element composition of all compounds in the human metabolome database.

The ultimate goal would be to have one or more functions that allow to identify
peaks in (MS1) spectra representing isotopes of the same compound based
exclusively on the m/z and intensity values, i.e. without knowing the chemical
formula of the compound. The functions should be modular and customizable.

# Isotope detection approaches

## CAMERA

CAMERA takes a simple approach based on hard coded lower and upper limits
for isotope peaks. How this limits were defined is unclear. For the m/z
differences of the peaks simple lower and upper limits are used, not
discriminating between e.g. C13 and N15 isotopes.


@Andrea, could you please describe here each other method (ideally the ones from
the papers I sent) briefly?


# Element counts in HMDB

In this section we analyze element compositions of compounds from the Human
Metabolome Database.

```{r libraries, warning = FALSE}
library("CompoundDb")
library("MetaboCoreUtils")
library("pander")
```

We load the HMDB database and extract the related compounds.

```{r}
cdb <- CompDb("data/CompDb.Hsapiens.HMDB.4.0.sqlite")
cmps <- compounds(cdb, columns = c("compound_id", "name", "formula", "exactmass"))
```

For each compound in the human metabolom database (HMDB) we compute how many 
atoms of C, H, N, O, P and S are present in it. 
We collect the counts in a data frame with a column for each element and a row 
for each compound in HMDB.

```{r}
CHNOPS <- c("C", "H", "N", "O", "P", "S", "Cl")
counts <- lapply(cmps$formula, function(frml) {
  res <- countElements(frml)
  row <- data.frame(matrix(0, nrow = 1, ncol = length(CHNOPS)))
  names(row) <- CHNOPS
  only <- intersect(CHNOPS, names(res))
  row[1, only] <- res[only]
  row
  })
counts_df <- do.call(rbind, counts)
rownames(counts_df) <- cmps$compound_id
```


We compute the mean and standard deviation for the number of each element in the 
molecules of HMDB.

```{r}
m_sd <- rbind(colMeans(counts_df), apply(counts_df, 2, sd))
rownames(m_sd) = c("mean", "standard deviation")
pandoc.table(m_sd, style = "rmarkdown")
```

We plot, for each of the considered elements,the distribution of the number of 
atoms of the given element in the compounds of HMDB.

```{r distribution-counts, fig.cap = "distribution of the number of atoms of CHNOPS element in the compounds of HMDB.", fig.width = 7, fig.height = 14}
par(mfrow = c(7, 1), mar = c(3, 4, 1, 0.5))
for (el in CHNOPS) {
    hist(counts_df[, el], xlab = "", main = el, breaks = 256)
}
# It's not clear to me why particularly for H the number of compounds  
# oscillates between odd and even numbers or almost seems like 
# there are two different behaviours
```

We next evaluate what the relationship between the compound's mass and its
number of specific atoms. We could use this later to determine a mass-dependent
factor to estimate the expected intensity of a certain isotope.


Idea:
- for each atom: plot compound masses against number of atoms, fit a linear
  model and get the coefficients from that model.
- for a given mass (m/z) of a peak we can use these to estimate the "expected
  average" number of e.g. C atoms in the formula of this unknown compound and
  use that to guess what the intensity of the isotope peak would be.

We define the absolute abundances of the isotopes for C, N, O, S, Cl.

```{r}
isotopes <- list(C = c(0.9893, 0.0107), 
                 N = c(0.99636, 0.00364), 
                 O = c(0.99757, 0.00038, 0.00205),
                 S = c(0.9499, 0.0075, 0.0425),
                 Cl = c(0.7576, 0.2424))
```

We compute for each molecule in HMDB the probability that the molecule has 
certain numbers of the isotopes of a given element. 

```{r}
pC13 <- sapply(counts_df$C, 
               function(n) ifelse(n > 0, 
                                  dmultinom(c(n - 1, 1), prob = isotopes$C), 
                                  0))
p2C13 <- sapply(counts_df$C,
                function(n) ifelse(n > 1, 
                                   dmultinom(c(n - 2, 2), prob = isotopes$C), 
                                   0))
pN15 <- sapply(counts_df$N, 
               function(n) ifelse(n > 0, 
                                  dmultinom(c(n - 1, 1), prob = isotopes$N), 
                                  0))
pO17 <- sapply(counts_df$O, 
               function(n) ifelse(n > 0, 
                                  dmultinom(c(n - 1, 1, 0), prob = isotopes$O), 
                                  0))
pO18 <- sapply(counts_df$O, 
               function(n) ifelse(n > 0, 
                                  dmultinom(c(n - 1, 0, 1), prob = isotopes$O), 
                                  0))
pS33 <- sapply(counts_df$S, 
               function(n) ifelse(n > 0, 
                                  dmultinom(c(n - 1, 1, 0), prob = isotopes$S), 
                                  0))
pS34 <- sapply(counts_df$S, 
               function(n) ifelse(n > 0, 
                                  dmultinom(c(n - 1, 0, 1), prob = isotopes$S), 
                                  0))
pCl37 <- sapply(counts_df$Cl, 
                function(n) ifelse(n > 0, 
                                   dmultinom(c(n - 1, 1), prob = isotopes$Cl), 
                                   0))
```

```{r}
#pC13_ <- sapply(counts_df$C, function(n) isotopes$C[2]*n)
# p2S34 <- sapply(counts_df$S, 
#                function(n) ifelse(n > 1, 
#                                   dmultinom(c(n - 2, 0, 2), prob = isotopes$S), 
#                                   0))
# boxplot(p2S34[p2S34>0]); quantile(p2S34[p2S34>0], 0.8)
```

```{r, fig.width = 7, fig.height = 20}
names <- c("C13", "2C13", "N15", "O17", "O18", "S33", "S34", "Cl37")
par(mfrow = c(8, 2))
for (name in names)
{
  namev <- paste0("p", name)
  v <- get(namev)[get(namev) > 0]
  hist(v, probability = TRUE, main = name, xlab = namev)
  boxplot(v, main = name)
  #print(paste0(name," : ", quantile(v, 0.8)))
}
```

Open questions/discussion

- [ ] Instead of considering the isotopes separately (e.g. C13, N15 etc), should
      we also *aggregate* them in a M+1, M+2 etc where M+1 would be all atoms
      for which the isotope has a m/z difference of ~ 1?
- [ ] Can we use the information from HMDB on average counts/proportions of
      elements in compounds to get a lower/upper estimate of the isotope
      intensity?





# Session information

The R version and packages used in this analysis are listed below.

```{r sessioninfo}
sessionInfo()
```
