---
title: "Evaluation of the isotoploogues function using the *standards* data set"
author: "Andrea Vicini, Johannes Rainer"
output:
  rmarkdown::html_document:
    highlight: pygments
    toc: true
    toc_depth: 3
    fig_width: 5
---

```{r style, echo = FALSE, results = 'asis', message = FALSE}
library(BiocStyle)
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

**Last modified:** `r file.info("isotopologues-standards.Rmd")$mtime`<br />
**Compiled**: `r date()`

```{r, echo = FALSE}
library(xcms)
library(MetaboCoreUtils)
library(Rdisop)
source("~/lcms-standards/R/writetoms.R")
```

We load the data, a table with the standards and a table (`stab`) containing the
mapping of features to adducts of standards (found in
preprocessing-standards.Rmd).

```{r}
load("data/RData/preprocessing-standards/mix01/data_refined.RData")
load("data/RData/preprocessing-standards/mix01/stab.RData")
std_dil <- read.table("data/standards_dilution.txt", sep= "\t", header = TRUE)
stab$formula <- std_dil[match(stab$name, std_dil$name), "formula"]
```

We filter the data keeping only MS1 data, get the file names and the names of 
the standards
```{r}
dataMS1 <- filterFile(data, file = which(data$mode == "FS"))
mzML_names <- dataMS1$mzML 
std_names <- std_dil[std_dil$mix == 1, "name"]
```

For every mapping in `stab` (of a feature to an adduct of a standard) and for 
every mzML file we consider the spectrum corresponding to the retention time of
the feature. Then we use the `isotopologues` function with parameter `seedMz` 
equal to the mz of the adduct in the mapping. That way we expect to find the
peaks originated by the adduct. Finally we see which chemical formula SIRIUS 
predicts from these peaks and if the formula of the adduct is within the 
top 3 formulas returned by SIRIUS.
Currently, SIRIUS doesn't support adducts with charge higher than 1 or with
2M, 3M etc. and skips them during import. Because of that we remove the mappings
involving these elements.

```{r}
std2ft <- stab[!grepl("\\dM+|]\\d+", stab$adduct), ]
```

We prepare a table summarizing the results.
```{r}
TAB <- data.frame(std2ft[rep(seq_len(nrow(std2ft)), each = length(mzML_names)), 
                         c("name", "adduct", "formula", "feature_id", "rtmed")],
                  mzML = rep(mzML_names, nrow(std2ft)),
                  number_peaks = 0,
                  sirius_best_hit = NA,
                  sirius_top_3 = FALSE)
```

Next we create the .ms files to pass to SIRIUS and compute the number of peaks 
for each file.
```{r}
fl_path <- "inputdir"
dir.create(fl_path, showWarnings = FALSE, recursive = TRUE)

for (i in seq_len(nrow(TAB))) {
  sp <- spectra(filterRt(filterFile(dataMS1, TAB[i, "mzML"]), 
                         TAB[i, "rtmed"] + c(0, 1)))
  pm <- cbind(mz = sp[[1]]@mz, intensity = sp[[1]]@intensity)
  adduct <- TAB[i, "adduct"]
  mzadduct <- mass2mz(getMolecule(TAB[i, "formula"])$exactmass, adduct)[1, 1]
  iso_groups <- isotopologues(pm, seedMz = mzadduct, ppm = 10) # or should I 
  # use positive/ negative definition according to the mode of the adduct? 
  if(length(iso_groups)) {
    fl <- paste0(fl_path, "/", i, ".ms")
    ms1 <- pm[iso_groups[[1]], ]
    TAB[i, "number_peaks"] <- length(iso_groups[[1]])
    writetoms(outputms = fl, ms1 = ms1, compound = TAB[i, "name"], ion = adduct)
  }
}
```

We specify the directory where to save the results from SIRIUS and call it.
```{r, echo = FALSE}
outputdir <- "outputdir"
if(dir.exists(outputdir))
  unlink(outputdir, recursive = TRUE)
dir.create(outputdir)

PATH_TO_SIRIUS <- '/Applications/sirius.app/Contents/MacOS/sirius'
system2(PATH_TO_SIRIUS, args = c('--allow-ms1-only', '--input', "inputdir",
                                 '--output', outputdir, 'formula'))

```

We read the results obtained with SIRIUS and fill the above table .
```{r}
dirs <- list.dirs(outputdir, recursive = FALSE)

for (dir in dirs) {
  outputfile <- paste0(dir, "/formula_candidates.tsv")
  i <- as.numeric(strsplit(dir, "_")[[1]][2])
  if (file.exists(outputfile)) {
    ranking <- read.table(outputfile, sep = "\t",
                          header = TRUE)[, "molecularFormula"]
    if(TAB[i, "formula"] %in% head(ranking, 3))
      TAB[i, "sirius_top_3"] <- TRUE
    TAB[i, "sirius_best_hit"] <- ranking[1]
  }
}
```

```{r , results = "asis", echo = FALSE}
library(pander)
pandoc.table(TAB, style = "rmarkdown")
```

# Session information

The R version and packages used in this analysis are listed below.

```{r sessioninfo}
sessionInfo()
```
