# Introduction

Mixes of standards have been solved in water or added to human serum sample
pools in two different concentration and these samples were measured with the
LC-MS setup from Eurac (used also to generate the CHRIS untargeted metabolomics
data). Assignment of retention time and observed ion can be performed for each
standard based on MS1 information such as expected m/z but also based on the
expected difference in signal between samples with low and high concentration of
the standards. Finally, experimental fragment spectra provide the third level of
evidence. Thus, the present data set allows to annotate features to standards
based on 3 levels of evidence: expected m/z, difference in measured intensity
and MS2-based annotation.

The goal of this analysis is to create an in-house reference library for
our untargeted LC-MS setup. For that we:

- determine the retention time of each standard in water and serum.
- define which ions/adducts are measured for each standard (with relative
  abundances, i.e. which is most highly abundant ion etc).
- extract all MS2 spectra for each standard, match them against reference
  spectra and store them in a database.

The approach consists of the following steps:

- **Identification of potential features for a certain standard**. Features
  matching pre-defined adducts are identified. These are further sub-setted
  keeping only features which match the expected difference in intensities
  between *high* and *low* concentration samples. Resulting features are in
  addition grouped based in similar retention time (5 seconds) and similar
  abundance across samples (correlation > 0.6).
- **MS2 spectra of all selected features are matched against reference spectra
  for the standard from HMDB**. If no match is found and the ion is different
  than `[M+H]+` or `[M-H]-` a comparison between neutral loss spectra is
  performed in addition.
- **MS2 spectra of all selected features are matched against all HMDB and
  Massbank**. This is to evaluate specificity of the fragment spectra.

Based on the available information retention times/features are annotated to
standards using different confidence levels:

- **D** (lowest confidence): based on evidence 1: signal needs to be
  higher in samples with higher concentration of the standard. Measured m/z has
  to match the m/z of an ion of the compound.
- **C**: based on evidence 1 and 2. In addition to **D**, also the MS2
  spectrum for that feature needs to match with at least a low similarity to the
  reference spectra for that compound from HMDB.
- **B**: same as **C** but similarity to a reference spectrum for the standard
  is higher than 0.7.
- **A** (highest confidence): same as **B**, but the experimental MS2 spectrum
  does not match a MS2 spectrum from any other compound in HMDB or MassBank with
  a similarity > 0.7.

These confidence levels are for the annotation of the feature(s) to the
standard. Features with similar retention times and abundances across samples
(as well as eventually similar peak shapes) *inherit* the highest confidence
level from other features but get assigned a **-** to the inherited
confidence. **D** confidence levels are only given if there is a single feature
or for features with retention times similar to higher confidence features from
the other polarity or other matrix.

## Generation of the reference database

The lab-internal reference database will then be created based on the manually
defined and selected features and MS2 spectra. Feature data will be added as
*ion* information to the database. Features detected in water samples will be
preferred, but if a certain ion was only detected in serum or if the retention
times between serum and water features differs by more than 10 seconds they will
be in addition also added. All (manually selected) MS2 spectra (whether in water
or in serum) will be added.

Information from features stored in the database are:

- `original_sample`: the name of the sample mix.
- `sample_matrix`: either water or serum pool.
- `ion_adduct`: the adduct definition.
- `relative_ion_intensity`: the relative intensity of the ion (relative to all
  other ions of the same compound in the same sample matrix).
- `polarity`: the polarity; negative or positive.
- `compound_id`: the (HMDB) ID of the compound.
- `ion_mz`: the (measured) m/z of the feature (`mzmed`).
- `ion_rt`: the rounded, measured retention time of the feature (`rtmed`).
- `confidence_level`: the manually defined confidence level.
- `feature_id`: the feature ID.

Information from MS2 spectra stored on the database are:

- `original_file`: the name of the mzML file. This information should allow to
  link the spectrum to the sample mix etc.
- `adduct`: the adduct definition.
- `compound_id`: the (HMDB) ID of the compound.
- `acquisitionNum`: the spectrum ID/index within the mzML file.
- `polarity`: the polarity.
- `rtime`: the actual retention time of the spectrum.
- `precursorMz`: the precursor m/z.
- `collision_energy`: the collision energy.
- `instrument`: the MS instrument (`"Sciex TripleTOF 5600+"`).
- `instrument_type`: the setup (`"LC-ESI-QTOF"`).
- `confidence`: either `"low"` (no match against reference spectrum) or `"high"`
  (matches reference spectrum).
  

# TEMP

What to store from features. Add features predominantly from water. If none are
found there, or if a certain adduct was not found there, add from serum. Avoid
duplicated ion/adduct information for the same polarity (i.e. a single ion
should be added for a compound).

- feature_id
- original_sample: mix 01...
- sample_matrix: water|serum pool
- ion_adduct
- relative_ion_intensity: relative to the other ions [0-100]
- confidence_level: ABCD.
- rtmed -> ion_mz
- mzmed -> ion_rt
- compound_id
- polarity

```{r eval = FALSE}
## Create a data.frame with that info.
fts <- data.frame(feature_id = ,
                  compound_id,
                  ion_adduct,
                  sample_matrix,
                  original_sample,
                  confidence_level)
## Extract remaining variables from the features:
ion_mz, ion_rt, relative_ion_intensity, 
## Set relative ion intensity

idb <- insertIon(idb, fts)
```

What to store from spectra; add all spectra from water (should be the *purest*).

- confidence: high, low
- adduct
- rtime
- precursor m/z
- acquisitionNum
- polarity
- original_file
- collision_energy
- compound_id
- instrument "Sciex TripleTOF 5600+"
- instrument_type "LC-ESI-QTOF"

```{r eval = FALSE}
## Manually define the spctrum names/IDs.
sps_insert <- data.frame(original_spectrum_id,
                         compound_id,
                         confidence,
                         adduct,
                         instrument,
                         instrument_type)
## Get remaining from Spectra
rtime, precursor_mz, acquisitionNum, polarity, original_file, collision_energy
## Add the ion_id info by getting the just inserted ion info from the database
## identify the entries with feature_id, original_sample and sample_matrix
idb <- insertSpectra(sps, columns = c(the columns to insert))

```

```{r eval = FALSE}
## create a new IonDb from the CompDb. If file exists, remove all ions and
## spectra for the current mix.
```
