---
title: "Identifying mix01 standards in serum"
author: "Andrea Vicini, Vinicius Verri Hernandes, Johannes Rainer"
output:
  rmarkdown::html_document:
    highlight: pygments
    toc: true
    toc_float: true
    toc_depth: 3
    fig_width: 5
---


```{r, include = FALSE, cached = FALSE}
knitr::read_chunk("R/match-standards-chunks.R")
MIX <- 1
MATRIX <- "Serum"
POLARITY <- "POS"
```

```{r, libraries, echo = FALSE, message = FALSE}
```
```{r, general-settings, echo = FALSE, message = FALSE}
```

**Current version: 0.9.0, 2022-03-09.**

# Introduction

Mixes of standards have been solved in water or added to human serum sample
pools in two different concentration and these samples were measured with the
LC-MS setup from Eurac (used also to generate the CHRIS untargeted metabolomics
data). Assignment of retention time and observed ion can be performed for each
standard based on MS1 information such as expected m/z but also based on the
expected difference in signal between samples with low and high concentration of
the standards. Finally, experimental fragment spectra provide the third level of
evidence. Thus, the present data set allows to annotate features to standards
based on 3 levels of evidence: expected m/z, difference in measured intensity
and MS2-based annotation.

A detailed description of the approach is given in [match-standards-introduction.Rmd](match-standards-introduction.Rmd).

The list of standards that constitute the present sample mix are listed below
along with the expected retention time and the most abundant adduct for positive
and negative polarity as defined manually in a previous analysis by Mar
Garcia-Aloy.

```{r, standards-table, echo = FALSE, results = "asis"}
```

# Data import

We next load the data and split it according to matrix and polarity.

```{r, data-import}
```

We next also load the reference databases we will use to compare the
experimental MS2 spectra against. Below we thus load HMDB and MassBank data. We
create in addition *neutral loss* versions of the databases. Since HMDB does not
provide precusor m/z values we assume the fragment spectra to represent `[M+H]+`
ions (respectively `[M-H]-` ions for negative polarity) and add the m/z for
these adducts as *precursor m/z*.

```{r, load-reference-databases, message = FALSE}
```


# Initial evaluation of available MS2 data

Before performing the actual analysis that involves chromatographic peak
detection and evaluation of feature abundances, we compare all experimental MS2
spectra against the reference libraries. This provides already a first hint for
which standards a valid MS2 spectrum was recorded (and hence would allow
annotation based on evidence 3) and what related retention time might be.

We first extract all MS2 spectra from the respective data files and process them
by first removing all peaks with an intensity below 5% of the highest peak
signal, then scaling all intensities to a value range between 0 and 100 and
finally remove all MS2 spectra with less than 2 peaks.

```{r prepare-all-ms2}
fls <- fileNames(data_all)[data_all$polarity == "POS"]
```
```{r, all-ms2}
```

We next match these spectra against all reference spectra from HMDB for the
standards in `r MIX_NAME`. 

The settings for the matching are defined below. These will be used for all MS2
spectra similarity calculations. All matching spectra with a similarity larger
0.7 are identified.

```{r, compare-spectra-param}
```
 
Standards for which no MS2 spectrum in HMDB is present are listed in
the table below.

```{r, echo = FALSE, results = "asis"}
tmp <- std_dilution01[!(std_dilution01$HMDB %in% hmdb_std$compound_id), ]
pandoc.table(tmp[, c("name", "HMDB", "formula")], style = "rmarkdown",
             caption = "Standards for which no reference spectrum is available")
```

The table below lists all standards and the number of matching reference spectra
(along with their retention times).

```{r, table-all-ms2-hmdb, echo = FALSE, results = "asis"}
```

We repeat the same analysis for the negative polarity data (code not shown
again).

```{r, echo = FALSE}
fls <- fileNames(data_all)[data_all$polarity == "NEG"]
```
```{r, all-ms2, echo = FALSE}
```
```{r, table-all-ms2-hmdb, echo = FALSE, results = "asis"}
```

We next perform the full analysis of the data set involving MS1 and MS2 data.


# Serum, positive polarity

We now perform the analysis on the samples with the standards solved in
serum acquired in positive polarity mode.

```{r}
POLARITY <- "POS"
data <- filterFile(data_all, which(data_all$polarity == POLARITY))
hmdb <- hmdb_pos
hmdb_nl <- hmdb_pos_nl
```

## Data pre-processing

We perform the pre-processing of our data set which consists of the
chromatographic peak detection followed by a peak refinement step to reduce peak
detection artifacts, the correspondence analysis to group peaks across samples
and finally the gap-filling to fill in missing peak data from samples in which
no chromatographic peak was detected.

Some notes on the settings for the pre-processing:

- The `bw` parameter for the correspondence is larger than usual because adding
  the standards in higher concentrations caused considerable retention time
  shifts for some. A higher `bw` will allow to group also mis-aligned
  chromatographic peaks - but will not allow to discriminate between closely
  eluting ions.
- The `binSize` was also slightly increased to avoid splitting of features with
  similar m/z values.

```{r, preprocessing, eval = !file.exists(paste0(RDATA_PATH, "processed_data_", POLARITY, ".RData")), message = FALSE}
```

```{r, echo = FALSE}
load(paste0(RDATA_PATH, paste0("processed_data_", POLARITY, ".RData")))
data_FS <- filterFile(data, which(data$mode == "FS"),
                      keepFeatures = TRUE)
```

## Signal intensity difference

We compute the difference in (log2) signals between samples with high and low
concentration of the standards and calculate the p-value for this difference
using the Student's t-test.

```{r, abundance-difference}
```

## Identification of features matching standards

We now identify all features matching the any of the pre-defined set of adducts
for the standards of mix `r MIX`. Matching features are further subsetted to
those that show an at least twice as high signal in samples with high
concentration compared to those with low concentrations. Also, features with a
signal in low concentration but no detectable signal in high concentration are
removed.

```{r, define-adducts-pos}
```
```{r, match-features}
```

For most of the standards (`r length(unique(mD$target_name))` out of 
`r nrow(std_dilution01)`) in mix `r MIX` at least one feature was found
matching the standards adducts' m/z. 

```{r, table-standard-no-feature, results = "asis", echo = FALSE}
```

In the next sections we investigate for each standard which assignment would be
the correct one or, for those for which no signal was detected, why that was the
case.


## Standards with matching features

While for some standards a matching feature was found we still need to evaluate
whether this matching is correct. For each standard we thus first evaluate the
EICs for all matching features, then we match their MS2 spectra (if available)
against reference libraries. To ensure correct assignment of a feature, its
retention time and eventually related MS2 spectra, we consider the following
criteria to determine the annotation confidence:

- feature(s) was/were found with m/z matching those of adduct(s) of the 
  standard.
- signal is higher for samples with higher concentration.
- MS2 spectra matches reference spectra for the standard (if available).
- MS2 spectra don't match MS2 spectra of other compounds.


### 3-Phosphoglyceric Acid

```{r, echo = FALSE}
std <- "3-Phosphoglyceric Acid"
```

```{r, table-feature-matches, results = "asis", echo = FALSE, message = FALSE}
```

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SP")
```

A considerable number of features has been assigned to this standard based on
m/z matching. Based on their retention times, these seem however to represent
ions from different compounds.

The cleaned MS2 spectra for the features matched to `r std` are shown below
(peaks with an intensity below 5% of the maximum peak and spectra with less
than 2 peaks where removed).

```{r mix01-serum-3-phosphoglyceric-acid-ms2, fig.cap = "MS2 spectra for features matched to 3-Phosphoglyceric Acid.", echo = FALSE, eval = length(std_ms2) > 0}
par(mar = c(4.2, 4.5, 1.5, 0.5))
xl <- range(unlist(mz(std_ms2)))
plotSpectra(std_ms2, xlim = xl)
```

We next match the extracted MS2 spectra against the reference spectra for 
`r std` from HMDB.

```{r, plot-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for 3-Phoshoglyceric Acid from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

It seems that no extracted MS2 spectra matches well the reference ones from
3-Phosphoglyceric Acid. The maximum score is around 0.08 and is attained by a
spectrum from FT02217. The only two matches with score greater than 0.07 are
shown below.

```{r mix01-serum-3-phosphoglyceric-acid-mirror-hmdb, echo = FALSE, fig.cap = "Mirror plots"}
std_ms2_sel <- plot_select_ms2(std_ms2, std_hmdb, 0.07, ppm = csp@ppm, tolerance = csp@tolerance)
```

The spectra don't seem to match well.

In addition we compare all spectra against reference spectra from MassBank for
the selected standard. 

```{r, plot-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for 3-Phoshoglyceric Acid from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

Since all the extracted MS2 spectra are associated to ions different from
[M+H]+ we also perform a comparison between the neutral loss spectra from HMDB
and MassBank.

```{r, plot-ms2-hmdb-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for 3-Phoshoglyceric Acid from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r, plot-ms2-mbank-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for 3-Phoshoglyceric Acid from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

No match was found. In addition we match (**all**) the MS2 spectra for the
matched features against all spectra from HMDB or MassBank identifying reference
spectra with a similarity larger than 0.7. The results (if any spectra matched)
are shown in the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

For completeness, we also perform a neutral loss spectra comparison with HMDB
and MassBank.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2_nl, hmdb_nl, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, mbank_nl, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

#### Summary

- Annotation of FT02217 (RT=250, `[M+Na]+` ion) to `r std` with confidence level
  **C**? Or **D** because the score of the match between MS2 spectra was close
  to 0? 
- FT02579 was grouped together FT02217 so it should get the same confidence
  level? 
- Reference MS2: F08.S0419 (with a very *low* grading?)
- What about the other features. Should we give them confidence **D** or can
  we exclude them with other considerations?


### Acetylhistidine

```{r, echo = FALSE}
std <- "Acetylhistidine"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SP")
```

The cleaned MS2 spectra for the features matched to `r std` are shown below
(peaks with an intensity below 5% of the maximum peak and spectra with less
than 2 peaks where removed).

```{r mix01-serum-acetylhistidine-ms2, fig.cap = "MS2 spectra for features matched to acetylhistidine.", echo = FALSE}
par(mar = c(4.2, 4.5, 1.5, 0.5))
xl <- range(unlist(mz(std_ms2)))
plotSpectra(std_ms2, xlim = xl)
```

Next we compare them against the reference spectra for `r std` from HMDB.
The results are shown in the heatmap below.

```{r, plot-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for Acetylhistidine from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

Some of the MS2 spectra for FT02018 and FT02021 have a high similarity against
MS2 spectra from `r std`. Mirror plots for the best matching spectra are shown
below.

```{r mix01-serum-acetylhistidine-mirror-hmdb, echo = TRUE, fig.cap = "Mirror plots"}
std_ms2_sel <- plot_select_ms2(std_ms2, std_hmdb, 0.5,
                               ppm = csp@ppm, tolerance = csp@tolerance)
```

These matches are very nice. In addition we perform also the match against
reference spectra for that standard from MassBank.

```{r, plot-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for acetylhistidine from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

Surprisingly no spectrum matches MassBank.

FT02728 and FT02370 were considered as from [M+2Na-H]+ and [M+Na]+
respectively. We check if matches beetween neutral loss spectra would be found.

```{r, plot-ms2-hmdb-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for acetylhistidine from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

The same spectra match also with their neutral loss versions.

```{r, plot-ms2-mbank-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for acetylhistidine from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

One spectrum seems to match.

```{r mix01-serum-acetylhistidine-mirror-massbank-nl, echo = TRUE, fig.cap = "Mirror plots"}
tmp_res <- plot_select_ms2(std_ms2_nl, std_mbank_nl, 0.5,
                           ppm = csp_nl@ppm, tolerance = csp_nl@tolerance)
```

This spectrum matches however with a single peak only (the precursor) while the
fragment peaks differ.

In addition we match (**all**) the MS2 spectra for the matched features against
all spectra from HMDB or MassBank identifying reference spectra with a
similarity larger than 0.7. The results (if any spectra matched) are shown in
the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

For completeness, we also perform a neutral loss spectra comparison with HMDB
and MassBank.

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, hmdb_nl, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, mbank_nl,
              sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```


#### Summary

- FT02018 (RT=182.5, `[M+H]+` ion) matches MS2 spectra of `r std` but also
  those of others: confidence **B**.
- FT02021 (RT=198.1, `[M+H]+` ion); confidence **B**.
- Reference MS2: (F07.S0638?), (F07.S0810?), F08.S0632, F08.S0806.


### Betaine

```{r, echo = FALSE}
std <- "Betaine"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SP")
```

The cleaned MS2 spectra for the features matched to `r std` are shown below
(peaks with an intensity below 5% of the maximum peak and spectra with less
than 2 peaks where removed).

```{r mix01-serum-betaine-ms2, fig.cap = "MS2 spectra for features matched to betaine.", echo = FALSE}
par(mar = c(4.2, 4.5, 1.5, 0.5))
xl <- range(unlist(mz(std_ms2)))
plotSpectra(std_ms2, xlim = xl)
```

Next we compare them against the reference spectra for `r std` from HMDB.
The results are shown in the heatmap below.

```{r, plot-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for betaine from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

MS2 spectra for FT00638 (F07.S0573, F07.S0602, F08.S0601) and FT00642
(F07.S0573, F07.S0602, F08.S0601, in fact the same) match the reference spectra
from Betaine with high scores. From the EIC FT0522 seems to be mostly noise and
no *real* compound data.

```{r mix01-serum-betaine-mirror-hmdb, echo = TRUE, fig.cap = "Mirror plots"}
std_ms2_sel <- plot_select_ms2(std_ms2, std_hmdb, 0.7,
                               ppm = csp@ppm, tolerance = csp@tolerance)
```

In addition we compare against MassBank spectra.

```{r, plot-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for betaine from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

The same MS2 spectra match, albeit with a lower similarity.

We check in addition if matches between neutral loss spectra would be found.

```{r, plot-ms2-hmdb-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for Betaine from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

Neutral loss spectra don't match because there seems to be an offset (by 1.008)
in the precursor m/z values.

```{r, plot-ms2-mbank-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for Betaine from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

In all these cases only the 0 m/z peaks was matching, but no other peak.

In addition we match (**all**) the MS2 spectra for the matched features against
all spectra from HMDB or MassBank identifying reference spectra with a
similarity larger than 0.7. The results (if any spectra matched) are shown in
the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

For completeness, we also perform a neutral loss spectra comparison with HMDB
and MassBank (requiring a high similarity).

```{r, echo = FALSE, message = FALSE, results = "asis", eval = TRUE}
std_ms2_nl <- neutralLoss(std_ms2, PrecursorMzParam())
perform_match(std_ms2_nl, hmdb_nl, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl, similarity = 0.9)
```

```{r, echo = FALSE, message = FALSE, results = "asis", eval = TRUE}
perform_match(std_ms2_nl, mbank_nl, similarity = 0.9,
              sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```


#### Summary

- MS2 spectra for FT00638 (RT=163, [M+H]+ ion) matches Betaine with confidence B.
- The same for FT00642.
- Reference MS2: F08.S0573, F07.S0602, F07.S0601.


### C3 Carnitine

```{r, echo = FALSE}
std <- "C3 Carnitine"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SP")
```

The cleaned MS2 spectra for the features matched to `r std` are shown below
(peaks with an intensity below 5% of the maximum peak and spectra with less
than 2 peaks where removed).

```{r mix01-serum-C3-carnitine-ms2, fig.cap = "MS2 spectra for features matched to C3 Carnitine.", echo = FALSE}
par(mar = c(4.2, 4.5, 1.5, 0.5))
xl <- range(unlist(mz(std_ms2)))
plotSpectra(std_ms2, xlim = xl)
```

Next we compare them against the reference spectra for `r std` from HMDB.
The results are shown in the heatmap below.

```{r, plot-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for C3 Carnitine from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r, plot-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for C3 Carnitine from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

No reference spectrum is available in HMDB or MassBank.

In addition we match (**all**) the MS2 spectra for the matched features against
all spectra from HMDB or MassBank identifying reference spectra with a
similarity larger than 0.7. The results (if any spectra matched) are shown in
the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

For completeness, we also perform a neutral loss spectra comparison with HMDB
and MassBank.

```{r, echo = FALSE, message = FALSE, results = "asis", eval = TRUE}
std_ms2_nl <- neutralLoss(std_ms2, PrecursorMzParam())
perform_match(std_ms2_nl, hmdb_nl, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl, similarity = 0.7)
```

```{r, echo = FALSE, message = FALSE, results = "asis", eval = TRUE}
perform_match(std_ms2_nl, mbank_nl, similarity = 0.7,
              sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

#### Summary

- FT02650 (RT=162.6, `[M+NH4]+`): confidence level **D**? Maybe EICs for 
  FT02063, FT02649, FT03045 are not that good.


### CDP

```{r, echo = FALSE}
std <- "CDP"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SP")
```

The cleaned MS2 spectra for the features matched to `r std` are shown below
(peaks with an intensity below 5% of the maximum peak and spectra with less
than 2 peaks where removed).

```{r mix01-serum-cdp-ms2, fig.cap = "MS2 spectra for features matched to CDP.", echo = FALSE}
par(mar = c(4.2, 4.5, 1.5, 0.5))
xl <- range(unlist(mz(std_ms2)))
plotSpectra(std_ms2, xlim = xl)
```

Next we compare them against the reference spectra for `r std` from HMDB.
The results are shown in the heatmap below.

```{r, plot-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for CDP from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

Both spectra (F07.S0864 and F07.S0892) of feature FT06278 match to one
reference spectrum of CDP with high similarity.

```{r mix01-serum-cdp-mirror-hmdb, echo = TRUE, fig.cap = "Mirror plots"}
std_ms2_sel <- plot_select_ms2(std_ms2, std_hmdb, 0.7,
                               ppm = csp@ppm, tolerance = csp@tolerance)
```

Unfortunately, the match based on a single peak.

In addition we compare against MassBank spectra.

```{r, plot-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for CDP from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

No match against MassBank can be found.

We check in addition if matches between neutral loss spectra would be found.

```{r, plot-ms2-hmdb-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for CDP from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

Not unexpectedly, the same spectra match also with their neutral losses.

```{r, plot-ms2-mbank-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for CDP from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

We don't have a good match against MassBank.

In addition we match (**all**) the MS2 spectra for the matched features against
all spectra from HMDB or MassBank identifying reference spectra with a
similarity larger than 0.7. The results (if any spectra matched) are shown in
the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

For completeness, we also perform a neutral loss spectra comparison with HMDB
and MassBank (requiring a high similarity).

```{r, echo = FALSE, message = FALSE, results = "asis", eval = TRUE}
std_ms2_nl <- neutralLoss(std_ms2, PrecursorMzParam())
perform_match(std_ms2_nl, hmdb_nl, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl, similarity = 0.9)
```

```{r, echo = FALSE, message = FALSE, results = "asis", eval = TRUE}
perform_match(std_ms2_nl, mbank_nl, similarity = 0.9,
              sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

#### Summary

- FT06278 (RT=260.8, `[M+H]+`): confidence level **B**.
- Reference spectra: F23.S0892, F23.S0864?


### Creatine

```{r, echo = FALSE}
std <- "Creatine"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SP")
```

The cleaned MS2 spectra for the features matched to `r std` are shown below
(peaks with an intensity below 5% of the maximum peak and spectra with less
than 2 peaks where removed).

```{r mix01-serum-creatine-ms2, fig.cap = "MS2 spectra for features matched to creatine.", echo = FALSE}
par(mar = c(4.2, 4.5, 1.5, 0.5))
xl <- range(unlist(mz(std_ms2)))
plotSpectra(std_ms2, xlim = xl)
```

Next we compare them against the reference spectra for `r std` from HMDB.
The results are shown in the heatmap below.

```{r, plot-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for creatine from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

MS2 spectra for FT00870 (F07.S0633 and F08.S0630) match with a low similarity
against reference spectra from HMDB. Mirror plots for these are shown below.

```{r mix01-serum-creatine-mirror-hmdb, echo = TRUE, fig.cap = "Mirror plots"}
std_ms2_sel <- plot_select_ms2(std_ms2, std_hmdb, 0.2,
                               ppm = csp@ppm, tolerance = csp@tolerance)
```

In addition we compare against MassBank spectra.

```{r, plot-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for creatine from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

Also here there is a single match with a low similarity.

We check in addition if matches between neutral loss spectra would be found.

```{r, plot-ms2-hmdb-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for creatine from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

Neutral loss version of MS2 spectrum F23.S0621 for feature FT01667 matches the
(neutral loss) reference spectrum for creatine from MassBank (see below). The
same is true for the spectra from FT00575, FT00578 and FT00577.

```{r, plot-ms2-mbank-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for creatine from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

In addition we match (**all**) the MS2 spectra for the matched features against
all spectra from HMDB or MassBank identifying reference spectra with a
similarity larger than 0.7. The results (if any spectra matched) are shown in
the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

For completeness, we also perform a neutral loss spectra comparison with HMDB
and MassBank (requiring a high similarity).

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
std_ms2_nl <- neutralLoss(std_ms2, PrecursorMzParam())
perform_match(std_ms2_nl, hmdb_nl, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl, similarity = 0.9)
```

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, mbank_nl, similarity = 0.9,
              sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

#### Summary

- MS2 spectra for FT00578, FT00577 and FT00575 match spectra from creatinine.
- FT00870 (RT=176, `[M+H]+` ion): confidence **C**: the MS2 spectrum F07.S0633
  matches only `r std` with a low similarity.
- FT01667 (RT=173.3, `[M+2Ma-H]+` ion): neutral loss spectrum of F07.S0621
  matches reference spectrum with high simlarity. Confidence **B**?.
- FT01307 inherit confidence from above.
- Reference spectra: F07.S0621 (and F07.S0633?).

### Dimethylglycine

```{r, echo = FALSE}
std <- "Dimethylglycine"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

Based on their retention time, it seems that these features represent signal
from several different compounds.

We next plot the EIC for the assigned feature and visually inspect these.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SP")
```

The only MS2 spectra for the features matched to `r std` are shown below
(peaks with an intensity below 5% of the maximum peak and spectra with less
than 2 peaks where removed).

```{r mix01-serum-dimethylglycine-ms2, fig.cap = "MS2 spectra for features matched to dimethylglycine", echo = FALSE}
par(mar = c(4.2, 4.5, 1.5, 0.5))
xl <- range(unlist(mz(std_ms2)))
plotSpectra(std_ms2, xlim = xl)
```

We next match the extracted MS2 spectrum against the reference spectra for 
`r std` from HMDB.

```{r, plot-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectrum for selected features against reference spectra for Dimethylglycine from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

The spectrum does thus not match any reference spectra from `r std`. In addition
we compare it aginst the reference spectra from MassBank for the selected
standard.

```{r, plot-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectrum for selected features against reference spectra for Dimethylglycine from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

Again, we don't obtain any match. Since the extracted spectrum is associated to
an ion different from [M+H]+ ([M+2Na-H]+) we also perform a comparison between
the neutral loss spectra from HMDB and MassBank.

```{r, plot-ms2-hmdb-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for Dimethylglycine from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r, plot-ms2-mbank-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for Dimethylglycine from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

In the first case we still don't get any match, in the second however we have
some reference spectra that do match with the extracted spectrum. We plot below
the matches with similarity higher than 0.5.

```{r mix01-serum-dimethylglycine-mirror-hmdb, echo = FALSE, fig.cap = "Mirror plots"}
std_ms2_sel <- plot_select_ms2(std_ms2_nl, std_mbank_nl, 0.5, ppm = csp@ppm, tolerance = csp@tolerance)
```

In addition we match (**all**) the MS2 spectra for the
matched features against all spectra from HMDB or MassBank identifying reference
spectra with a similarity larger than 0.7. The results (if any spectra matched)
are shown in the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

For completeness, we also perform a neutral loss spectra comparison with HMDB
and MassBank.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2_nl, hmdb_nl, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, mbank_nl, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

The neutral loss version of the only available MS2 spectra beside `r std`
strangely matches a lot of other reference compounds.

#### Summary

- Should we assign FT01182 (RT = 174.6, `[M+2Na-H]+` ion) confidence level *B*
  because of the neutral loss matches against Mbank?

### Glycerol

```{r, echo = FALSE}
std <- "Glycerol"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

We next plot the EIC for the assigned feature and visually inspect it.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SP")
```


```{r mix01-serum-glycerol-ms2, fig.cap = "MS2 spectra for features matched to glycerol", echo = FALSE}
if (length(std_ms2)) {
    par(mar = c(4.2, 4.5, 1.5, 0.5))
    xl <- range(unlist(mz(std_ms2)))
    plotSpectra(std_ms2, xlim = xl)
} else {
    cat("No MS2 spectra found for", std, "\n")
}
```

Unfortunately no MS2 spectra is available.

#### Summary

- FT00947 with confidence **D**? But its EIC maybe doesn't look that good
- No MS2 spectra

### L-Glutamic Acid

```{r, echo = FALSE}
std <- "L-Glutamic Acid"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

Based on their retention time, it seems that these features represent signal
from several different compounds.

We next plot the EIC for the assigned feature and visually inspect these.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SP")
```

The cleaned MS2 spectra for the features matched to `r std` are shown below
(peaks with an intensity below 5% of the maximum peak and spectra with less
than 2 peaks where removed).

```{r mix01-serum-L-Glutamic-Acid-ms2, fig.cap = "MS2 spectra for features matched to L-Glutamic Acid.", echo = FALSE}
if (length(std_ms2)) {
    par(mar = c(4.2, 4.5, 1.5, 0.5))
    xl <- range(unlist(mz(std_ms2)))
    plotSpectra(std_ms2, xlim = xl)
} else {
    cat("No MS2 spectra found for", std, "\n")
}
```

We have thus MS2 spectra from FT01566 (first two plots), FT01931 (third plot),
FT01189 (last two plots).

Next we compare them against the reference spectra for `r std` from HMDB. The
results are shown in the heatmap below.

```{r, plot-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for L-Glutamic Acid from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

The MS2 spectra for FT00819 seem to have quite high similarity
against HMDB MS2 spectra from `r std`. Mirror plots for the best matching
spectra (similarity higher than 0.9) are shown below.

```{r mix01-serum-L-Glutamic-Acid-mirror-hmdb, echo = TRUE, fig.cap = "Mirror plots"}
std_ms2_sel <- plot_select_ms2(std_ms2, std_hmdb, 0.9, ppm = csp@ppm,
                               tolerance = csp@tolerance)
```

Thus, the best matching spectra are F23.S0695 and F24.S0682.

We also perform the comparison against reference spectra from MassBank and as
before we find matches involving F23.S0695 and F24.S0682 (from FT01189).

```{r, plot-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for L-Glutamic Acid from MassBank", echo = FALSE, fig.width = 5, fig.height = 5}
```


```{r mix01-serum-L-Glutamic-Acid-mirror-mbank, echo = TRUE, fig.cap = "Mirror plots"}
tmp_res <- plot_select_ms2(std_ms2, std_mbank, 0.7, ppm = csp@ppm,
                               tolerance = csp@tolerance)
```
We also perform the comparison using the neutral loss version of the spectra.

```{r, plot-ms2-hmdb-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for L-Glutamic-Acid from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r, plot-ms2-mbank-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for L-Glutamic-Acid from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

No match is found for HMDB (?) while for MassBank F23.S0695 and F24.S0682
(from FT01189) are matched again. Additionally also F23.S0636 (from FT01566) 
matches but with lower score. Below we plot the matches for this last spectrum
with similarity > 0.4.

```{r mix01-serum-L-Glutamic-Acid-mirror-mbank_nl, echo = TRUE, fig.cap = "Mirror plots"}
sim <- sim["FT01566_F23.S0636", ]
tmp_res <- plot_select_ms2(std_ms2_nl["FT01566_F23.S0636"], std_mbank_nl, 0.4,
                           ppm = csp@ppm, tolerance = csp@tolerance)
```

In addition we match (**all**) the MS2 spectra for the matched features against
all spectra from HMDB or MassBank identifying reference spectra with a
similarity larger than 0.7. The results (if any spectra matched) are shown in
the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

For completeness, we also perform a neutral loss spectra comparison with HMDB
and MassBank.

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, hmdb_nl, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, mbank_nl,
              sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```
  
MS2 spectra associated to feature FT0693, FT1018 have been matched to reference
spectra from L-Glutamic Acid but also from other compounds with relatively high
scores in both cases.

No match is found for HMDB (?) while for MassBank F23.S0695 and F24.S0682
(from FT01189) are matched again. Additionally also F23.S0636 (from FT01566) 
matches but with lower score. Below we plot the matches for this last spectrum
with similarity > 0.4.

#### Summary

- FT01189 (RT=182.5, `[M+H]+` ion): confidence level **B**. Its MS2 spectra
  (F23.S0695 and F24.S0682) match `r std`, but also other compounds.
- FT00845 (RT=180.1, `[M+H-NH3]+`] ion): confidence level **B-**.
- FT01566 (RT=179.9, `[M+Na]+`] ion): confidence level **C**?
- Reference MS2 spectra: F23.S0695 and F24.S0682.

### Myo-Inositol

```{r, echo = FALSE}
std <- "Myo-Inositol"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

Based on their retention time, it seems that these features represent signal
from several different compounds.

We next plot the EIC for the assigned feature and visually inspect these.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SP")
```

The cleaned MS2 spectra for the features matched to `r std` are shown below
(peaks with an intensity below 5% of the maximum peak and spectra with less
than 2 peaks where removed).

```{r mix01-serum-myo-inositol-ms2, fig.cap = "MS2 spectra for features matched to myo-inositol.", echo = FALSE}
if (length(std_ms2)) {
    par(mar = c(4.2, 4.5, 1.5, 0.5))
    xl <- range(unlist(mz(std_ms2)))
    plotSpectra(std_ms2, xlim = xl)
} else {
    cat("No MS2 spectra found for", std, "\n")
}
```

Next we compare them against the reference spectra for `r std` from HMDB and
MassBank. The results are shown in the heatmaps below.

```{r, plot-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for myo-inositol from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r, plot-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for myo-inositol from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

None of the extracted spectra matches reference spectra from `r std`. We also
check if there are matches between neutral loss spectra.

```{r, plot-ms2-hmdb-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for myo-inositol from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r, plot-ms2-mbank-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for myo-inositol from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

We don't find any match also in this case. In addition we match (**all**) the
MS2 spectra for the matched features against all spectra from HMDB or MassBank
identifying reference spectra with a similarity larger than 0.7. The results
(if any spectra matched) are shown in the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

For completeness, we also perform a neutral loss spectra comparison with HMDB
and MassBank.

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, hmdb_nl, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, mbank_nl,
              sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

#### Summary

- Extracted spectra don't match any reference spectra
- All features assigned with confidence with confidence **D**? (maybe FT02106,
  FT02108 EICs look better)


### Pyruvic Acid

```{r, echo = FALSE}
std <- "Pyruvic Acid"
```

The table below lists all features that were matched to one of the adducts of 
`r std` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SP")
```

We next look if there are MS2 spectra associated to this feature but we don't
find any.

```{r mix01-serum-pyruvic-acid-ms2, fig.cap = "MS2 spectra for features matched to pyruvic-acid", echo = FALSE}
if (length(std_ms2)) {
    par(mar = c(4.2, 4.5, 1.5, 0.5))
    xl <- range(unlist(mz(std_ms2)))
    plotSpectra(std_ms2, xlim = xl)
} else {
    cat("No MS2 spectra found for", std, "\n")
}
```


#### Summary

- No MS2 spectra available.


### Suberic Acid

```{r, echo = FALSE}
std <- "Suberic Acid"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

Based on their retention time, it seems that these features represent signal
from a few different compounds.

We next plot the EIC for the assigned feature and visually inspect these.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SP")
```

The cleaned MS2 spectra for the features matched to `r std` are shown below
(peaks with an intensity below 5% of the maximum peak and spectra with less
than 2 peaks where removed).

```{r mix01-serum-suberic-acid-ms2, fig.cap = "MS2 spectra for features matched to suberic acid.", echo = FALSE}
par(mar = c(4.2, 4.5, 1.5, 0.5))
xl <- range(unlist(mz(std_ms2)))
plotSpectra(std_ms2, xlim = xl)
```

We have thus MS2 spectra for feature around 140. Next we compare them
against the reference spectra for `r std` from HMDB and MassBank. The results
are shown in the heatmaps below.

```{r, plot-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for suberic acid from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r, plot-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for suberic acid from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

Strangely every match has score equal to 0 in both cases. In addition we
also look for matches between neutral loss spectra.

```{r, plot-ms2-hmdb-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for suberic acid from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

As for HMDB, we find matches for F24.S0524, F23.S0504 (from FT00499) with low
score.

```{r mix01-suberic-acid-mirror-hmdb_nl, echo = TRUE, fig.cap = "Mirror plots"}
tmp_res <- plot_select_ms2(std_ms2_nl, std_hmdb_nl, 0.2,
                               ppm = csp@ppm, tolerance = csp@tolerance)
```

```{r, plot-ms2-mbank-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for suberic acid from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r mix01-suberic-acid-mirror-mbank_nl, echo = TRUE, fig.cap = "Mirror plots"}
tmp_res <- plot_select_ms2(std_ms2_nl, std_mbank_nl, 0.2,
                               ppm = csp@ppm, tolerance = csp@tolerance)
```

In addition we match (**all**) the MS2 spectra for the matched features against
all spectra from HMDB or MassBank identifying reference spectra with a
similarity larger than 0.7. The results (if any spectra matched) are shown in
the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

For completeness, we also perform a neutral loss spectra comparison with HMDB
and MassBank.

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, hmdb_nl, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, mbank_nl,
              sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

The extracted spectra match with high similarity reference spectra 
of other compounds than `r std`.

#### Summary

- FT00499 (RT = 141.3, `[M+2Na]2+` ion) with confidence level  **C** (taking
  into account neutral loss spectra) and related MS2 (F23.S0504, F24.S0524).
- The other matched features with level **D**? Or looking to the EICs exclude
  some (e.g FT02375)? 

### Xanthine

```{r, echo = FALSE}
std <- "Xanthine"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

Based on their retention time, it seems that these features represent signal
from two compounds.

We next plot the EIC for the assigned feature and visually inspect these.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SP")
```

The cleaned MS2 spectra for the features matched to `r std` are shown below
(peaks with an intensity below 5% of the maximum peak and spectra with less
than 2 peaks where removed).

```{r mix01-serum-xanthine-ms2, fig.cap = "MS2 spectra for features matched to xanthine", echo = FALSE}
par(mar = c(4.2, 4.5, 1.5, 0.5))
xl <- range(unlist(mz(std_ms2)))
plotSpectra(std_ms2, xlim = xl)
```

We have thus MS2 spectra for the set of features with retention time around 140.
Next we compare them against the reference spectra for `r std` from HMDB. The
results are shown in the heatmap below.

```{r, plot-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for Xanthine from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

The MS2 spectra for FT01285 (F23.S0505 and F24.S0515) have the highest
similarity against MS2 spectra from `r std`. Mirror plots for the best matching
spectra are shown below.

```{r mix01-serum-xanthine-mirror-hmdb, echo = TRUE, fig.cap = "Mirror plots"}
std_ms2_sel <- plot_select_ms2(std_ms2, std_hmdb, 0.7,
                               ppm = csp@ppm, tolerance = csp@tolerance)
```
We repeat the comparison using MassBank reference spectra and we obtain similar
results as shown below.

```{r, plot-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for Xanthine from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r mix01-serum-xanthine-mirror-mbank, echo = TRUE, fig.cap = "Mirror plots"}
tmp_res <- plot_select_ms2(std_ms2, std_mbank, 0.7,
                               ppm = csp@ppm, tolerance = csp@tolerance)
```

We also check if there are matches between neutral loss spectra.

```{r, plot-ms2-hmdb-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for Xanthine from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r, plot-ms2-mbank-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for Xanthine from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

Based on neutral loss spectra comparison with MassBank also spectra F23.S0525
(and maybe F24.S0532) from feature FT02002 seem to match. 

```{r mix01-serum-xanthine-mirror-massbank-nl, echo = TRUE, fig.cap = "Mirror plots"}
tmp_res <- plot_select_ms2(std_ms2_nl, std_mbank_nl, 0.5,
                           ppm = csp_nl@ppm, tolerance = csp_nl@tolerance)
```
The spectra from FT02002 consit mainly of one peak so maybe the associated
matches are not that reliable.

In addition we match (**all**) the MS2 spectra for the matched features against
all spectra from HMDB or MassBank identifying reference spectra with a
similarity larger than 0.7. The results (if any spectra matched) are shown in
the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

For completeness, we also perform a neutral loss spectra comparison with HMDB
and MassBank.

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, hmdb_nl, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, mbank_nl,
              sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

Based on both HMDB and MassBank the extracted spectra match reference spectra of
compounds different from `r std` with higher similarity.

#### Summary

- FT01285 (RT=141.4, `[M+H]+` ion): confidence level **B**.
- FT00924 (RT=141.2, `[M+H-NH3]+` ion): confidence level **B-**.
- FT01908 (RT=141.2, `[M+K]+` ion): confidence level **B-**.
- FT01645 (RT=141.7, `[M+Na]+` ion): confidence level **B-**.
- FT02002 (RT=142.7, `[M+2Na-H]+` ion): confidence level **B-**.
- Reference MS2: F23.S0505 and F24.S0515 (from FT01285).

### Uric acid

```{r, echo = FALSE}
std <- "Uric acid"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

Based on their retention time, it seems that these features represent signal
from two compounds.

We next plot the EIC for the assigned feature and visually inspect these.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SP")
```

Signal is very low for all of the features. And also no MS2 spectra were
found.

```{r mix01-uric-acid-ms2, fig.cap = "MS2 spectra for features matched to Uric acid", echo = FALSE}
if (length(std_ms2)) {
    par(mar = c(4.2, 4.5, 1.5, 0.5))
    xl <- range(unlist(mz(std_ms2)))
    plotSpectra(std_ms2, xlim = xl)
} else {
    cat("No MS2 spectra found for", std, "\n")
}
```


#### Summary

# Serum, negative polarity

We now perform the analysis on the samples with the standards still solved in
pure serum but acquired in negative polarity mode.

```{r}
POLARITY <- "NEG"
data <- filterFile(data_all, which(data_all$polarity == POLARITY))
hmdb <- hmdb_neg
hmdb_nl <- hmdb_neg_nl
```

```{r, preprocessing, eval = !file.exists(paste0(RDATA_PATH, "processed_data_", POLARITY, ".RData")), message = FALSE}
```

```{r, echo = FALSE}
load(paste0(RDATA_PATH, paste0("processed_data_", POLARITY, ".RData")))
data_FS <- filterFile(data, which(data$mode == "FS"),
                      keepFeatures = TRUE)
```

## Signal intensity difference

We compute the difference in (log2) signals between samples with high and low
concentration of the standards and calculate the p-value for this difference
using the Student's t-test.

```{r, abundance-difference}
```

## Identification of features matching standards

We now identify all features matching the any of the pre-defined set of adducts
for the standards of mix `r MIX`. Matching features are further subsetted to
those that show an at least twice as high signal in samples with high
concentration compared to those with low concentrations. Also, features with a
signal in low concentration but no detectable signal in high concentration are
removed.

```{r, define-adducts-neg}
```
```{r, match-features}
```

For most of the standards (`r length(unique(mD$target_name))` out of 
`r nrow(std_dilution01)`) in mix `r MIX` at least one feature was found
matching the standards adducts' m/z. 

```{r, table-standard-no-feature, results = "asis", echo = FALSE}
```

In the next sections we investigate for each standard which assignment would be
the correct one or, for those for which no signal was detected, why that was the
case.

## Standards with matching features

### 3-Phosphoglyceric Acid

```{r, echo = FALSE}
std <- "3-Phosphoglyceric Acid"
```

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

We next plot the EIC for the assigned feature and visually inspect these.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SN")
```

```{r mix01-serum-neg-3-phosphoglyceric-acid-ms2, fig.cap = "MS2 spectra for features matched to 3-Phosphoglyceric Acid.", echo = FALSE}
if (length(std_ms2)) {
    par(mar = c(4.2, 4.5, 1.5, 0.5))
    xl <- range(unlist(mz(std_ms2)))
    plotSpectra(std_ms2, xlim = xl)
} else {
    cat("No MS2 spectra found for", std, "\n")
}
```

#### Summary

- Beacuse of the new excluding criteria we loose FT0911 (because its
  `mean_high` is `NA` while `mean_low` it is not) which is the only feature
  with associated MS2 spectra.


### Acetylhistidine

```{r, echo = FALSE}
std <- "Acetylhistidine"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

Based on their retention time, it seems that these features represent signal
from several different compounds.

We next plot the EIC for the assigned feature and visually inspect these.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SN")
```

```{r mix01-serum-neg-acetylhistidine-ms2, fig.cap = "MS2 spectra for features matched to acetylhistidine.", echo = FALSE}
par(mar = c(4.2, 4.5, 1.5, 0.5))
xl <- range(unlist(mz(std_ms2)))
plotSpectra(std_ms2, xlim = xl)

```

Next we compare them against the reference spectra for `r std` from HMDB.
The results are shown in the heatmap below.

```{r plot-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for Acetylhistidine from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

The MS2 spectra F32.S0549, F31.S0586 (and maybe F32.S0599) from feature FT1073
have a quite high similarity against MS2 some spectra from `r std`.
Mirror plots for the best matching spectra are shown below.

```{r mix01-serum-neg-acetylhistidine-mirror-hmdb, echo = TRUE, fig.cap = "Mirror plots"}
std_ms2_sel <- plot_select_ms2(std_ms2, std_hmdb, 0.5,
                               ppm = csp@ppm, tolerance = csp@tolerance)
```

In addition, we perform also the match against reference spectra of `r std`
from MassBank.

```{r, plot-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for acetylhistidine from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

Surprisingly no spectrum matches MassBank. We also check if there are matches 
between neutral loss spectra for both HMDB and MassBank.

```{r, plot-ms2-hmdb-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for acetylhistidine from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r, plot-ms2-mbank-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for acetylhistidine from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

As for HMDB we don't find any. In the case of MassBank we find that the same
extracted spectra as in the first comparison using HMDB without neutral loss
computation (but the scores are very low).In addition we match (**all**) the MS2
spectra for the matched features against all spectra from HMDB or MassBank
identifying reference spectra with a similarity larger than 0.7. The results
(if any spectra matched) are shown in the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

For completeness, we also perform a neutral loss spectra comparison with HMDB
and MassBank.

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, hmdb_nl, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, mbank_nl,
              sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

#### Summary

- FT1073 (RT=177.1, `[M-H]-` ion): confidence level **B**. MS2 spectra match
  also carnosine.
- Reference MS2: F32.S0549 (also matches carnosine), F31.S0586.

### Betaine

```{r, echo = FALSE}
std <- "Betaine"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

Based on their retention time these features might represent signal
from the same compound.

We next plot the EIC for the assigned feature and visually inspect these.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SN")
```

We next extract the MS2 spectra for all these features and match them against
the reference spectra for `r std` from HMDB.

```{r mix01-serum-neg-betaine-ms2, fig.cap = "MS2 spectra for features matched to betaine.", echo = FALSE}
if (length(std_ms2)) {
    par(mar = c(4.2, 4.5, 1.5, 0.5))
    xl <- range(unlist(mz(std_ms2)))
    plotSpectra(std_ms2, xlim = xl)
} else {
    cat("No MS2 spectra found for", std, "\n")
}
```

#### Summary

- FT0616 (RT=156.2, `[M+HCOO]-` ion): confidence level **D**.
- FT0496 (RT=159.6, `[M+Cl]-` ion): confidence level as above.
- No MS2 spectra associated to the features.


### CDP

```{r, echo = FALSE}
std <- "CDP"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

Based on their retention time, it seems that these features represent signal
from several different compounds.

We next plot the EIC for the assigned feature and visually inspect these. EICs
look rather noisy and large retention time shifts between the samples are
visible.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SN")
```

```{r mix01-serum-neg-CDP-ms2, fig.cap = "MS2 spectra for features matched to CDP", echo = FALSE}
if (length(std_ms2)) {
    par(mar = c(4.2, 4.5, 1.5, 0.5))
    xl <- range(unlist(mz(std_ms2)))
    plotSpectra(std_ms2, xlim = xl)
} else {
    cat("No MS2 spectra found for", std, "\n")
}
```

### Dimethylglycine

```{r, echo = FALSE}
std <- "Dimethylglycine"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

We next plot the EIC for the only feature matched to `r std` and show it below.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SN")
```

```{r mix01-serum-neg-dimethylglycine-ms2, fig.cap = "MS2 spectra for features matched to dimethylglycine", echo = FALSE}
if (length(std_ms2)) {
    par(mar = c(4.2, 4.5, 1.5, 0.5))
    xl <- range(unlist(mz(std_ms2)))
    plotSpectra(std_ms2, xlim = xl)
} else {
    cat("No MS2 spectra found for", std, "\n")
}
```

Unfortunately no MS2 spectra associated to FT0087 is available.

#### Summary

- FT0087 (RT=177, `[M-H]-` ion): confidence level **D**.


### Glycerol

```{r, echo = FALSE}
std <- "Glycerol"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

Also for `r std` we have only one match. We next plot the and show the EIC
of the realted feature.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SN")
```

```{r mix01-serum-neg-glycerol-ms2, fig.cap = "MS2 spectra for features matched to glycerol", echo = FALSE}
if (length(std_ms2)) {
    par(mar = c(4.2, 4.5, 1.5, 0.5))
    xl <- range(unlist(mz(std_ms2)))
    plotSpectra(std_ms2, xlim = xl)
} else {
    cat("No MS2 spectra found for", std, "\n")
}
```

Unfortunately no MS2 spectra is available.

#### Summary

- very low signal. Seems nothing properly detected.

### L-Glutamic Acid

```{r, echo = FALSE}
std <- "L-Glutamic Acid"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

Based on their retention time, it seems that these features represent signal
from 2 different compounds.

We next plot the EIC for the assigned feature and visually inspect these.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SN")
```

We next extract the MS2 spectra for all these features and match them against
the reference spectra for `r std` from HMDB.

```{r mix01-serum-neg-L-Glutamic-Acid-ms2, fig.cap = "MS2 spectra for features matched to L-Glutamic Acid.", echo = FALSE}
if (length(std_ms2)) {
    par(mar = c(4.2, 4.5, 1.5, 0.5))
    xl <- range(unlist(mz(std_ms2)))
    plotSpectra(std_ms2, xlim = xl)
} else {
    cat("No MS2 spectra found for", std, "\n")
}
```

Next we compare them against the reference spectra for `r std` from HMDB. The
results are shown in the heatmap below.

```{r, plot-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for L-Glutamic Acid from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

The MS2 spectra F31.S0592, F32.S0560, F32.S0611 (from FT0433) have quite
high similarity against some MS2 spectra from `r std`. Mirror plots for the
best matching spectra (similarity higher than 0.7) are shown below.

```{r mix01-serum-neg-L-Glutamic-Acid-mirror-hmdb, echo = TRUE, fig.cap = "Mirror plots"}
std_ms2_sel <- plot_select_ms2(std_ms2, std_hmdb, 0.7,
                               ppm = csp@ppm, tolerance = csp@tolerance)
```

We perform also the match against reference spectra for that standard from
MassBank obtaining similar results as shown in the heatmap below.

```{r, plot-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for L-Glutamic Acid from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

In addition we match (**all**) the MS2 spectra for the matched features against
all spectra from HMDB or MassBank identifying reference spectra with a
similarity larger than 0.7. The results (if any spectra matched) are shown in
the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

Extracted spectra match also reference spectra of compounds different from
`r std`. 

#### Summary

- FT0433(RT=176.4, `[M-H]-` ion): confidence level **B**.
- Reference MS2: F31.S0592, F32.S0560, F32.S0611.


### Myo-Inositol

```{r, echo = FALSE}
std <- "Myo-Inositol"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

Based on their retention time, it seems that these features represent signal
from several different compounds.

We next plot the EIC for the assigned feature and visually inspect these.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SN")
```

The only available MS2 spectrum available for the above features is shown below.

```{r mix01-serum-neg-myo-inositol-ms2, fig.cap = "MS2 spectra for features matched to myo-inositol.", echo = FALSE}
if (length(std_ms2)) {
    par(mar = c(4.2, 4.5, 1.5, 0.5))
    xl <- range(unlist(mz(std_ms2)))
    plotSpectra(std_ms2, xlim = xl)
} else {
    cat("No MS2 spectra found for", std, "\n")
}
```

Next we compare it against the reference spectra for `r std` from HMDB
and MassBank.

```{r, plot-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for myo-inositol from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r, plot-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for myo-inositol from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

In both cases the extracted spectrum does not match any reference spectra from
`r std`. Since feature FT1334, the feature associated to the available spectrum,
was matched to an ion (`[M+Cl]-`) different from `[M-Cl]-` we also look at
neutral loss spectra.

```{r, plot-ms2-mbank-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for myo-inositol from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r, plot-ms2-hmdb-nl-heatmap, fig.cap = "Similarities of MS2 neutral loss spectra for selected features against reference neutral loss spectra for myo-inositol from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

As for the MassBank comparison we don't find any match while for the HMDB one
we find a few matches with low score. Below we report the only match with score
above 0.2.

```{r mix01-serum-myo-inositol-mirror-massbank-nl, echo = TRUE, fig.cap = "Mirror plots"}
tmp_res <- plot_select_ms2(std_ms2_nl, std_hmdb_nl, 0.2,
                           ppm = csp_nl@ppm, tolerance = csp_nl@tolerance)
```

In addition we match (**all**) the MS2 spectra for the matched features against
all spectra from HMDB or MassBank identifying reference spectra with a
similarity larger than 0.7. The results (if any spectra matched) are shown in
the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

For completeness, we also perform a neutral loss spectra comparison with HMDB
and MassBank.

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, hmdb_nl, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, mbank_nl,
              sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```
In any case, no match with similarity higher than 0.7 is found.

#### Summary

- FT1334 (RT = 193.7, `[M+Cl]-` ion) with confidence **C** (because of the
  neutral loss spectra match)?
- FT1460 (RT = 192.2, `[M+HCOO]-` ion) and FT0834 (RT = 192.2, `[M-H]-` ion)
  inheriting confidence level from FT1334.
- other features with confidence **D**?

### propionic acid

```{r, echo = FALSE}
std <- "propionic acid"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

Based on their retention time, it seems that these features represent signal
from several different compounds.

We next plot the EIC for the assigned feature and visually inspect these.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SN")
```

```{r mix01-serum-neg-propionic-acid-ms2, fig.cap = "MS2 spectra for features matched to myo-inositol.", echo = FALSE}
if (length(std_ms2)) {
    par(mar = c(4.2, 4.5, 1.5, 0.5))
    xl <- range(unlist(mz(std_ms2)))
    plotSpectra(std_ms2, xlim = xl)
} else {
    cat("No MS2 spectra found for", std, "\n")
}
```
```{r, mix01-serum-neg-propionic-acid-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for myo-inositol from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r, mix01-serum-neg-propionic-acid-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for myo-inositol from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

The extracted MS2 spectrum doesn't match any reference spectra of `r std` from
HMDB or MassBank. In addition we match (**all**) the MS2 spectra for the
matched features against all spectra from HMDB or MassBank identifying
reference spectra with a similarity larger than 0.7. The results (if any
spectra matched) are shown in the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```


#### Summary

Only a single MS2 spectra that doesn't match with any reference spectrum and
looking at the EICs the intensity is relatively low for all features .

### Pyruvic Acid


```{r, echo = FALSE}
std <- "Pyruvic Acid"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```


```{r, echo = FALSE}
plot_eics(data, std, tmp, "SN")
```

We next look if there are MS2 spectra associated to the above features.

```{r mix01-serum-neg-pyruvic-acid-ms2, fig.cap = "MS2 spectra for features matched to pyruvic-acid", echo = FALSE}
std_ms2 <- extract_ms2(data, rownames(tmp))
if (length(std_ms2)) {
    par(mar = c(4.2, 4.5, 1.5, 0.5))
    xl <- range(unlist(mz(std_ms2)))
    plotSpectra(std_ms2, xlim = xl)
} else {
    cat("No MS2 spectra found for", std, "\n")
}
```

Next we compare them against the reference spectra for `r std` from HMDB and
MassBank.

```{r, plot-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for Pyruvic Acid from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r, plot-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for Pyruvic Acid from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

No match is found in both cases. Although FT0044 and FT0043 are both from ion
`[M-H]-`, we also perform a neutral loss comparison. While for HMDB no match is 
found, surprisingly we get matches in the MassBank case. Below we plot those
with a similarity score greater than 0.7.

```{r, plot-ms2-hmdb-nl-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for Pyruvic Acid from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r, plot-ms2-mbank-nl-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for Pyruvic Acid from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

```{r mix01-serum-pyruvic-acid-mirror-massbank-nl, echo = TRUE, fig.cap = "Mirror plots"}
tmp_res <- plot_select_ms2(std_ms2_nl, std_mbank_nl, 0.7,
                           ppm = csp_nl@ppm, tolerance = csp_nl@tolerance)
```

In addition we match (**all**) the MS2 spectra for the matched features against
all spectra from HMDB or MassBank identifying reference spectra with a
similarity larger than 0.7. The results (if any spectra matched) are shown in
the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
std_ms2_sel <- Spectra()
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

For completeness, we also perform a neutral loss spectra comparison with HMDB
and MassBank.

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, hmdb_nl, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

```{r, echo = FALSE, message = FALSE, results = "asis", eval = ALL_NL_MATCH}
perform_match(std_ms2_nl, mbank_nl,
              sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp_nl)
```

No match involving `r std` was found but similarity with MS2 spectra for
Ascorbic acid are high.


#### Summary
We have two features associated to the same ion. In particular, based on
neutral loss spectra comparison against MassBank:
- FT0043 (RT = 41.7, `[M-H]-` ion, associated spectrum F32.S0174) with
  confidence **B**.
- FT0328 (RT = 40.3, `[M+HCOO]-` ion) inheriting confidence from FT0043?
  Although its EIC doesn't look that good.
- FT0044 (RT = 65.1, `[M-H]-` ion, associated spectra F31.S0237, F32.S0267)
  with confidence **B**.
- FT0329 (RT = 62.9, `[M+HCOO]-` ion) inheriting confidence from FT0044?
  Although its EIC doesn't look that good.

Only MS2 spectra from FT0043 are available but they don't match reference
spectra of `r std` (but those of ascorbic acid).

### Suberic Acid

```{r, echo = FALSE}
std <- "Suberic Acid"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

Based on their retention time, it seems that these features represent signal
from 3(or 4?) different compounds.

We next plot the EIC for the assigned feature and visually inspect these.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SN")
```

The cleaned MS2 spectra for the features matched to `r std` are shown below
(peaks with an intensity below 5% of the maximum peak and spectra with less
than 2 peaks where removed).

```{r mix01-serum-neg-suberic-acid-ms2, fig.cap = "MS2 spectra for features matched to suberic acid.", echo = FALSE}
par(mar = c(4.2, 4.5, 1.5, 0.5))
xl <- range(unlist(mz(std_ms2)))
plotSpectra(std_ms2, xlim = xl)
```

Next we compare them against the reference spectra for `r std` from HMDB. The results are shown in
the heatmap below.

```{r, plot-ms2-hmdb-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for suberic acid from HMDB.", echo = FALSE, fig.width = 5, fig.height = 5}
```

Al the extracted spectra seem to match and F31.S0148 from FT0757 and FT0758
have a particularly high similarity score for some matches. Mirror plots for
the best matching spectra are shown below.

```{r mix01-serum-neg-suberic-acid-mirror-hmdb, echo = TRUE, fig.cap = "Mirror plots"}
std_ms2_sel <- plot_select_ms2(std_ms2, std_hmdb, 0.7,
                               ppm = csp@ppm, tolerance = csp@tolerance)
```

In addition we perform also the match against reference spectra for that
standard from MassBank.

```{r, plot-ms2-mbank-heatmap, fig.cap = "Similarities of MS2 spectra for selected features against reference spectra for Suberic Acid from MassBank.", echo = FALSE, fig.width = 5, fig.height = 5}
```

In addition we match (**all**) the MS2 spectra for the matched features against
all spectra from HMDB or MassBank identifying reference spectra with a
similarity larger than 0.7. The results (if any spectra matched) are shown in
the two following tables.

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, hmdb, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```

```{r, echo = FALSE, message = FALSE, results = "asis"}
perform_match(std_ms2, mbank, sv = c("rtime", "target_name", "score"),
              name = "target_name", param = csp)
```


#### Summary

- FT0757 (RT=37.3, `[M-H]-` ion): confidence level **A**.
- FT1399 (RT=37.3, `[M+HCOO]-` ion) inherits confidence level **A** from FT0757?
  But its EIC has very low intensity in comparison.
- FT0758 (RT=64.1, `[M-H]-` ion): confidence level **A**?
- Reference MS2: F31.S0148, F31.S0260.


### Uric acid

```{r, echo = FALSE}
std <- "Uric acid"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

We show the EICs for the features above.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SN")
```

We next look if there are MS2 spectra associated to the above features.

```{r mix01-serum-neg-uric-acid-ms2, fig.cap = "MS2 spectra for features matched to pyruvic-acid", echo = FALSE}
if (length(std_ms2)) {
    par(mar = c(4.2, 4.5, 1.5, 0.5))
    xl <- range(unlist(mz(std_ms2)))
    plotSpectra(std_ms2, xlim = xl)
} else {
    cat("No MS2 spectra found for", std, "\n")
}
```


#### Summary

- No MS2 spectra is available, the EIC of the only matched feature has low
  signal and maybe don't look good.

### Xanthine

```{r, echo = FALSE}
std <- "Xanthine"
```

The table below lists all features that were matched to one of the adducts of 
`r print(std)` that have in addition also on average a twice as high signal in
samples with higher concentrations than in those with lower concentrations.

```{r, table-feature-matches, results = "asis", echo = FALSE}
```

Based on their retention time, it seems that these features represent signal
from 3 different compounds.

We next plot the EIC for the assigned feature and visually inspect these.

```{r, echo = FALSE}
plot_eics(data, std, tmp, "SN")
```

We next extract the MS2 spectra for all these features and match them against
the reference spectra for `r std` from HMDB.

```{r mix01-serum-neg-xanthine-ms2, fig.cap = "MS2 spectra for features matched to xanthine", echo = FALSE}
if (length(std_ms2)) {
    par(mar = c(4.2, 4.5, 1.5, 0.5))
    xl <- range(unlist(mz(std_ms2)))
    plotSpectra(std_ms2, xlim = xl)
} else {
    cat("No MS2 spectra found for", std, "\n")
}
```

#### Summary

- No MS2 spectra available.
- FT1090 (RT=140.2, `[M-H]-` ion): confidence level **D**. The EIC looks good.
- FT0491 was grouped together with FT1090 but its EIC has low intensity and
  maybe doesn't look that good.

## Standards without any matching feature

### C3 Carnitine
### Creatine

# Session information

The R version and packages used in this analysis are listed below.

```{r sessioninfo}
sessionInfo()
```
