---
title: "Preprocessing of the *standards* data set"
author: "Andrea Vicini, Johannes Rainer"
output:
  rmarkdown::html_document:
    highlight: pygments
    toc: true
    toc_depth: 3
    fig_width: 5
---

```{r style, echo = FALSE, results = 'asis', message = FALSE}
library(BiocStyle)
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

**Last modified:** `r file.info("preprocessing-standards.Rmd")$mtime`<br />
**Compiled**: `r date()`

```{r}
mix <- 1
mix_name <- paste0("Mix", ifelse(mix<10,paste0(0,mix),mix)) 
```

```{r general-settings, echo = FALSE}
IMAGE_PATH <- paste0("images/preprocessing-standards/", tolower(mix_name),"/")
RDATA_PATH <- paste0("data/RData/preprocessing-standards/", tolower(mix_name), "/")
dir.create(IMAGE_PATH, showWarnings = FALSE)
dir.create(RDATA_PATH, showWarnings = FALSE)

## Define the *base* path where mzML files can be found. This is
## /data/massspec/mzML/ on the cluster
MZML_PATH <- "~/mix01/" #"/data/massspec/mzML/"

library(BiocParallel)
#' Set up parallel processing using 2 cores
if (.Platform$OS.type == "unix") {
    register(bpstart(MulticoreParam(2)))
} else {
    register(bpstart(SnowParam(2)))
}
```

# Introduction

In this document we perform the preprocessing and analysis of mzML files to
determine retention times and measured ions for the *standards*. These standards
are a collection of ~ 250 pure standards of polar metabolites which were used to
setup the HILIC-based LC-MS protocols to measure the polar metabolome in human
serum samples.


# Data preprocessing

In this section we perform the preprocessing of the LC-MS(/MS) data for all
mzML files of one mix of standards.

Below we load all required libraries and the definition of the dataset.

```{r libraries}
library(xcms)
library(pander)
library(RColorBrewer)
library(magrittr)
library(Rdisop)
library(MetaboCoreUtils)
library(MsCoreUtils)
library(MsFeatures)
library(SummarizedExperiment)
library(CompMetaboTools)
library(DBI)
library(dplyr)

std_serum_files <- read.table("data/std_serum_files.txt", header = TRUE)
```

We next subset the data set to files with samples from a single standards mix.

```{r}
std_serum_files01 <- std_serum_files[which(std_serum_files$type == mix_name), ]
```

The list of standards that constitute the present sample mix are listed below
along with the expected retention time and the most abundant adduct for positive
and negative polarity as defined in a previous analysis.

```{r, echo = FALSE, results = "asis"}
std_dilution <- read.table("data/standards_dilution.txt",
                           sep= "\t", header = TRUE)
std_dilution01 <- std_dilution[std_dilution$mix == mix, ]
table1 <- std_dilution01[, c("name", "formula", "RT", "POS", "NEG")]
pandoc.table(table1, style = "rmarkdown", caption = paste0("Standards of ",
                                                           mix_name))
```

We next import the MS data from all mzML files of this mix.

```{r load-data, warning = FALSE, eval = !file.exists(paste0(RDATA_PATH, "data.RData"))}
fls <- paste0(MZML_PATH, std_serum_files01$folder, "/", std_serum_files01$mzML)
data <- readMSData(fls, pdata = new("NAnnotatedDataFrame", std_serum_files01),
                   mode = "onDisk")
save(data, file = paste0(RDATA_PATH, "data.RData"))
```


```{r load-data-cached, echo = FALSE, eval = file.exists(paste0(RDATA_PATH, "data.RData"))}
load(paste0(RDATA_PATH, "data.RData"))
```

We next create the base peak chromatogram (BPC) for each file and define colors
for the different sets of samples.

```{r}
bpc <- chromatogram(data, aggregationFun = "max")

#' Define a new variable combining the polarity and the matrix in which the
#' standards are solved.
tmp <- rep("Water", length(fileNames(data)))
tmp[grep("^QC", data$class)] <- "Serum"
matrix_pol <- paste0(tmp, "_", data$polarity)
data$matrix_pol <- matrix_pol

#' Define the colors for each type
col_matrix_pol <- brewer.pal(5, "Set1")[c(1, 2, 4, 5)]
names(col_matrix_pol) <- c("Water_POS", "Water_NEG", "Serum_NEG", "Serum_POS")
```

The BPC are shown below.

```{r bpc-mix01, fig.path = IMAGE_PATH, fig.cap = "Base peak chromatogram. Different colours for polarity and matrix. The vertical grey dotted lines indicate the retention time where an ion of a standard present in the current mix is expected.", fig.width = 10, fig.height = 5}
plot(bpc, col = paste0(col_matrix_pol[data$matrix_pol], 80))
legend("topright", col = col_matrix_pol,
       legend = names(col_matrix_pol), lwd = 1)
abline(v = std_dilution01$RT, lty = 3, col = "grey")
```

As expected, the signal in serum samples (orange and purple) seems to be higher
for the retention time regions in which a high number of compounds is expected
(20-60, 110-200 seconds). Surprisingly, also the samples with the mix solved in
water yield high signals for positive polarity (in the region from 60-120
seconds, but also around 160 seconds). In general, samples measured in positive
polarity seem to have higher intensities than the corresponding samples measured
in negative polarity.

Next we perform the chromatographic peak detection on all samples and
subsequently *refine* also the chromatographic peaks.

```{r peakdetection, eval = !file.exists(paste0(RDATA_PATH, "data_peakdetection.RData"))}
cwp <- CentWaveParam(ppm = 50,
                     peakwidth = c(2, 20),
                     snthresh = 5,
                     mzdiff = 0.001,
                     prefilter = c(4, 300),
                     noise = 100,
                     integrate = 2)

data <- findChromPeaks(data, param = cwp) 
save(data, file = paste0(RDATA_PATH, "data_peakdetection.RData"))
```

```{r peakdetection-cached, echo = FALSE, eval = file.exists(paste0(RDATA_PATH, "data_peakdetection.RData"))}
load(paste0(RDATA_PATH, "data_peakdetection.RData"))
```

```{r refinement, message = FALSE, warning = FALSE, eval = !file.exists(paste0(RDATA_PATH, "data_refined.RData"))}
mnp <- MergeNeighboringPeaksParam(expandRt = 2.5, expandMz = 0.001,
                                  minProp = 3/4)
data <- refineChromPeaks(data, param = mnp)
save(data, file = paste0(RDATA_PATH, "data_refined.RData"))
```

```{r refinement-cached, echo = FALSE, eval = file.exists(paste0(RDATA_PATH, "data_refined.RData"))}
load(paste0(RDATA_PATH, "data_refined.RData"))
```

We next evaluate the signal for Xanthine. For this we first calculate first its
monoisotopic mass based on the compound's chemical formula and then the m/z of
its [M+H]+ ion. 

```{r xanthine-mz}
frml_xan <- std_dilution01[std_dilution01$name == "Xanthine", "formula"]
mass_xan <- getMolecule(frml_xan)$exactmass    # calculate mass
mz_xan <- mass2mz(mass_xan, "[M+H]+")[1, 1]    # get the m/z for [M+H]+
mzr_xan <- mz_xan + c(1, -1) * ppm(mz_xan, 50) # calculate m/z range
```


```{r Xanthine-detected_peaks, fig.path = IMAGE_PATH, fig.cap = "Detected peaks for Xanthine.", fig.width = 7, fig.height = 5}
chr <- chromatogram(filterFile(data, file = which(data$polarity == "POS")), 
                    rt = c(130, 150), mz = mzr_xan)
col_samples <- col_matrix_pol[chr$matrix_pol]
plot(chr, col = col_samples, 
     peakBg = paste0(col_samples[chromPeaks(chr)[, "sample"]], 40))
legend("topright", col = col_matrix_pol,
       legend = names(col_matrix_pol), lwd = 1, cex = 1)
# The last two peaks have not been detected
```

- [ ] table with the number of peaks per file.

```{r, echo = FALSE, results = "asis"}
pks <- chromPeaks(data)
nppf <- cbind(data$mzML, as.numeric(table(pks[, "sample"])))
colnames(nppf) <- c("file name", "number of peaks")
pandoc.table(nppf, style = "rmarkdown", caption = "Number of peaks per file")
```

We split the data set for positive and negative polarity

```{r}
data_pos <- filterFile(data, file = which(data$polarity == "POS"))
data_neg <- filterFile(data, file = which(data$polarity == "NEG"))
```

- [ ] run correspondence analysis (separately for the positive and negative
      polarity).

```{r correspondence}
pdp <- PeakDensityParam(sampleGroups = data_pos$matrix_pol, bw = 1.8,
                        minFraction = 0.6, binSize = 0.02)
data_pos <- groupChromPeaks(data_pos, param = pdp)
data_neg <- groupChromPeaks(data_neg, param = pdp) 
save(data_pos,data_neg, file = paste0(RDATA_PATH, "data_postcorrespondence.RData"))
```

```{r correspondence-cached, echo = FALSE, eval = file.exists(paste0(RDATA_PATH, "data_postcorrespondence.RData"))}
load(paste0(RDATA_PATH, "data_postcorrespondence.RData"))
```

```{r}
fD_pos <- featureDefinitions(data_pos)
fD_neg <- featureDefinitions(data_neg)
#' feature intensity matrix
fmat_pos <- featureValues(data_pos, value = "into", method = "sum")
fmat_neg <- featureValues(data_neg, value = "into", method = "sum")
```


```{r}
#' Per-feature summary.
head(featureSummary(data_pos))
head(featureSummary(data_neg))
#' Percentage of missing values before filling
sum(is.na(fmat_pos)) / length(fmat_pos) # positive polarity
sum(is.na(fmat_neg)) / length(fmat_neg) # negative polarity
```

```{r fillChromPeaks, echo = FALSE, eval = file.exists(paste0(RDATA_PATH, "data_postfilling.RData"))}
data_pos <- fillChromPeaks(data_pos, param = ChromPeakAreaParam())
data_neg <- fillChromPeaks(data_neg, param = ChromPeakAreaParam())
save(data_pos,data_neg, file = paste0(RDATA_PATH, "data_postfilling.RData"))
```

```{r fillChromPeaks-cached, echo = FALSE, eval = file.exists(paste0(RDATA_PATH, "data_postfilling.RData"))}
load(paste0(RDATA_PATH, "data_postfilling.RData"))
```

```{r}
fmat_pos_f <- featureValues(data_pos, value = "into", method = "sum")
fmat_neg_f <- featureValues(data_neg, value = "into", method = "sum")

#' Percentage of missing values after filling
sum(is.na(fmat_pos_f)) / length(fmat_pos_f) # positive polarity
sum(is.na(fmat_neg_f)) / length(fmat_neg_f) # negative polarity

```

We start to consider features for water and also subset to FS mode only

Choose substetting
```{r}
# Maybe these can become parameters of a function
pol <- "POS" # "POS" or "NEG"
matr <- "Water" # "Water" or "Serum"
fmat <- get(paste0("fmat_",tolower(pol)))
fmat_f <- get(paste0("fmat_",tolower(pol),"_f"))
fD <- get(paste0("fD_",tolower(pol)))
data_subs <- get(paste0("data_",tolower(pol))) #
mp=paste0(matr,"_",pol)

```


```{r}
keep_features <- rowSums(!is.na(fmat[, which(data_subs$matrix_pol == mp 
                                 & data_subs$mode == "FS")])) >= 2
fmat_mp <- fmat_f[ keep_features, which(data_subs$matrix_pol == mp 
                                 & data_subs$mode == "FS")]

# fmat_mp <- fmat_f[, which(data_subs$matrix_pol == mp 
#                                  & data_subs$mode == "FS")]
  
# Exclude rows with NA
fmat_mp <- na.omit(fmat_mp)
```

and log2 transform the abundances (and keep only rows without missing values)

```{r}
fmat_mpl2 <- log2(fmat_mp)
```

We perform a t-test to see if there is a significant positive difference between
samples with high concentrations of standards and samples with low concentration

```{r}
# H0: high==low vs H1: H0^c
high <- grep("High", colnames(fmat_mpl2)) #c(1,3,6)
low <- grep("Low", colnames(fmat_mpl2)) #c(2,4,5)

pvalues <- apply(fmat_mpl2, 1, function(x) {
    res <- t.test(x[high], x[low], mu = 0)
    c(pvalue = res$p.value,
      M = unname(res$estimate[1] - res$estimate[2]))
})
pvalues <- t(pvalues)
```


```{r Volcano-plot-pvalues, fig.path = IMAGE_PATH, fig.cap = "Volcano plot of p-values.", fig.width = 7, fig.height = 5}
# Volcano plot
plot(pvalues[, "M"], -log10(pvalues[, "pvalue"]), pch = 21, col = "#00000080",
     bg = "#00000020", xlab = "M", ylab = expression(-log[10]~(p)))
```

What's not clear to me:
- Why do we have so many features *significant*?
- Why do we have features with higher abundance in low compared to high?

TODO:
- [ ] (pandoc) table with the *significant* features (those with pvalue < 0.05
      and a M-value > 3 (8-times higher concentration in high)): columns:
      feature id, mzmed, rtmed (from `featureDefinitions`), M, pvalue. Order
      them by rtmed.
      

```{r, echo = FALSE, results = "asis"}
#sign_f_indexes <- which(pvalues[, "pvalue"] < 0.05 & pvalues[, "M"] > 3)
sign_f_id <- rownames(fmat_mp)[which(pvalues[, "pvalue"] < 0.05 & 
                                       pvalues[, "M"] > 3)]
sign_f <- cbind(sign_f_id, fD[sign_f_id,c("mzmed", "rtmed")], 
               pvalues[sign_f_id,])
sign_f <- sign_f[order(sign_f[, "rtmed" ], decreasing = FALSE),]
pandoc.table(data.frame(sign_f), style = "rmarkdown", 
             caption = "Significant features")
```
- [ ] plot all significant features (export as png file). Use the
      `featureChromatograms` function (with parameters `expandRt = 2` and
      `filled = TRUE`) to extract the EIC and plot that. Add also a vertical
      line to the plot where we expect one of the standards from mix01 to be
      measured. Look through all plots and see if maybe some have ~ the same
      retention time and similar peak shape (could be different adducts/ions of
      the same compound).

```{r sign-featureChromatograms}
f_chrs <- featureChromatograms(data_subs, expandRt = 2, 
                               features = sign_f_id, filled = TRUE)
save(f_chrs, file = paste0(RDATA_PATH, "f_chrs_", mp, ".RData"))
```



```{r plotfeatures}
# Plot all the features and save their plot as a png files in the 
# selectedfeatures folder 
if(!dir.exists(paste0(IMAGE_PATH, "selectedfeatures", mp)))
  dir.create(paste0(IMAGE_PATH, "selectedfeatures", mp))
col_hl <- rep(brewer.pal(5, "Set1")[3], ncol(f_chrs)) # green
col_hl[low] <- brewer.pal(5, "Set1")[4] # purple
for (i in 1:nrow(f_chrs))
{
png(filename = paste0(IMAGE_PATH, "selectedfeatures", mp,"/feature", i, ".png"), 
    width=700, height=500)
chr_obj=f_chrs[i, match(colnames(fmat_mp), f_chrs$mzML)]
plot(chr_obj, col = col_hl, 
     peakBg = paste0(col_hl[chromPeaks(chr_obj)[, "sample"]], 40))
legend("topright", col = brewer.pal(5, "Set1")[c(3, 4)],
       legend = c("high", "low"), lwd = 1, cex = 1)
abline(v = std_dilution01$RT, lty = 3)
dev.off()
}

```

- [ ] let's see how to proceed from here... maybe feature grouping? Install
      `RforMassSpectrometry/MsFeatures` for that and have a look at its
      documentation and vignette. We want to group features if the have about
      the same retention time and then have also correlated abundances.

# Grouping of features      
```{r}
plot(sign_f$rtmed, sign_f$mzmed,
     xlab = "retention time", ylab = "m/z", main = "features",
     col = "#00000060")
abline(v= table1$RT, lty = 2)
grid()
```

```{r}
featureGroups(data_subs) <- NA_character_
featureDefinitions(data_subs)[sign_f$sign_f_id, "feature_group"] <- "FG"
table(featureGroups(data_subs))
```

 Grouping of features by similar retention time
```{r}
data_subs <- groupFeatures(data_subs, param = SimilarRtimeParam(4))
table(featureGroups(data_subs))
```

```{r}
library(pheatmap)
fvals <- log2(featureValues(data_subs, filled = TRUE, 
                            method = "sum")[sign_f$sign_f_id, ])

cormat <- cor(t(fvals), use = "pairwise.complete.obs")
ann <- data.frame(fgroup = featureDefinitions(data_subs)[sign_f$sign_f_id, 
                                                         "feature_group"])
rownames(ann) <- rownames(cormat)

res <- pheatmap(cormat, annotation_row = ann, cluster_rows = TRUE,
                cluster_cols = TRUE)
```

Grouping of features by abundance correlation across samples
```{r}
data_subs <- groupFeatures(data_subs, 
                           AbundanceSimilarityParam(threshold = 0.7, 
                                                    transform = log2), 
                           filled = TRUE, method = "sum")
table(featureGroups(data_subs))
```

```{r}
plotFeatureGroups(data_subs)
grid()
```

Grouping of features by EIC correlation
```{r EIC-grouping, echo = FALSE, eval = !file.exists(paste0(RDATA_PATH, "data_post_grouping", mp, ".RData"))}
data_subs <- groupFeatures(data_subs, EicCorrelationParam(threshold = 0.7, n = 2,
                                                  clean = TRUE))
save(data_subs, file = paste0(RDATA_PATH, "data_post_grouping", mp, ".RData"))
table(featureGroups(data_subs))
```

```{r grouping-cached, echo = FALSE, eval = file.exists(paste0(RDATA_PATH, "data_post_grouping", mp, ".RData"))}
load(paste0(RDATA_PATH, "data_post_grouping", mp, ".RData"))
```


```{r, echo = FALSE, results = "asis"}
# Table with significant features and groups of features
fD=featureDefinitions(data_subs)
sign_f <- cbind(sign_f_id, fD[sign_f_id,c("mzmed", "rtmed")], 
               pvalues[sign_f_id,], 
               group = fD[sign_f_id, c("feature_group")])
sign_f <- sign_f[order(sign_f[, "rtmed" ]),]
sign_f <- sign_f[order(sign_f[, "group" ]),]
rtmed_group=split(sign_f$rtmed, sign_f$group) %>%
vapply(FUN = mean, numeric(1))
rtmed_group=rep(rtmed_group, unname(table(sign_f$group)))
sign_f <- cbind(sign_f,rtmed_group)
sign_f <- sign_f[order(sign_f[, "rtmed_group" ]),]
pandoc.table(data.frame(sign_f), style = "rmarkdown", caption = "Significant features")
```

```{r}
plotFeatureGroups(data_subs)
grid()
```

```{r}
# Some groups splitted: FG.010.001, FG.002.001
fts <- grep("FG.002.001", featureGroups(data_subs))
eics <- featureChromatograms(data_subs, features = fts,
                             filled = TRUE, n = 1)
```

```{r}
cols <- rainbow(length(fts))
names(cols) <- unique(featureGroups(data_subs)[fts])

par(mfrow = c(1, 2))
plotOverlay(eics, col = cols[featureGroups(data_subs)[fts]], lwd = 2)
plotOverlay(normalize(eics), col = cols[featureGroups(data_subs)[fts]], lwd = 2)
```

Compute exact masses of standars
```{r}
frmls <- std_dilution01[, "formula"]
masses <- sapply(frmls, function(formula) getMolecule(formula)$exactmass)
```

Compute m/z of possible adducts of the standars
```{r}
#mz_adducts <- mass2mz(masses, c("[M+H]+","[M+Na]+"))
mz_adducts <- mass2mz(masses, adductNames(ifelse(pol == "POS", "positive", "negative")))
ord <- order(mz_adducts[, 1])
frmls <- frmls[ord]
info_stds <- std_dilution01 [ord, ]
mz_adducts <- mz_adducts[ord, ]
```

mzmed of significant features
```{r}
mzmed_f <- sign_f$mzmed
names(mzmed_f) <- sign_f$sign_f_id
mzmed_f <- sort(mzmed_f)
```

Compute matrix of matchings
```{r}
#closest(sort(sign_f$mzmed),sort(mz_adducts[,1]), tolerance = 0, ppm = 20)
matches_mat <-apply(mz_adducts, 2, function(col) closest(mzmed_f, col,
                                                         tolerance = 0, ppm = 20))
rownames(matches_mat) <- names(mzmed_f)
```


```{r}
# from adduct to feature
rownames(matches_mat)[which(!is.na(matches_mat[, "[M+H]+"]))] 
matches_mat[which(!is.na(matches_mat[,"[M+H]+"])), "[M+H]+"] 
# from feature to adduct
colnames(matches_mat)[which(!is.na(matches_mat["FT0463", ]))] 
frmls[matches_mat["FT0463", which(!is.na(matches_mat["FT0463", ]))]] 
sign_f[which(sign_f[,"sign_f_id"] == "FT0463"), ]
```


```{r}
get_adducts <- function(ft, matches_mat, info_stds = NULL)
{
  add_name <- colnames(matches_mat)[which(!is.na(matches_mat[ft, ]))]
  res <- cbind(add_name, 
               info_stds[matches_mat[ft, which(!is.na(matches_mat[ft, ]))], ])
  res
}
#get_adducts(ft="FT0463", matches_mat, info_stds[,c("name", "formula")])
```

```{r}
matchings=NULL
for (i in 1:nrow(sign_f))
{
  addcts <- get_adducts(ft = sign_f$sign_f_id[i], matches_mat, 
                     info_stds[, c("name", "formula")])
  if(nrow(addcts)>0)
    matchings <- rbind(matchings,cbind(id = sign_f$sign_f_id[i],
                                    group = sign_f$group[i],addcts)) 
}
```

Matchings per group feature
```{r}
matchings_groups <- lapply(unique(matchings$group), function(x) 
  matchings[which(matchings$group == x), c("id", "add_name", "name", "formula")])
names(matchings_groups)=unique(matchings$group)
```

Plot of the matchings attempt. Features belonging to the same group have been 
connected with a straight line 

```{r matchings-plot, fig.path = IMAGE_PATH, fig.cap = "Plot of the matchings. Features belonging to the same group have been connected with a straight line.", fig.width = 6, fig.height = 3}
par(mar = c(5.1, 5, 4.1, 2.1))
angle_labels <- 30
addmatched <- unique(matchings$add_name)
xlim <- c(0, nrow(matchings) + 1)
ylim <- c(0, length(addmatched) + 1)
x <- match(matchings_groups[[1]][, "id"], matchings$id)
y <- match(matchings_groups[[1]][, "add_name"], addmatched)
plot(x, y, xlim = xlim, ylim = ylim, type = "l", xlab ="features", 
     ylab = "adducts", xaxt = "n", yaxt= "n")
axis(side = 1, at = 1:nrow(matchings), labels = FALSE)
axis(side = 2, at = 1:length(addmatched), labels = FALSE)
abline(v = 1:nrow(matchings), h = 1:length(addmatched), col = "gray85", lty = 3)
text(par("usr")[1], 1:length(addmatched), labels = addmatched, srt = 0, pos = 2,
     xpd = TRUE, cex= 0.7)
text(1:nrow(matchings), par("usr")[3]-1, labels = matchings$id, srt = 0, pos = 1, 
     xpd = TRUE, cex= 0.7, srt= 90)
points(x, y, pch = 16, 
       col=rainbow(15)[match(matchings_groups[[1]][, "name"], table1$name)])
text(x, y, labels = matchings_groups[[1]][, "name"], cex = 0.7, pos = 3, 
     srt = angle_labels)

for (i in 2: length(matchings_groups))
{
  x <- match(matchings_groups[[i]][, "id"], matchings$id)
  y <- match(matchings_groups[[i]][, "add_name"], addmatched)
  points(x, y, xlim = xlim, ylim = ylim, type= "l")
  points(x, y, pch = 16, 
         col=rainbow(15)[match(matchings_groups[[i]][, "name"], table1$name)])
  text(x, y, labels = matchings_groups[[i]][, "name"], cex= 0.7, pos = 3, 
       srt = angle_labels)
}

```
Here compare the retention time of adducts that are both in the found matchings 
and in the standards_dilution.txt file
```{r}
compare_rt <- NULL
for (i in 1:nrow(matchings))
{
  idx <- which(table1[,"name"] == matchings$name[i] & 
                 table1[,pol] == matchings$add_name[i])
  if(length(idx) > 0) 
    compare_rt=rbind(compare_rt, cbind(matchings[i, ], RT = table1[idx, "RT"]))
}

compare_rt <- cbind(compare_rt, sign_f[match(compare_rt$id,sign_f$sign_f_id),c("rtmed","rtmed_group")])
rownames(compare_rt)=1:nrow(compare_rt)
pandoc.table(compare_rt[,c("name", "add_name", "rtmed", "RT")], style = "rmarkdown")
```

```{r}
compound <- table1[,c("name", "formula", "RT", pol)]
feature_to_compound <- matchings[,c("id", "add_name", "name")]
colnames(feature_to_compound) <- c("feature_id", "add_name", "comp_name") 
feature <- data.frame(sign_f)
colnames(feature) <- c("id", "mzmed", "rtmed", "pvalue", "M", "f_group", "rtmed_group")

con <- dbConnect(RSQLite::SQLite(), ":memory:")
dbWriteTable(con, "feature", feature, append = TRUE)
dbWriteTable(con, "compound", compound, append = TRUE)
dbWriteTable(con, "feature_to_compound", feature_to_compound, 
             append = TRUE)
dbplyr::src_dbi(con)

dbplyr::src_dbi(con)
summary_table <- tbl(con, sql("SELECT compound.name, add_name, feature.id, 
                           feature.f_group, feature.rtmed, feature.M, 
                           feature.pvalue, compound.POS, compound.RT 
                  FROM compound
                  INNER JOIN feature_to_compound ON compound.name == comp_name
                  INNER JOIN feature ON feature_id == feature.id"))

summary_table
```

# Session information

The R version and packages used in this analysis are listed below.

```{r sessioninfo}
sessionInfo()
```
