---
title: "Preprocessing of the *standards* data set"
author: "Andrea Vicini, Johannes Rainer"
output:
  rmarkdown::html_document:
    highlight: pygments
    toc: true
    toc_depth: 3
    fig_width: 5
---

```{r style, echo = FALSE, results = 'asis', message = FALSE}
library(BiocStyle)
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

**Last modified:** `r file.info("preprocessing-standards.Rmd")$mtime`<br />
**Compiled**: `r date()`

# Introduction

In this document we perform the preprocessing and analysis of mzML files to
determine retention times and measured ions for the *standards*. These standards
are a collection of ~ 250 pure standards of polar metabolites which were used to
setup the HILIC-based LC-MS protocols to measure the polar metabolome in human
serum samples.


# Data preprocessing

In this section we perform the preprocessing of the LC-MS(/MS) data for all
mzML files of one mix of standards.

- [ ] load the *data/std_serum_files.txt* file.
- [ ] subset to the first mix (Mix01).
- [ ] load also the *data/standards_dilution.txt* file, subset to the first mix
  (column `"mix"`) and include a table with columns *name*, *formula*, *RT*, 
  *POS*, *NEG*.
- [ ] import the data (both POS and NEG into the same variable/object).
- [ ] create a base peak chromatogram (BPC) of all samples (use a different
  color for polarity and *matrix* (i.e. QC/serum and water)).
- [ ] perform chromatographic peak detection.
- [ ] table with the number of peaks per file.
- [ ] let's see how to proceed from here...


# Session information

The R version and packages used in this analysis are listed below.

```{r sessioninfo}
sessionInfo()
```
