---
title: "Dose-response"
author: "Mar Garcia-Aloy"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false
    toc_float: true
---

```{r settings, echo = FALSE, results = "hide", message = FALSE}
mixnum <- 11      # specify which MIX
polarity <- "NEG" # specify "POS" or "NEG"

#' Define the path where we can find the mzML files:
MZML_PATH <- "C:/Users/mgarciaaloy/Documents/mzML/"
#if (!file.exists(MZML_PATH))
#    stop("Can not find the directory with the mzML files: ", MZML_PATH)
```

# Preliminaries
```{r, message=FALSE}
source("R/listpeaks.R")

library(xcms)
library(reshape2)
library(ggplot2)
library(RColorBrewer)
library(Rdisop)
library(CompoundDb)
```

# Data import
In the file `standards_dilution_files.txt` there is the information regarding 
the injection sequence.
```{r}
injections <- read.table("data/standards_dilution_files.txt", # import the file
                         sep = "\t", header = TRUE, as.is = TRUE)
myfiles <- injections$mzML # get file names
myfiles <- myfiles[grep(polarity, myfiles)]# select files names of our polarity
(myfiles <- myfiles[grep(paste0("MIX ", mixnum, "[A-Z]"), myfiles)])
```

In the file `standards_dilution.txt` there is the information regarding 
the standards that are in the samples.
```{r}
std_info <- read.table("data/standards_dilution.txt",
                      sep = "\t", header = TRUE, as.is = TRUE)
std_info <- std_info[!is.na(std_info[, grep(polarity, colnames(std_info))]) &
                       !(is.na(std_info$RT)) &
                       std_info$mix == mixnum,]
std_info$mzneut = NA
for(i in seq(nrow(std_info))){
  if(grepl("C", std_info$formula[i])){std_info$mzneut[i] = 
    getMolecule(as.character(std_info$formula[i]))$exactmass}else{
      std_info$mzneut[i] = paste(std_info$formula[i])}
}
write.csv(std_info, "x.csv", row.names = FALSE)
std_info = read.csv("x.csv")
std_info
```

According to the information provided in the sheet `Dilution` of the file `Dynamic Range.xlxs` the concentrations of STDs in the samples are the following:
```{r}
ppm <- c(0.09765625, 0.09765625, 0.1953125, 0.1953125, 0.390625, 0.390625, 0.78125, 0.78125, 1.5625, 1.5625, 3.125, 3.125, 6.25, 6.25, 12.5, 12.5, 25, 25, 50, 50)
```


# Plots  
## Peak detection
Apply the `listpeaks()` function in order to detect the peaks of ISs:
```{r, warning=FALSE}
mzvalues <- c()
for(i in seq(nrow(std_info))){
  mzvalues <- c(mzvalues, 
                unlist(mass2mz(std_info$mzneut[i], 
                               adduct = 
                                 as.character(
                                   std_info[i, 
                                            grep(polarity, colnames(std_info))]
                                   ))))
}
rtvalues <- std_info$RT

peaktb <- listpeaks(file = paste0(MZML_PATH, myfiles),
                    mz = mzvalues,
                    param = CentWaveParam(peakwidth = c(2, 10), 
                                          ppm = 30))
peaktb$name <- std_info$name[peaktb$row]

# calculate the difference between "rtmed" and the expected retention time:
peaktb$rt_delta <- abs(std_info$RT[peaktb$row] - peaktb$rt)

# subset the peak matrix to rows (peaks) for which this difference 
# (the absolute value of the difference) is smaller than 60 seconds:
peaktb <- subset(peaktb, rt_delta < 60)

# among the subsetted peaks, choose the one with the highest intensity:
pl <- split(peaktb, paste(peaktb$row, peaktb$column))
pl <- lapply(pl, function(z){
  z[which.max(z$into),]
})
peaktb <- do.call(rbind, pl)

peaktb_wide <- dcast(peaktb, peaktb$name ~ peaktb$colum, value.var = "into")
peaktb_wide <- setNames(data.frame(t(peaktb_wide[,-1])), peaktb_wide[,1])
peaktb_wide[is.na(peaktb_wide)] <- 0

```

```{r}
getPalette = colorRampPalette(brewer.pal(9, "YlOrRd"))
mycols = getPalette(length(myfiles))
myplots = list()
for(i in seq(ncol(peaktb_wide))){
   tb <- data.frame(sample=seq(length(myfiles)), intensity=peaktb_wide[,i])
   tb$ppm <- ppm
   myplots[[i]] <- 
     ggplot(tb, aes(ppm, intensity)) + 
     geom_point(color = mycols) + theme_classic()  + 
     geom_smooth(method = "lm") + 
     ggtitle(colnames(peaktb_wide)[i]) + 
     geom_text(x = 2, y = max(tb$intensity), label = expression('R'^2*'=')) + 
     geom_text(x = 7, y =  max(tb$intensity), 
               label = round(summary(lm(peaktb_wide[,i]~ppm))$r.squared, 3))
}
if(mixnum<10){
  filename = 
    paste0("standard_dilution_dose_response_0", mixnum, "_", polarity, ".pdf")
} else if(mixnum>9){
    filename = 
    paste0("standard_dilution_dose_response_", mixnum, "_", polarity, ".pdf")
  }

ggsave(filename, 
       gridExtra::marrangeGrob(grobs = myplots, top="", 
                               layout_matrix = 
                                 rbind(c(1,2,3), c(4,5,6), c(7,8,9), 
                                       c(10,11,12), c(13,14,15))), 
       width=21, height=15)
```
