---
title: "Testing isotopologue function"
author: "Andrea Vicini"
output:
  rmarkdown::html_document:
    highlight: pygments
    toc: true
    toc_depth: 3
    fig_width: 5
---

```{r style, echo = FALSE, results = 'asis', message = FALSE}
library(BiocStyle)
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

**Last modified:** `r file.info("isotope-intensity-estimation.Rmd")$mtime`<br />
**Compiled**: `r date()`


```{r, echo = FALSE}
library(MetaboCoreUtils)
library(enviPat)
library(Rdisop)
library(MetaboCoreUtils)
library(MsCoreUtils)
data(isotopes)
subst_def <- read.table(paste0("~/lcms-standards/data/txt/isotope-intensity-",
                            "estimation/hmdb_subst.txt"), header = TRUE)
```

In this document I will test the isotopologues function firstly on simulated
peak matrices from true isotope patterns and then on measured spectra.

Before that, I define a tentative function to visualize the groups returned by 
isotopologues function.
```{r}
plot_groups <- function(x, i_groups = list(), cex = 0.7, pos = 3, xlim = NULL, ...){
  if(!is.null(xlim)){
    idxs <- which(x[, 1] >= xlim[1] & x[, 1] <= xlim[2])
  } else
    idxs <- seq_len(nrow(x))
  plot(x[idxs, ], type = "h", ylim = c(0, 1.1 * max(x[idxs, 2])), ...)
  col <- rainbow(length(i_groups))
  for(i in seq_along(i_groups)){
    group <- intersect(i_groups[[i]], idxs)
    points(x[group,], type = "h", col = col[i], ...)
    text(x[group,], labels = rep(i, length(group)), cex = cex, pos = pos) 
  }
}
```

## Testing on simulated peak matrices

Here I define another helper function to create peak matrices from isotope 
patterns of compounds with formula in `chemforms` (keeping for each of them the 
peaks with absolute intensity grater than `treshold`). The isotope patterns are 
concatenated one after the other and then sorted increasingly according to the 
mass.

```{r}
artificial_pm <- function(isotopes, chemforms, threshold = 0.001)
{
  iso_p <- isopattern(isotopes, chemforms , threshold = threshold, rel_to = 2)
  x <- do.call(rbind, lapply(iso_p, function(p) p[, 1:2]))
  frmls <- do.call(c, lapply(names(iso_p), function(nm) rep(nm, nrow(iso_p[[nm]]))))
  ord <- order(x[, "m/z"])
  x <- x[ord, ]
  frmls <- frmls[ord]
  colnames(x)<- c("mz", "intensity")
  data.frame(x, frmls)
}
```

We select a few compounds that are present in HMBD.

```{r}
smpl <- data.frame(formula = c("C32H62O4", "C21H20O13S", "C31H60O5", 
                               "C24H38O12", "C26H34O11", "C25H22O13"), 
                   exactmass = c(510.4648105, 512.0624619, 512.4440750, 
                                 518.2363267, 522.2101119, 530.1060408))
```

```{r, fig.width = 15, fig.height = 10}
threshold = 10^(-3)
x <- artificial_pm(isotopes, smpl$formula, threshold)
expected_groups <- lapply(smpl$formula, function(f) which(x[,"frmls"] == f))
i_groups <- isotopologues(x[, 1:2], substDefinition = subst_def, ppm = 1)

par(mfrow = c(2, 1))
plot_groups(x[, 1:2], i_groups, main = "groups found")
plot_groups(x[, 1:2], expected_groups, main = "expected groups")
i_groups
# expected_groups
```

All the peaks are correctly identified. We add some random noise to the peaks. 
The function seems to be able to separate the peaks from compounds from the 
noise most of the times and it still groups the peaks correctly. In general, 
the higher the ppm we select or the number of generated random peaks the higher 
is the chance that some of them are found to be compatible

```{r}
# write.table(x, file = "~/MetaboCoreUtils/tests/simulated_spectrum.txt",
#             sep = "\t", row.names = TRUE)
```

```{r, fig.width = 15, fig.height = 10}
set.seed(123); n_noise = 50
noise <- data.frame(mz = runif(n_noise, min = min(x$mz), max = max(x$mz)),
                    intensity = runif(n_noise, min(x$intensity), max(x$intensity)),
                    frmls = rep("noise", n_noise))
x_n <- rbind(x, noise)
x_n <- x_n[order(x_n$mz), ]
expected_groups_n <- lapply(smpl$formula, function(f) which(x_n[,"frmls"] == f))
i_groups <- isotopologues(x_n[, 1:2], substDefinition = subst_def, ppm = 1)

par(mfrow = c(2, 1))
plot_groups(x_n[, 1:2], i_groups, main = "groups found")
plot_groups(x_n[, 1:2], expected_groups_n, main = "expected groups")
i_groups
# expected_groups
```

In the previous examples the ppm was set to 1. With a higher ppm some problems 
may occurr as shown in the example below (everything is the same as before
but the tolerance has been set to a higher value). When two simulated isotope 
patterns are close it may happen that the one on the left (whose first peak is 
processed first) subtracts peaks from the one the left (which became compatible 
after increasing the ppm). From the plot below it is not clearly visible but 
from the indexes it is.

```{r, fig.width = 15, fig.height = 10}
i_groups <- isotopologues(x[, 1:2], substDefinition = subst_def, ppm = 20)

par(mfrow = c(2, 1))
plot_groups(x[, 1:2], i_groups, main = "groups found")
plot_groups(x[, 1:2], expected_groups, main = "expected groups")
i_groups
# expected_groups
```

Now I add a error on the mz values in the peak matrix. The result is affected 
by this operation in a significant way. Firstly we have to use a higher ppm in 
such a way that the function can match the new perturbed m/zs. But increasing 
the ppm doesn't always allow to capture all the peaks correctly 
as in the case without error addition. A reason can be the 
fact that some mz values related to different substitutions, say A and B, of 
the same compound are really close. When a small error is added to them the 
function might see as closest to the substitution A the peak related to 
substitution B and viceversa. So it expects that peak related to substitution B 
has intensity within the bounds of substitution A and as a result of that 
the peak is not selected. This happens in the example below with the peaks that 
have distance around +1 from the first peak.

```{r, fig.width = 15, fig.height = 10}
set.seed(1234)
x_n <- x
x_n$mz <- x_n$mz + rnorm(n = nrow(x_n), mean = 0, sd = 0.001)
x_n <- x_n[order(x_n$mz), ]
i_groups <- isotopologues(x_n[, 1:2], substDefinition = subst_def,
                           tolerance = 0, ppm = 20)

par(mfrow = c(2, 1))
plot_groups(x_n[, 1:2], i_groups, main = "groups found")
plot_groups(x_n[, 1:2], expected_groups, main = "expected groups")
i_groups
# expected_groups
```

Now we add an error on the intensity values in the peak matrix and it 
seems to affect the results less than a deviation on m/z.
```{r, fig.width = 15, fig.height = 10}
set.seed(123)
x_n <- x
x_n$intensity <- x_n$intensity * ( 1 + rnorm(n = nrow(x_n), mean = 0, sd = 0.05))
x_n <- x_n[order(x_n$mz), ] 
i_groups <- isotopologues(x_n[, 1:2], substDefinition = subst_def, ppm = 1)

par(mfrow = c(2, 1))
plot_groups(x_n[, 1:2], i_groups, main = "groups found")
plot_groups(x_n[, 1:2], expected_groups, main = "expected groups")
i_groups
# expected_groups
```

## Testing on real data spectra

Here I apply the function to some spectra containing signal of standards (which
are reported in the table below).

```{r, echo = FALSE, results = "asis"}
std_dilution <- read.table("data/standards_dilution.txt",
                           sep= "\t", header = TRUE)
table1 <- std_dilution[std_dilution$mix == 1, 
                       c("name", "formula", "RT", "POS", "NEG")]
pander::pandoc.table(table1, style = "rmarkdown",
             caption = paste0("Standards of ", 1))
```


```{r}
library(MSnbase)
fl <- "~/mix01/2020/2020_01/HighIS_Mix01_3_POS.mzML" 
data <- readMSData(fl, mode = "onDisk")
```

We select spectra containing each the signal of a given standard.

```{r}
# Xanthine: C5H4N4O2
sps_xan <- spectra(filterRt(data, rt = c(140, 141)))
pm_xan <- cbind(mz = sps_xan[[1]]@mz, intensity = sps_xan[[1]]@intensity)
# Acetylhistidine: C8H11N3O3
sps_ace <- spectra(filterRt(data, rt = c(180, 181)))
pm_ace <- cbind(mz = sps_ace[[1]]@mz, intensity = sps_ace[[1]]@intensity)
# Betaine:  C5H11NO2
sps_beta <- spectra(filterRt(data, rt = c(166, 167)))
pm_beta <- cbind(mz = sps_beta[[1]]@mz, intensity = sps_beta[[1]]@intensity)
# Creatine: C4H9N3O2
sps_crea <- spectra(filterRt(data, rt = c(173, 174)))
pm_crea <- cbind(mz = sps_crea[[1]]@mz, intensity = sps_crea[[1]]@intensity)
# Dimethylglycine: C4H9NO2
sps_dime <- spectra(filterRt(data, rt = c(177, 178)))
pm_dime <- cbind(mz = sps_dime[[1]]@mz, intensity = sps_dime[[1]]@intensity)
# L-Glutamic Acid: C5H9NO4
sps_lglu<- spectra(filterRt(data, rt = c(176, 177)))
pm_lglu <- cbind(mz = sps_lglu[[1]]@mz, intensity = sps_lglu[[1]]@intensity)
# Myo-Inositol: C6H12O6 | 193 | [M+Na]+ |
sps_myo<- spectra(filterRt(data, rt = c(193, 194)))
pm_myo <- cbind(mz = sps_myo[[1]]@mz, intensity = sps_myo[[1]]@intensity)
# 3-Phosphoglyceric Acid: C3H7O7P | 267 | [M+H]+  
sps_phos <- spectra(filterRt(data, rt = c(267, 268)))
pm_phos <- cbind(mz = sps_phos[[1]]@mz, intensity = sps_phos[[1]]@intensity)
# Suberic Acid:   C8H14O4    | 37  | [M+Na]+ 
sps_sub<- spectra(filterRt(data, rt = c(37, 38)))
pm_sub <- cbind(mz = sps_sub[[1]]@mz, intensity = sps_sub[[1]]@intensity)
```

Below we apply isotopologues on those spectra.
```{r} 
# iso_gr_xan <- isotopologues(pm_xan, substDefinition = subst_def, ppm = 10)
# iso_gr_ace <- isotopologues(pm_ace, substDefinition = subst_def, ppm = 10)
# iso_gr_beta <- isotopologues(pm_beta, substDefinition = subst_def, ppm = 10)
# iso_gr_crea <- isotopologues(pm_crea, substDefinition = subst_def, ppm = 10)
# 
# 
# par(mfrow=c(2, 2))
# plot_groups(pm_xan, iso_gr_xan, main = "xan ")
# plot_groups(pm_ace, iso_gr_ace, main = "ace")
# plot_groups(pm_beta, iso_gr_beta, main = "beta")
# plot_groups(pm_crea, iso_gr_crea, main = "crea")
```

We compute the mz of the selected standards.
```{r}
mz_xan <- mass2mz(getMolecule("C5H4N4O2")$exactmass, "[M+H]+")[1, 1]
mz_ace <- mass2mz(getMolecule("C8H11N3O3")$exactmass, "[M+H]+")[1, 1]
mz_beta <- mass2mz(getMolecule("C5H11NO2")$exactmass, "[M+H]+")[1, 1]
mz_crea <- mass2mz(getMolecule("C4H9N3O2")$exactmass, "[M+H]+")[1, 1]
mz_dime <- mass2mz(getMolecule("C4H9NO2")$exactmass, "[M+H]+")[1, 1]
mz_lglu <- mass2mz(getMolecule("C5H9NO4")$exactmass, "[M+H]+")[1, 1]
mz_myo <- mass2mz(getMolecule("C6H12O6")$exactmass, "[M+Na]+")[1, 1]
mz_phos <- mass2mz(getMolecule("C3H7O7P")$exactmass, "[M+H]+")[1, 1]
mz_sub <- mass2mz(getMolecule("C8H14O4")$exactmass, "[M+Na]+")[1, 1]
```

With the parameter `seedMz` we look for groups of peaks whose first peak is 
compatible with the mz of the selected standards. For some of them no group is
found.
```{r, fig.height = 8, fig.width=15}
iso_gr_xan <- isotopologues(pm_xan, substDefinition = subst_def, 
                             seedMz = mz_xan, ppm = 20)
iso_gr_ace <- isotopologues(pm_ace, substDefinition = subst_def,
                             seedMz = mz_ace, ppm = 20)
iso_gr_beta <- isotopologues(pm_beta, substDefinition = subst_def, 
                              seedMz = mz_beta, ppm = 20)
iso_gr_crea <- isotopologues(pm_crea, substDefinition = subst_def, 
                              seedMz = mz_crea, ppm = 20)
iso_gr_dime <- isotopologues(pm_dime, substDefinition = subst_def, 
                              seedMz = mz_dime, ppm = 20)
iso_gr_lglu <- isotopologues(pm_lglu, substDefinition = subst_def, 
                              seedMz = mz_lglu, ppm = 20)
iso_gr_myo <- isotopologues(pm_myo, substDefinition = subst_def, 
                              seedMz = mz_myo, ppm = 20)
iso_gr_phos <- isotopologues(pm_phos, substDefinition = subst_def, 
                              seedMz = mz_phos, ppm = 20)
iso_gr_sub <- isotopologues(pm_sub, substDefinition = subst_def, 
                              seedMz = mz_sub, ppm = 20)
par(mfrow=c(3, 3))
plot_groups(pm_xan, iso_gr_xan, xlim = mz_xan + c(- 0.5, 6), main = "Xanthine")
plot_groups(pm_ace, iso_gr_ace, xlim = mz_ace + c(- 0.5, 6), main = "Acetylhistidine") 
plot_groups(pm_beta, iso_gr_beta, xlim = mz_beta + c(- 0.5, 6), main = "Betaine")
plot_groups(pm_crea, iso_gr_crea, xlim = mz_crea + c(- 0.5, 6), main = "Creatine")
plot_groups(pm_dime, iso_gr_dime, xlim = mz_dime + c(- 0.5, 6), main = "Dimethylglycine")
plot_groups(pm_lglu, iso_gr_lglu, xlim = mz_lglu + c(- 0.5, 6), main = "L-Glutamic Acid")
plot_groups(pm_myo, iso_gr_myo, xlim = mz_myo + c(- 0.5, 6), main = "Myo-Inositol")
plot_groups(pm_phos, iso_gr_phos, xlim = mz_phos + c(- 0.5, 6), main = "3-Phosphoglyceric Acid")
plot_groups(pm_sub, iso_gr_sub, xlim = mz_sub + c(- 0.5, 6), main = "Suberic Acid")
```

Now we give the found peaks to Rdisop and see if it predicts the right 
formula. Since I think that Rdisop expects close peaks in the patter to be
combined, I define the following simple helper function to do that.

```{r}
combine_close_peaks <- function(x, tolerance = 0, ppm = 0) {
  gs <- MsCoreUtils::group(x[, 1], tolerance = tolerance, ppm = ppm)
  res_int <- tapply(x[, 2], gs, sum)
  res_mz <- tapply(x[, 1]*x[, 2], gs, sum)/res_int
  cbind(mz = res_mz, intensity = res_int)
}
```

```{r}
# Xanthine C5H4N4O2
res <- combine_close_peaks(pm_xan[iso_gr_xan[[1]], ], tolerance = 0.1)
head(decomposeIsotopes(res[, "mz"], res[, "intensity"], z = 1), 3)
```

```{r}
# Acetylhistidine C8H11N3O3
res <- combine_close_peaks(pm_ace[iso_gr_ace[[1]], ], tolerance = 0.1)
head(decomposeIsotopes(res[, "mz"], res[, "intensity"], z = 1), 3)
```

```{r}
# Betaine  C5H11NO2
res <- combine_close_peaks(pm_beta[iso_gr_beta[[1]], ], tolerance = 0.1)
head(decomposeIsotopes(res[, "mz"], res[, "intensity"], z = 1), 3)
```

For the case of Xanthine the right formula is not returned by Rdisop. For the 
case of Acetylhistidine the its formula with a additional H is among those 
returned by Rdisop though its score is very low. For the case of Betaine its 
formula is returned with a additional H and score 1.

<!---
```{r}
isop_xan_Rdisop <- getIsotope(getMolecule("C5H4N4O2", z = 1), seq(1,4))
head(decomposeIsotopes(isop_xan_Rdisop[1, ], isop_xan_Rdisop[2, ], z = 1), 3)
```


```{r}
isop_xan_enviPat <- isopattern(isotopes, "C5H4N4O2", rel_to = 2)[[1]][, 1:2]
head(decomposeIsotopes(isop_xan_enviPat[, "m/z"], 
                       isop_xan_enviPat[, "abundance"], z = 1), 3)
```


```{r}
isop_xan_combined <- combine_close_peaks(isop_xan_enviPat, tolerance = 0.1)
head(decomposeIsotopes(isop_xan_combined[, "mz"], 
                       isop_xan_combined[, "intensity"], z = 1), 3)
```
--->

# Session information

The R version and packages used in this analysis are listed below.

```{r sessioninfo}
sessionInfo()
```

