---
title: "Feature annotation"
subtitle: "Internal Standards: Matrix effect"
author: "Mar Garcia-Aloy"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false
    toc_float: true
---

```{r include = FALSE}
knitr::opts_chunk$set(echo=TRUE)

startpoint <- Sys.time()
```

# Preliminaries
## Parameters
```{r}
polarity <- "NEG" # specify "POS" or "NEG"
study <- "IS"     # specify "IS" or "QC"
mz.d <- 0.01
```

## Libraries
```{r, message=FALSE}
library(Rdisop)
library(CompoundDb)
```

## Functions
```{r}
source("https://raw.githubusercontent.com/EuracBiomedicalResearch/lcms-standards/master/R/which_within.R")
```

## Data import
```{r}
load(paste0("filter_group_", study, "_", polarity, ".RData"))
```

## Standards library
### IS basic
```{r}
std_info <- read.table(
  "https://raw.githubusercontent.com/EuracBiomedicalResearch/lcms-standards/master/data/internal_standards.txt",
  sep = "\t", header = TRUE, as.is = TRUE)
std_info <- std_info[!is.na(std_info[, grep(polarity, 
                                            colnames(std_info))]),]
std_info <- std_info[!grepl("Isoleucine", std_info$name),]
rownames(std_info) <- seq(nrow(std_info))

std_info$mzneut = NA
std_info$mz <- NA
for(i in seq(nrow(std_info))){
  if(grepl("C", std_info$formula[i])){std_info$mzneut[i] = 
    getMolecule(as.character(std_info$formula[i]))$exactmass}else{
      std_info$mzneut[i] = as.numeric(std_info$formula[i])}
  std_info$mz[i] <- unlist(
    mass2mz(std_info$mzneut[i],
            adduct = as.character(
              std_info[i, grep(polarity, colnames(std_info))])))
}
```

### IS isotopic pattern
```{r}
is_ip <- read.table(
  gsub("std_serum", "data/is_isotopic_pattern.txt", getwd()),
  sep = "\t", header = TRUE, as.is = TRUE)
is_ip <- is_ip[is_ip$polarity == polarity,]
```

### In-house identifications
```{r}
ih_cmp <- read.table(
  gsub("std_serum", "data/inhouse_cmps.txt", getwd()),
  sep = "\t", header = TRUE, as.is = TRUE)

ih_ion <- read.table(
  gsub("std_serum", "data/inhouse_ions.txt", getwd()),
  sep = "\t", header = TRUE, as.is = TRUE)

ih_feat <- read.table(
  gsub("std_serum", "data/inhouse_feat.txt", getwd()),
  sep = "\t", header = TRUE, as.is = TRUE)


ih_cmp$mzneut = NA
ih_cmp$mz <- NA
if(polarity == "NEG"){ ion = -1.007276 
} else if(polarity == "POS"){ ion = 1.007276 }

for(i in seq(nrow(ih_cmp))){
  
  if(grepl("C", ih_cmp$formula[i])){
    ih_cmp$mzneut[i] = 
      getMolecule(as.character(ih_cmp$formula[i]))$exactmass
  } else {
      ih_cmp$mzneut[i] = as.numeric(ih_cmp$formula[i])
      }
  
  ih_cmp$mz[i] <- ih_cmp$mzneut[i] + ion
}


ions.long = data.frame(matrix(ncol=5,nrow=0))
colnames(ions.long) = c("ID", "rt", "mode", "mz", "assignation")

ions.loop <- data.frame(matrix(ncol=5, nrow=nrow(ih_cmp)))
colnames(ions.loop) <- c("ID", "rt", "mode", "mz", "assignation")

ions.loop$ID <- ih_cmp$cmp_ID
ions.loop$rt <- ih_cmp$RT
ions.loop$mode <- polarity
ions.loop$mz <- ih_cmp$mz
if(polarity == "NEG"){
  ions.loop$assignation <- "[M-H]-"
}else if(polarity == "POS"){
    ions.loop$assignation = "[M+H]+"}
ions.long = rbind(ions.long, ions.loop)

for(i in 4:ncol(ih_ion)){
  for(j in seq(nrow(ih_ion))){
    if(ih_ion[j,i]!=0){
      ions.loop = data.frame(matrix(ncol=5,nrow=1))
      colnames(ions.loop) = c("ID", "rt", "mode", "mz", "assignation")
      ions.loop$ID = colnames(ih_ion)[i]
      ions.loop$rt = 
        ih_cmp$RT[which(ih_cmp$cmp_ID == colnames(ih_ion)[i])]
      ions.loop$mode <- polarity
      
      if(ih_ion$type[j] == "dimer"){
        ions.loop$mz = (ih_cmp[ih_cmp$cmp_ID == colnames(ih_ion)[i], 
                               "mzneut"])*2 + ion + ih_ion$delta[j]
        # close statment "if ion "j" is a dimer"
      } else if(ih_ion$type[j] == "trimer"){
        ions.loop$mz = (ih_cmp[ih_cmp$cmp_ID == colnames(ih_ion)[i], 
                               "mzneut"])*3 + ion + ih_ion$delta[j]
        # close statment "if ion "j" is a trimer"
      } else if(ih_ion$type[j] == "charged"){
        ions.loop$mz = (ih_cmp[ih_cmp$cmp_ID == colnames(ih_ion)[i], 
                               "mz"] + ion)/2
        ions.loop$assignation <- ih_ion$ion_id[j]
        # close statment "if ion `j` is a charged"
      }else {
        ions.loop$mz = (ih_cmp[ih_cmp$cmp_ID == colnames(ih_ion)[i], 
                               "mz"]) + ih_ion$delta[j]
      } # close statment "if ion "j" is NOT  dimer / trimer / charged"
      ions.loop$assignation <- ih_ion$ion_id[j]
      ions.long = rbind(ions.long, ions.loop)
    } # close statment "if ion "j" in cmp "i" is detected"...
  } # close loop ion "j"
} # close loop in-house compound "i"

if(polarity == "POS"){
  ions.long$assignation = gsub("\\*", "+", ions.long$assignation)
} else if(polarity == "NEG"){
  ions.long$assignation = gsub("\\*", "-", ions.long$assignation)
}
ions.long <- rbind(ions.long, ih_feat)

tmp <- std_info[which(!std_info$name %in% ih_cmp$comments), ]
tmp <- subset(tmp, select = c("name", "RT", "mzneut", "mz"))
tmp <- cbind(tmp, std_info[which(!std_info$name %in% ih_cmp$comments),
                           grep(polarity, colnames(std_info))])
colnames(tmp) <- colnames(ions.long)
tmp$mode <- polarity
ions.long <- rbind(tmp, ions.long)
```

# Annotation
```{r}
feat_filter <- feat_filter[order(feat_filter$cmp), ]
feat_filter$compound <- NA
feat_filter$annotation <- NA
```

## Isotopes, adducts, dimers & fragments
```{r}
feat_filter$isotope <- NA
feat_filter$dimer <- NA
feat_filter$adduct <- NA
feat_filter$fragment <- NA
feat_filter$charged <- NA

c <- levels(factor(feat_filter$cmp))

for(i in seq(length(levels(factor(feat_filter$cmp))))){
  c.i <- c[i]
  tmp <- feat_filter[feat_filter$cmp == c[i], ]
  
  if(nrow(tmp) > 1){
    tmp2 <- rbind(
      combn(tmp$mzmed, 2), # combine in pairs all mz of compound "z"
      abs(apply(combn(tmp$mzmed, 2), 2, diff))) # calculate their diff
    
    # Isotopes ---------------------------------------------------------
    ##### 13C
    # subset pairs with a relationship of 1.003355 (13C):
    tmp3 <- data.frame(tmp2[ ,
                             which(tmp2[3,] > (1.003355 - mz.d) & 
                                     tmp2[3,] < (1.003355 + mz.d))])
    if(ncol(tmp3) > 0){
      # consider only pairs in which the ion with higher mz is lower 
      # (less than the half) than the other
      mi <- c()
      for(j in seq(ncol(tmp3))){
        mi <- c(mi, 
                mean(data_NONimp[,rownames(tmp)[tmp$mzmed==tmp3[2,j]]] /
                       data_NONimp[,rownames(tmp)[tmp$mzmed==tmp3[1,j]]],
                     na.rm = TRUE))
      }
      tmp3 <- data.frame(tmp3[, which(mi < 0.5)])
      
      # write the annotations
      for(j in seq(ncol(tmp3))){
        tmp$isotope[tmp$mzmed == tmp3[1,j]] <- paste("i_13C", j, sep = "-")
        tmp$isotope[tmp$mzmed == tmp3[2,j]] <- paste("i_13C", j, sep = "-")
      }
      
      # add info to feat_filter matrix
      idx <- which(rownames(feat_filter) %in% rownames(tmp))
      feat_filter$isotope[idx] <- tmp$isotope
      
    }
    
    ##### 34S
    tmp3 <- data.frame(tmp2[ ,
                             which(tmp2[3,] > (1.995796 - mz.d) & 
                                     tmp2[3,] < (1.995796 + mz.d))])
    if(ncol(tmp3) > 0){
      mi <- c()
      for(j in seq(ncol(tmp3))){
        mi <- c(mi, 
                mean(data_NONimp[,rownames(tmp)[tmp$mzmed==tmp3[2,j]]] /
                       data_NONimp[,rownames(tmp)[tmp$mzmed==tmp3[1,j]]],
                     na.rm = TRUE))
      }
      tmp3 <- data.frame(tmp3[, which(mi < 0.5)])
      
      for(j in seq(ncol(tmp3))){
        tmp$isotope[tmp$mzmed == tmp3[1,j]] <- paste("i_34S", j, sep = "-")
        tmp$isotope[tmp$mzmed == tmp3[2,j]] <- paste("i_34S", j, sep = "-")
      }
      
      idx <- which(rownames(feat_filter) %in% rownames(tmp))
      feat_filter$isotope[idx] <- tmp$isotope
    }
    
    
    # Dimers -----------------------------------------------------------
    tmp3 <- combn((tmp$mzmed), 2) # combine in pairs all mz of cmp "z"
    tmp3[1, ] <- tmp3[1,] - 1.007276
    tmp3[1, ] <- (tmp3[1,]*2) + 1.007276
    tmp3 <- rbind(tmp3, abs(tmp3[1,] - tmp3[2,]))
    di <- which(tmp3[3,] > 0 & tmp3[3,] < 0.02)
    for(j in seq(length(di))){
      tmp$dimer[tmp$mzmed == tmp2[1, di[j]]] <- paste("d", j, sep = "-")
      tmp$dimer[tmp$mzmed == tmp2[2, di[j]]] <- paste("d", j, sep = "-")
    }
    idx <- which(rownames(feat_filter) %in% rownames(tmp))
    feat_filter$dimer[idx] <- tmp$dimer
    
    
    # Adducts ----------------------------------------------------------
    ##### NH4
    tmp3 <- data.frame(tmp2[ ,
                             which(tmp2[3,] > (17.02654 - mz.d) & 
                                     tmp2[3,] < (17.02654 + mz.d))])
    for(j in seq(ncol(tmp3))){
      tmp$adduct[tmp$mzmed == tmp3[1,j]] <- paste("a_NH4", j, sep = "-")
      tmp$adduct[tmp$mzmed == tmp3[2,j]] <- paste("a_NH4", j, sep = "-")
    }
    idx <- which(rownames(feat_filter) %in% rownames(tmp))
    feat_filter$adduct[idx] <- tmp$adduct
    
    ##### Na
    tmp3 <- data.frame(tmp2[ ,
                             which(tmp2[3,] > (21.98198 - mz.d) & 
                                     tmp2[3,] < (21.98198 + mz.d))])
    for(j in seq(ncol(tmp3))){
      tmp$adduct[tmp$mzmed == tmp3[1,j]] <- paste("a_Na", j, sep = "-")
      tmp$adduct[tmp$mzmed == tmp3[2,j]] <- paste("a_Na", j, sep = "-")
    }
    idx <- which(rownames(feat_filter) %in% rownames(tmp))
    feat_filter$adduct[idx] <- tmp$adduct
    
    ##### K
    tmp3 <- data.frame(tmp2[ ,
                             which(tmp2[3,] > (37.95589 - mz.d) & 
                                     tmp2[3,] < (37.95589 + mz.d))])
    for(j in seq(ncol(tmp3))){
      tmp$adduct[tmp$mzmed == tmp3[1,j]] <- paste("a_K", j, sep = "-")
      tmp$adduct[tmp$mzmed == tmp3[2,j]] <- paste("a_K", j, sep = "-")
    }
    idx <- which(rownames(feat_filter) %in% rownames(tmp))
    feat_filter$adduct[idx] <- tmp$adduct
    
    # Fragments --------------------------------------------------------
    ##### NH3
    tmp3 <- data.frame(tmp2[ ,
                             which(tmp2[3,] > (17.02654 - mz.d) & 
                                     tmp2[3,] < (17.02654 + mz.d))])
    for(j in seq(ncol(tmp3))){
      tmp$fragment[tmp$mzmed == tmp3[1,j]] <- paste("f_NH3", j, sep = "-")
      tmp$fragment[tmp$mzmed == tmp3[2,j]] <- paste("f_NH3", j, sep = "-")
    }
    idx <- which(rownames(feat_filter) %in% rownames(tmp))
    feat_filter$fragment[idx] <- tmp$fragment
    
    # Charged ----------------------------------------------------------
    #### C13
    tmp3 <- t(expand.grid(tmp$mzmed, tmp$mzmed))
    tmp3[1, ] <- (tmp3[1,] + ion + 1.003355) / 2
    tmp3 <- rbind(tmp3, abs(tmp3[1,] - tmp3[2,]))
    di <- which(tmp3[3,] > 0 & tmp3[3,] < 0.02)
    for(j in seq(length(di))){
      tmp$charged[tmp$mzmed == tmp3[1, di[j]]] <- paste("c13C", j, sep = "-")
      tmp$charged[tmp$mzmed == tmp3[2, di[j]]] <- paste("c13C", j, sep = "-")
    }
    idx <- which(rownames(feat_filter) %in% rownames(tmp))
    feat_filter$charged[idx] <- tmp$charged

    #### C12
    tmp3 <- t(expand.grid(tmp$mzmed, tmp$mzmed))
    tmp3[1, ] <- (tmp3[1,] + ion) / 2
    tmp3 <- rbind(tmp3, abs(tmp3[1,] - tmp3[2,]))
    di <- which(tmp3[3,] > 0 & tmp3[3,] < 0.02)
    for(j in seq(length(di))){
      tmp$charged[tmp$mzmed == tmp3[1, di[j]]] <- paste("c", j, sep = "-")
      tmp$charged[tmp$mzmed == tmp3[2, di[j]]] <- paste("c", j, sep = "-")
    }
    idx <- which(rownames(feat_filter) %in% rownames(tmp))
    feat_filter$charged[idx] <- tmp$charged
    
  }
}
```

## In-house compounds
```{r}
for(i in seq(nrow(feat_filter))){
  mz.i <- feat_filter$mzmed[i]
  rt.i <- feat_filter$rtmed[i]
  id.i <- ions.long[unlist(which_within(mz.i, ions.long$mz)), ]
  id.i <- id.i[unlist(which_within(rt.i, id.i$rt, mzd = 20)), ]
  if(nrow(id.i) > 1){
    id.i <- id.i[which(abs(((mz.i-id.i$mz)/id.i$mz)*1e6) < 5), ]
  }
  if(nrow(id.i) > 1){
    if(max(abs(rt.i - id.i$rt)) > 10){
      id.i <- id.i[which.min(abs(rt.i - id.i$rt)), ]
    }
  }
  
  if(nrow(id.i) == 1){
    feat_filter$compound[i] <- id.i$ID
    feat_filter$annotation[i] <- as.character(id.i$assignation)
  } else if(nrow(id.i) == 2){
    if(id.i$mz[1] == id.i$mz[2]){
      feat_filter$compound[i] <- 
        paste(ih_cmp$comments[ih_cmp$cmp_ID == id.i$ID[1]], "/", 
              ih_cmp$comments[ih_cmp$cmp_ID == id.i$ID[2]])
      feat_filter$annotation[i] <- as.character(
        paste(id.i$assignation[1], "/", id.i$assignation[2])
      )
    }
  }else if(nrow(id.i) > 2) {
    print(paste(i, rownames(feat_filter)[i], ": there is some CONFLICT"))
  }
}
```


## IS, isotopic pattern
I'm going to annotate their isotopes considering both the mz value 
& the abundance:
```{r}
for(i in seq(nrow(std_info))){
  cmp.i <- std_info$name[i]
  tmp <- feat_filter$cmp[which(feat_filter$compound == cmp.i)]
  tmp <- feat_filter[feat_filter$cmp == tmp, ]
  
  abc <- is_ip[is_ip$name == cmp.i, ]
  
  # check which experimental mz (tmp) are within theoretical mz (abc)
  def <- which_within(abc$mz, tmp$mzmed, mzd = 0.015)
  if(length(def)>0){
    for(j in seq(length(def))){
      names(def)[[j]] <- j
    }
    def <- def[lapply(def,length)>0]
    if(length(def)>0){
      ghi <- unlist(def)
      names(ghi) <- names(def)
    
      # check matching with abundance
      # calculate the proportion between mz.i & molecular ion (ft)
      ft <- rownames(tmp)[!is.na(tmp$annotation)]
      if(length(ft)>1){
        ft <- ft[which.max(tmp$int[rownames(tmp) %in% ft])]
      }
      
      if(length(def)>1){
        def <- # experimental abundances
          colMeans(data_NONimp[, rownames(tmp)[ghi]] / 
                     data_NONimp[,ft], na.rm = TRUE)*100
      } else if(length(def) == 1){
        def <- 
          mean(data_NONimp[, rownames(tmp)[ghi]]/data_NONimp[,ft], na.rm = TRUE)*100
      }
      deft <- # theorical abundances
        abc$abundance[as.numeric(names(ghi))]
    
      # the diferences in the proportional abundances is < 5?
      tmp$annotation[ghi][which(abs(def-deft) < 5)] <- 
        abc$annotation[as.numeric(names(ghi))][which(abs(def-deft) < 5)]
      tmp$compound[ghi][which(abs(def-deft) < 5)] <- 
        abc$name[as.numeric(names(ghi))][which(abs(def-deft) < 5)]
    
      tmp$annotation[ghi][which(is.na(deft))] <- 
        abc$annotation[as.numeric(names(ghi))][which(is.na(deft))]
      tmp$compound[ghi][which(is.na(deft))] <- 
        abc$name[as.numeric(names(ghi))][which(is.na(deft))]
    
      idx <- which(rownames(feat_filter) %in% rownames(tmp))
      feat_filter$annotation[idx] <- tmp$annotation
      feat_filter$compound[idx] <- tmp$compound
  }}
}
```

## Other checkings
Assegurar que tots els ions que tenen el mateix nom, estiguin en el mateix group
```{r}
c <- levels(factor(feat_filter$compound))
for(i in seq(length(c))){
  tmp <- feat_filter[!is.na(feat_filter$compound), ]
  tmp <- tmp[tmp$compound == c[i], ]
  if(length(levels(factor(tmp$cmp))) > 1){
    idx <- which(feat_filter$cmp %in% levels(factor(tmp$cmp)))
    feat_filter$cmp[idx] <- levels(factor(tmp$cmp))[1]
  }
  
}
length(levels(factor(feat_filter$cmp)))
```

Llistar els grups que tenen diferents compostos:
```{r}
c <- levels(factor(feat_filter$cmp))
for(i in seq(length(c))){
  tmp <- feat_filter[!is.na(feat_filter$cmp), ]
  tmp <- tmp[tmp$cmp == c[i], ]
  if(length(levels(factor(tmp$compound))) > 1){
    print(paste(i, levels(factor(tmp$cmp)), levels(factor(tmp$compound))))
  }
}
```

Indicate compound name for identified compounds
```{r}
tmp <- ih_cmp[ih_cmp$identified == "YES", ]
tmp <- tmp[which(tmp$cmp_ID %in% feat_filter$compound), ]
for(i in seq(nrow(tmp))){
  idx <- which(feat_filter$compound == tmp$cmp_ID[i])
  feat_filter$compound[idx] <- tmp$comments[i]
}
```

```{r}
if(polarity == "NEG"){
  feat_filter$annotation <- gsub("\\[2M\\+Na\\]\\+", "[2M+Na-2H]-", feat_filter$annotation)
}
```


# Table
```{r}
tb.ann <- 
  data.frame(matrix(ncol = 5, 
                    nrow = length(levels(factor(feat_filter$cmp)))))
colnames(tb.ann) <- c("C", "Compound", "RT", "Annotations", "Comments")
c <- levels(factor(feat_filter$cmp))
for(i in seq(nrow(tb.ann))){
  tb.ann$C[i] <- c[i]
  tmp <- feat_filter[feat_filter$cmp == c[i], ]
  if(length(tmp$compound[!is.na(tmp$compound)]) > 0){
    tb.ann$Compound[i] <- tmp$compound[!is.na(tmp$compound)][1]
  }
  tb.ann$RT[i] <- round(mean(tmp$rtmed))
  tb.ann$Annotations[i] <- paste(round(tmp$mzmed, 4), tmp$annotation, 
                                 collapse = "; ")
  if(length(which(ih_cmp$cmp_ID==tb.ann$Compound[i]))>0){
    tb.ann$Comments[i] <- 
      ih_cmp$comments[which(ih_cmp$cmp_ID==tb.ann$Compound[i])]
  }
  if(tb.ann$Compound[i] %in% std_info$name){
    tb.ann$Comments[i] <- "STD"
  }
}
```


# Save output
```{r}
ls()
save(list = c(
  # XCMS data
  "raw_data", "xdata", "MZML_PATH", "myfiles",
  "cwp", "mnp", "pdp1", "pdp2", "pgp", "fcp",
  
  "data_imp", "data_NONimp", "feat",
  
  # Annotations
  "feat_filter", "tb.ann"),
  file = paste0("annotations_", study, "_", polarity, ".RData"))
```


# Session information
```{r}
Sys.time()-startpoint

devtools::session_info()
```

