---
title: "Formula finder"
author: "Mar Garcia-Aloy, Vinicius Veri"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false
    toc_float: true
---

```{r startpoint, include = FALSE}
startpoint <- Sys.time()
```

In this code we are going to test the efficiency of the designed code 
to deduce the molecular formula of specific ions.
We'll do that using the data from the experiment done by Vinicius Veri on January-2020. 


# Experiment description

Three replicates of the following samples were injected in the same 
order as they are named in the list below:  

- **`LowIS_Mix`**: Mix of standards prepared in Water/ACN (50:50) at 5ppm.     
- **`HighIS_Mix`**: Mix of standards prepared in Water/ACN (50:50) at 50ppm.  
- **`QC`**: QC samples (serum) without STD.  
- **`QC_LowIS_Mix`**: QC samples (serum) spicked with LowIS_Mix solution.  
- **`QC_HighIS_Mix`**: QC samples (serum) spicked with HighIS_Mix solution.  


# Preliminaries

## Parameters

```{r parameters}
## MZML_PATH <- "C:/Users/mgarciaaloy/Documents/mzML/"
## MZML_PATH <- "C:/Users/vveri/Documents/stds_ID_files_trial/"
MZML_PATH <- "/data/massspec/mzML/"

polarity.all <- c("POS", "NEG") 
```

## Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(xcms)
library(Rdisop)
library(CompoundDb)
library(doParallel)
library(magrittr)
library(SummarizedExperiment)
library(CHNOSZ)
library(kableExtra)

C_rule <- function(x, error = 0.1){
  C <- round(c((x/1.1) - ((x/1.1)*error), (x/1.1) + ((x/1.1)*error)))
  return(C)
}
H_rule <- function(C, N){
  H <- 2*C + N + 2
  return(H)
}
RPU_rule <- function(C=0, H=0, N=0, P=0, Si=0, Cl=0, Fl=0, I=0){
  RPU <- (C+Si) - (H+Cl+Fl+I)/2 + (N+P)/2 + 1
  return(RPU)
}
```


## Parallel processing

```{r parallel-proc}
## ncores <- detectCores()-1
## registerDoParallel(ncores)
## register(DoparParam(), default = TRUE)
register(SerialParam())
```



# Data import

## Theoretical info

```{r std-info}
std_info <- read.table(
  "../data/standards_dilution.txt",
  sep = "\t", header = TRUE, as.is = TRUE)

std_info$mz_POS <- NA
std_info$mz_NEG <- NA
for(i in 1:nrow(std_info)){
  std_info$mz_POS[i] <- unlist(
    mass2mz(getMolecule(as.character(std_info$formula[i]))$exactmass, 
            adduct = as.character(std_info$POS[i])))
  std_info$mz_NEG[i] <- unlist(
    mass2mz(getMolecule(as.character(std_info$formula[i]))$exactmass, 
            adduct = as.character(std_info$NEG[i])))
}
std_info$form_pred_POS <- NA
std_info$form_pred_NEG <- NA

mix.levels <- levels(factor(std_info$mix))
```


## Experimental info

```{r injections}
# Import table with file names:
injections <- read.table("../data/std_serum_files.txt", 
                         sep = "\t", header = TRUE, as.is = TRUE)

# Select files adquired in MS1:
injections <- injections[injections$mode == "FS", ]

# Focus only on the 2nd replicated of "HighIS_Mix" samples:
injections <- injections[grep("Water_Mix", injections$class), ]
injections <- injections[grep("_High", injections$class), ]
injections <- injections[grep("_2_", injections$mzML), ]
```


# Output

```{r output, warning = FALSE, echo = FALSE, message = FALSE}
cwp <- CentWaveParam(
  ppm = 50,
  peakwidth = c(2, 20),
  snthresh = 10,
  mzdiff = 0.001,
  prefilter = c(3, 500),
  noise = 100,
  integrate = 2)

for(k in 3#:(length(mix.levels))
){
  
  # Select STDs used in mix "k"
  std_info.i <- std_info[std_info$mix == mix.levels[k], ]
  if(mix.levels[k] == 5){
    std_info.i <- std_info.i[std_info.i$name != 
                               "Glycerophospho-inositol", ]
  } else if(mix.levels[k] == 9){
    std_info.i <- std_info.i[std_info.i$name != 
                               "Eplerone", ]
  } else if(mix.levels[k] == 17){
    std_info.i <- std_info.i[std_info.i$name != 
                               "3-Hydroxy-DL-kynurenine", ]
  }
  
  for (polarity in polarity.all) {
    if(polarity == "NEG"){ 
      pol <- -1
      ion <- -1.007276 # Assuming a protonated ion? always?
    } else if(polarity == "POS"){ 
      pol <- 1
      ion <- 1.007276 
    }
    idx <- injections$type == paste0(
      "Mix",
      sprintf(paste0("%0", ceiling(log10(10 + 1L)), "d"), 1:20)[k])
    filename <- paste0(injections$folder[idx], "/", injections$mzML[idx])
    #filename <- injections$mzML[idx]
    filename <- filename[grep(polarity, filename)]
    data_raw <- readMSData(paste0(MZML_PATH, filename),
                           mode = "onDisk")
    std_info.i$mz <- std_info.i[, paste0("mz_", polarity)]
    for(i in 1:(nrow(std_info.i))){
      if(!is.na(std_info.i$mz[i])){
        eic <- chromatogram(data_raw, aggregationFun = "max", 
                            mz = std_info.i$mz[i] + 0.01 * c(-1, 1))
        pks <- data.frame(chromPeaks(findChromPeaks(eic, param = cwp)))
        if(any(which((pks$rtmin - 10) < std_info.i$RT[i] & 
                     (pks$rtmax + 10) > std_info.i$RT[i]))){
          pks <- pks[which((pks$rtmin - 10) < std_info.i$RT[i] & 
                             (pks$rtmax + 10) > std_info.i$RT[i]), ]
          if(nrow(pks) > 1){
            pks <- pks[which.max(pks$maxo), ]
          }

          ## Have one function that returns the full scan for the peak position
          ## of a chrom peak. chromPeakFullScan -> returns a Spectra with one
          ## Spectrum per chromPeak.
          ## Have a helper function that takes an rt as input and returns the
          ## index of the matching spectrum. which.closest
          sps <- data_raw[[closest(pks$rt, rtime(data_raw),
                                   duplicates = "closest")]]
          sps2 <- as.data.frame(sps)
          ## sps <- data_raw %>%
          ##   filterRt(rt = pks$rt + 0.5 * c(-1,1)) %>%
          ##   spectra
          ## sps2 <- data.frame(sps[[2]])
          ## <- stops here.
          
          sps2$i100 <- (sps2$i / max(sps2$i))*100

          ## Have a function to get isotope pattern from a spectrum providing
          ## a starting m/z: isotopeIndex - could simply use closest for that?
          idx <- c(which.min((abs(std_info.i$mz[i] - sps2$mz) / 
                                std_info.i$mz[i])*1e6),
                   which.min((abs((std_info.i$mz[i] + 1.003355) - 
                                    sps2$mz) / 
                                (std_info.i$mz[i] + 1.003355))*1e6),
                   which.min((abs((std_info.i$mz[i] + 1.003355*2) - 
                                    sps2$mz) / 
                                (std_info.i$mz[i] + 1.003355*2))*1e6))
          sps2 <- sps2[idx, ]
          masses <- sps2$mz
          intensities <- sps2$i
          ## <- stops here.

          ## This below assumes we know already the formula?
          ## How could we do this if we don't know the formula yet? Run it
          ## with all elements initialized?
          if(grepl("Na", std_info.i[i, polarity])){
            molecules <- decomposeIsotopes(masses, intensities, ppm = 20,
                                           elements = initializeElements(
                                             c("C", "H", "N", "O", 
                                               "P", "S", "Na")))
          }else{
            molecules <- decomposeIsotopes(masses, intensities, ppm = 20)
          }

          ## Function that takes a formula as input and returns element counts.
          formulas <- data.frame(formula = getFormula(molecules))
          ## Use lapply(as.character(formulas$formula), countElements)
          tmp <- makeup(as.character(formulas$formula)) ## maybe use MetabCoreUtils?
          if(class(tmp) == "numeric"){
            formulas$C <- tmp["C"]
            formulas$H <- tmp["H"]
            formulas$O <- tmp["O"]
            formulas$N <- tmp["N"]
            formulas$S <- tmp["S"]
            formulas$P <- tmp["P"]
            formulas$Na <- tmp["Na"]
            formulas[is.na(formulas)] <- 0
          } else{
            formulas$C <- NA
            formulas$H <- NA
            formulas$O <- NA
            formulas$N <- NA
            formulas$S <- NA
            formulas$P <- NA
            formulas$Na <- NA
            for(j in 1:nrow(formulas)){
              if(length(tmp[[j]][names(tmp[[j]]) == "C"]) == 1){
                formulas$C[j] <- tmp[[j]][names(tmp[[j]]) == "C"]  
              } else {formulas$C[j] <- 0}
              if(length(tmp[[j]][names(tmp[[j]]) == "H"]) == 1){
                formulas$H[j] <- tmp[[j]][names(tmp[[j]]) == "H"]
              } else {formulas$H[j] <- 0}
              if(length(tmp[[j]][names(tmp[[j]]) == "O"]) == 1){
                formulas$O[j] <- tmp[[j]][names(tmp[[j]]) == "O"]
              } else {formulas$O[j] <- 0}
              if(length(tmp[[j]][names(tmp[[j]]) == "N"]) == 1){
                formulas$N[j] <- tmp[[j]][names(tmp[[j]]) == "N"]
              } else {formulas$N[j] <- 0}
              if(length(tmp[[j]][names(tmp[[j]]) == "S"]) == 1){
                formulas$S[j] <- tmp[[j]][names(tmp[[j]]) == "S"]
              } else {formulas$S[j] <- 0}
              if(length(tmp[[j]][names(tmp[[j]]) == "P"]) == 1){
                formulas$P[j] <- tmp[[j]][names(tmp[[j]]) == "P"]
              } else {formulas$P[j] <- 0}
              if(length(tmp[[j]][names(tmp[[j]]) == "Na"]) == 1){
                formulas$Na[j] <- tmp[[j]][names(tmp[[j]]) == "Na"]
              } else {formulas$Na[j] <- 0}
            } # close formula "j"
          }
          ## end here.

          ## NEW CODE: check if formulas match the expected intensities.
          formulas$C_rule <- FALSE
          if(intensities[1] < 1e6){
            tmp.error <- 0.4} else {tmp.error <- 1.0}
          tmp <- C_rule((intensities[2] / intensities[1])*100, tmp.error)
          formulas$C_rule[formulas$C >= tmp[1] & 
                            formulas$C <= tmp[2]] <- TRUE
          formulas$H_rule <- FALSE
          formulas$H_rule[formulas$H <= 
                            H_rule(formulas$C, formulas$N)] <- TRUE
          formulas$N_rule <- 
            formulas$N %% 2 == round(masses[1] - ion) %% 2
          formulas$RPU_rule <- FALSE
          formulas$RPU_rule[RPU_rule(
            C = formulas$C, H = formulas$H - pol, 
            N = formulas$N, P = formulas$P) >= 0] <- TRUE
          if(grepl("Na", std_info.i[i, polarity])){
            formulas$RPU_rule[RPU_rule(
              C = formulas$C, H = formulas$H, 
              N = formulas$N, P = formulas$P) %% 1 > 0] <- FALSE
          } else{
            formulas$RPU_rule[RPU_rule(
              C = formulas$C, H = formulas$H - pol, 
              N = formulas$N, P = formulas$P) %% 1 > 0] <- FALSE 
          }

          ## Also here: what if we don't have the formula already???
          if(grepl("Na", std_info.i[i, polarity])){
            formulas.ok <- 
              formulas[formulas$C_rule & formulas$H_rule & 
                         formulas$N_rule & formulas$RPU_rule &
                         formulas$Na == 1, ]
          }else{
            formulas.ok <- 
              formulas[formulas$C_rule & formulas$H_rule & 
                         formulas$N_rule & formulas$RPU_rule, ]
          }
          ## end here.
          
          if(nrow(formulas.ok) > 0){
            formulas.ok$isodev <- NA
            for(j in 1:nrow(formulas.ok)){
              formulas.ok$isodev[j] <- mean(abs(
                ((getIsotope(getMolecule(as.character(
                  formulas.ok$formula[j])), seq(1,3))[2,] / 
                    max(getIsotope(getMolecule(as.character(
                      formulas.ok$formula[j])), seq(1,3))[2,]))*100)[2] - 
                  (intensities / max(intensities)*100)[2]))
            } # close formula "j"
            formulas.ok <- formulas.ok[order(formulas.ok$isodev), ]
            
            if(nrow(formulas.ok) < 6){
              std_info[
                which(std_info$name == std_info.i$name[i]),
                paste0("form_pred_", polarity)] <- 
                paste(formulas.ok$formula, collapse = "; ") 
            } else{
              std_info[
                which(std_info$name == std_info.i$name[i]),
                paste0("form_pred_", polarity)] <- 
                paste0(paste(formulas.ok$formula[1:5], 
                             collapse = "; "), "; etc.")
            }
          } # close: if(nrow(formulas.ok) > 0)
          rm(pks)
        } # close if any pks
      } # close: if(!is.na(std_info.i$mz[i]))
    } # close std "i"
  } # close polarity "z"
} # close mix "k"
```

```{r}
#' @param x `numeric` with m/z values.
#' 
#' @param mz `numeric` with the m/z value of the main ion for which isotope
#'     peaks should be searched for.
#'
#' @return `integer` with the index of the main ion peak and its isotope peaks.
isotopeIndex <- function(x, mz = numeric()) {
}

```


# Table

```{r table, warning=FALSE}
kable(std_info[,c(1:3,5:8,13,14,17,18)]) %>%
  kable_styling(bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive"))
```


# Session information

```{r session}
Sys.time()-startpoint

devtools::session_info()
```
