---
title: "Formula finder"
author: "Mar Garcia-Aloy, Vinicius Veri"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false
    toc_float: true
---

```{r startpoint, include = FALSE}
startpoint <- Sys.time()
```

In this code we are going to test the efficiency of the designed code 
to deduce the molecular formula of specific ions.
We'll do that using the data from the experiment done by Vinicius Veri on January-2020. 


# Experiment description

Three replicates of the following samples were injected in the same 
order as they are named in the list below:  

- **`LowIS_Mix`**: Mix of standards prepared in Water/ACN (50:50) at 5ppm.     
- **`HighIS_Mix`**: Mix of standards prepared in Water/ACN (50:50) at 50ppm.  
- **`QC`**: QC samples (serum) without STD.  
- **`QC_LowIS_Mix`**: QC samples (serum) spicked with LowIS_Mix solution.  
- **`QC_HighIS_Mix`**: QC samples (serum) spicked with HighIS_Mix solution.  


# Preliminaries

## Parameters

```{r parameters}
MZML_PATH <- "C:/Users/mgarciaaloy/Documents/mzML/"
## MZML_PATH <- "C:/Users/vveri/Documents/stds_ID_files_trial/"
## MZML_PATH <- "/data/massspec/mzML/"

polarity.all <- c("POS", "NEG") 
```

## Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(xcms)
library(Rdisop)
library(CompoundDb)
library(doParallel)
library(magrittr)
library(SummarizedExperiment)
library(kableExtra)
library(MsCoreUtils)
library(MetaboCoreUtils)

C_rule <- function(x, error = 0.1){
  C <- round(c((x/1.1) - ((x/1.1)*error), (x/1.1) + ((x/1.1)*error)))
  return(C)
}
H_rule <- function(C, N){
  H <- 2*C + N + 2
  return(H)
}
RPU_rule <- function(C=0, H=0, N=0, P=0, Si=0, Cl=0, Fl=0, I=0){
  RPU <- (C+Si) - (H+Cl+Fl+I)/2 + (N+P)/2 + 1
  return(RPU)
}
```


## Parallel processing

```{r parallel-proc}
## ncores <- detectCores()-1
## registerDoParallel(ncores)
## register(DoparParam(), default = TRUE)
register(SerialParam())
```



# Data import

## Theoretical info

```{r std-info}
std_info <- read.table(
  "../data/standards_dilution.txt",
  sep = "\t", header = TRUE, as.is = TRUE)

std_info$mz_POS <- NA
std_info$mz_NEG <- NA
for(i in 1:nrow(std_info)){
  std_info$mz_POS[i] <- unlist(
    CompoundDb::mass2mz(getMolecule(as.character(std_info$formula[i]))$exactmass, 
            adduct = as.character(std_info$POS[i])))
  std_info$mz_NEG[i] <- unlist(
    CompoundDb::mass2mz(getMolecule(as.character(std_info$formula[i]))$exactmass, 
            adduct = as.character(std_info$NEG[i])))
}
std_info$form_pred_POS <- NA
std_info$form_pred_NEG <- NA

mix.levels <- levels(factor(std_info$mix))
```


## Experimental info

```{r injections}
# Import table with file names:
injections <- read.table("../data/std_serum_files.txt", 
                         sep = "\t", header = TRUE, as.is = TRUE)

# Select files adquired in MS1:
injections <- injections[injections$mode == "FS", ]

# Focus only on the 2nd replicated of "HighIS_Mix" samples:
injections <- injections[grep("Water_Mix", injections$class), ]
injections <- injections[grep("_High", injections$class), ]
injections <- injections[grep("_2_", injections$mzML), ]
```


# Output

```{r output, warning = FALSE, echo = FALSE, message = FALSE}
cwp <- CentWaveParam(
  ppm = 50,
  peakwidth = c(2, 20),
  snthresh = 10,
  mzdiff = 0.001,
  prefilter = c(3, 500),
  noise = 100,
  integrate = 2)

for(k in 4#:(length(mix.levels))
){
  
  # Select STDs used in mix "k"
  std_info.i <- std_info[std_info$mix == mix.levels[k], ]
  if(mix.levels[k] == 5){
    std_info.i <- std_info.i[std_info.i$name != 
                               "Glycerophospho-inositol", ]
  } else if(mix.levels[k] == 9){
    std_info.i <- std_info.i[std_info.i$name != 
                               "Eplerone", ]
  } else if(mix.levels[k] == 17){
    std_info.i <- std_info.i[std_info.i$name != 
                               "3-Hydroxy-DL-kynurenine", ]
  }
  
  for (polarity in polarity.all) {
    if(polarity == "NEG"){ 
      pol <- -1
      ion <- -1.007276 
    } else if(polarity == "POS"){ 
      pol <- 1
      ion <- 1.007276 
    }
    idx <- injections$type == paste0(
      "Mix",
      sprintf(paste0("%0", ceiling(log10(10 + 1L)), "d"), 1:20)[k])
    filename <- paste0(injections$folder[idx], "/", injections$mzML[idx])
    #filename <- injections$mzML[idx]
    filename <- filename[grep(polarity, filename)]
    data_raw <- readMSData(paste0(MZML_PATH, filename),
                           mode = "onDisk")
    std_info.i$mz <- std_info.i[, paste0("mz_", polarity)]
    for(i in 1:(nrow(std_info.i))
        ){
      if(!is.na(std_info.i$mz[i])){
        eic <- chromatogram(data_raw, aggregationFun = "max", 
                            mz = std_info.i$mz[i] + 0.01 * c(-1, 1))
        pks <- data.frame(chromPeaks(findChromPeaks(eic, param = cwp)))
        if(any(which((pks$rtmin - 30) < std_info.i$RT[i] & 
                     (pks$rtmax + 30) > std_info.i$RT[i]))){
          pks <- pks[which((pks$rtmin - 30) < std_info.i$RT[i] & 
                             (pks$rtmax + 30) > std_info.i$RT[i]), ]
          if(nrow(pks) > 1){
            pks <- pks[which.max(pks$maxo), ]
          }
          sps <- data_raw[[closest(pks$rt, rtime(data_raw),
                                   duplicates = "closest")]]
          sps2 <- as.data.frame(sps)
          sps2$i100 <- (sps2$i / max(sps2$i))*100
          
          idx <- c(which.min((abs(std_info.i$mz[i] - sps2$mz) / 
                                std_info.i$mz[i])*1e6),
                   which.min((abs((std_info.i$mz[i] + 1.003355) - 
                                    sps2$mz) / 
                                (std_info.i$mz[i] + 1.003355))*1e6),
                   which.min((abs((std_info.i$mz[i] + 1.003355*2) - 
                                    sps2$mz) / 
                                (std_info.i$mz[i] + 1.003355*2))*1e6))
          sps2 <- sps2[idx, ]
          masses <- sps2$mz
          intensities <- sps2$i
          if(grepl("Na", std_info.i[i, polarity])){
            molecules <- decomposeIsotopes(masses, intensities, ppm = 20,
                                           elements = initializeElements(
                                             c("C", "H", "N", "O", 
                                               "P", "S", "Na")))
          }else{
            molecules <- decomposeIsotopes(masses, intensities, ppm = 20)
          }
          formulas <- data.frame(formula = getFormula(molecules))
          tmp <- lapply(as.character(formulas$formula), countElements)
          if(class(tmp) == "numeric"){
            formulas$C <- tmp["C"]
            formulas$H <- tmp["H"]
            formulas$O <- tmp["O"]
            formulas$N <- tmp["N"]
            formulas$S <- tmp["S"]
            formulas$P <- tmp["P"]
            formulas$Na <- tmp["Na"]
            formulas[is.na(formulas)] <- 0
          } else{
            formulas$C <- NA
            formulas$H <- NA
            formulas$O <- NA
            formulas$N <- NA
            formulas$S <- NA
            formulas$P <- NA
            formulas$Na <- NA
            for(j in 1:nrow(formulas)){
              if(length(tmp[[j]][names(tmp[[j]]) == "C"]) == 1){
                formulas$C[j] <- tmp[[j]][names(tmp[[j]]) == "C"]  
              } else {formulas$C[j] <- 0}
              if(length(tmp[[j]][names(tmp[[j]]) == "H"]) == 1){
                formulas$H[j] <- tmp[[j]][names(tmp[[j]]) == "H"]
              } else {formulas$H[j] <- 0}
              if(length(tmp[[j]][names(tmp[[j]]) == "O"]) == 1){
                formulas$O[j] <- tmp[[j]][names(tmp[[j]]) == "O"]
              } else {formulas$O[j] <- 0}
              if(length(tmp[[j]][names(tmp[[j]]) == "N"]) == 1){
                formulas$N[j] <- tmp[[j]][names(tmp[[j]]) == "N"]
              } else {formulas$N[j] <- 0}
              if(length(tmp[[j]][names(tmp[[j]]) == "S"]) == 1){
                formulas$S[j] <- tmp[[j]][names(tmp[[j]]) == "S"]
              } else {formulas$S[j] <- 0}
              if(length(tmp[[j]][names(tmp[[j]]) == "P"]) == 1){
                formulas$P[j] <- tmp[[j]][names(tmp[[j]]) == "P"]
              } else {formulas$P[j] <- 0}
              if(length(tmp[[j]][names(tmp[[j]]) == "Na"]) == 1){
                formulas$Na[j] <- tmp[[j]][names(tmp[[j]]) == "Na"]
              } else {formulas$Na[j] <- 0}
            } # close formula "j"
          }
          
          formulas$C_rule <- FALSE
          #if(intensities[1] < 1e6){
            #tmp.error <- 0.4} else {tmp.error <- 1.0}
          if(as.data.frame(sps)$i[as.data.frame(sps)$mz == 
                                    sps2$mz[1]] < 1e6){tmp.error <- 0.4
                                    } else {tmp.error <- 1.0}
          tmp <- C_rule((intensities[2] / intensities[1])*100, tmp.error)
          formulas$C_rule[formulas$C >= tmp[1] & 
                            formulas$C <= tmp[2]] <- TRUE
          formulas$H_rule <- FALSE
          formulas$H_rule[formulas$H <= 
                            H_rule(formulas$C, formulas$N)] <- TRUE
          formulas$N_rule <- 
            formulas$N %% 2 == round(masses[1] - ion) %% 2
          formulas$RPU_rule <- FALSE
          formulas$RPU_rule[RPU_rule(
            C = formulas$C, H = formulas$H - pol, 
            N = formulas$N, P = formulas$P) >= 0] <- TRUE
          if(grepl("Na", std_info.i[i, polarity])){
            formulas$RPU_rule[RPU_rule(
              C = formulas$C, H = formulas$H, 
              N = formulas$N, P = formulas$P) %% 1 > 0] <- FALSE
          } else{
            formulas$RPU_rule[RPU_rule(
              C = formulas$C, H = formulas$H - pol, 
              N = formulas$N, P = formulas$P) %% 1 > 0] <- FALSE 
          }
          
          if(grepl("Na", std_info.i[i, polarity])){
            formulas.ok <- 
              formulas[formulas$C_rule & formulas$H_rule & 
                         formulas$N_rule & formulas$RPU_rule &
                         formulas$Na == 1, ]
          }else{
            formulas.ok <- 
              formulas[formulas$C_rule & formulas$H_rule & 
                         formulas$N_rule & formulas$RPU_rule, ]
          }
          if(nrow(formulas.ok) > 0){
            formulas.ok$isodev <- NA
            for(j in 1:nrow(formulas.ok)){
              formulas.ok$isodev[j] <- mean(abs(
                ((getIsotope(getMolecule(as.character(
                  formulas.ok$formula[j])), seq(1,3))[2,] / 
                    max(getIsotope(getMolecule(as.character(
                      formulas.ok$formula[j])), seq(1,3))[2,]))*100)[2] - 
                  (intensities / max(intensities)*100)[2]))
            } # close formula "j"
            formulas.ok <- formulas.ok[order(formulas.ok$isodev), ]
            
            if(nrow(formulas.ok) < 6
              ){
              std_info[
                which(std_info$name == std_info.i$name[i]),
                paste0("form_pred_", polarity)] <- 
                paste(formulas.ok$formula, collapse = "; ") 
            } else{
              std_info[
                which(std_info$name == std_info.i$name[i]),
                paste0("form_pred_", polarity)] <- 
                paste0(paste(formulas.ok$formula[1:5], 
                             collapse = "; "), "; etc. (n=", 
                       nrow(formulas.ok), ")")
            }
          } # close: if(nrow(formulas.ok) > 0)
          rm(pks)
        } # close if any pks
      } # close: if(!is.na(std_info.i$mz[i]))
    } # close std "i"
  } # close polarity "z"
} # close mix "k"
```


# Table

```{r table, warning=FALSE}
kable(std_info[,c(1:3,5:8,13,14,17,18)]) %>%
  kable_styling(bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive"))
```


# Session information

```{r session}
Sys.time()-startpoint

devtools::session_info()
```
