---
title: "Formula finder"
author: "Mar Garcia-Aloy, Vinicius Veri"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false
    toc_float: true
---

```{r startpoint, include = FALSE}
startpoint <- Sys.time()
```

In this code we are going to test the efficiency of the designed code 
to deduce the molecular formula of specific ions.  
We'll do that using the data from the experiment done by Vinicius Veri on January-2020. 

# Experiment description
Three replicates of the following samples were injected in the same 
order as they are named in the list below:  

- **`LowIS_Mix`**: Mix of standards prepared in Water/ACN (50:50) at 5ppm.     
- **`HighIS_Mix`**: Mix of standards prepared in Water/ACN (50:50) at 50ppm.  
- **`QC`**: QC samples (serum) without STD.  
- **`QC_LowIS_Mix`**: QC samples (serum) spicked with LowIS_Mix solution.  
- **`QC_HighIS_Mix`**: QC samples (serum) spicked with HighIS_Mix solution.  


# Preliminaries

## Parameters

```{r parameters}
MZML_PATH <- "Y:/mzML/"
#MZML_PATH <- "C:/Users/vveri/Documents/stds_ID_files_trial/"

polarity.all <- c("POS", "NEG") 
```

## Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(xcms)
library(Rdisop)
library(CompoundDb)
library(doParallel)
library(magrittr)
library(SummarizedExperiment)
library(CHNOSZ)
library(kableExtra)

extractSpectraData <- function(x) {
  if (is(x, "list")) {
    df <- DataFrame(do.call(rbind, lapply(x, MSnbase:::.spectrum_header)))
    df$mz <- NumericList(lapply(x, function(z) z@mz))
    df$intensity <- NumericList(lapply(x, function(z) z@intensity))
  } else if (inherits(x, "MSnExp")) {
    df <- DataFrame(MSnbase::fData(x))
    df$mz <- NumericList(MSnbase::mz(x))
    df$intensity <- NumericList(MSnbase::intensity(x))
  } else stop("'x' should be either a 'list' of 'Spectrum' objects or an ",
              "object extending 'MSnExp'")
  df
}

C_rule <- function(x){
  C <- round(c((x/1.1) - (x/1.1)/10, (x/1.1) + (x/1.1)/10))
  return(C)
}
H_rule <- function(C, N){
  H <- 2*C + N + 2
  return(H)
}
RPU_rule <- function(C=0, H=0, N=0, P=0, Si=0, Cl=0, Fl=0, I=0){
  RPU <- (C+Si) - (H+Cl+Fl+I)/2 + (N+P)/2 + 1
  return(RPU)
}
```


## Parallel processing

```{r}
ncores <- detectCores()-1
registerDoParallel(ncores)
register(DoparParam(), default = TRUE)
```



# Data import

## Theoretical info

```{r}
std_info <- read.table(
  "../data/standards_dilution.txt",
  sep = "\t", header = TRUE, as.is = TRUE)

std_info$mz_POS <- NA
std_info$mz_NEG <- NA
for(i in 1:nrow(std_info)){
  std_info$mz_POS[i] <- unlist(
    mass2mz(getMolecule(as.character(std_info$formula[i]))$exactmass, 
            adduct = as.character(std_info$POS[i])))
  std_info$mz_NEG[i] <- unlist(
    mass2mz(getMolecule(as.character(std_info$formula[i]))$exactmass, 
            adduct = as.character(std_info$NEG[i])))
}
std_info$form_pred_POS <- NA
std_info$form_pred_NEG <- NA

mix.levels <- levels(factor(std_info$mix))
```


## Experimental info

```{r}
# Import table with file names:
injections <- read.table("../data/std_serum_files.txt", 
                         sep = "\t", header = TRUE, as.is = TRUE)

# Select files adquired in MS1:
injections <- injections[injections$mode == "FS", ]

# Focus only on the 2nd replicated of "HighIS_Mix" samples:
injections <- injections[grep("Water_Mix", injections$class), ]
injections <- injections[grep("_High", injections$class), ]
injections <- injections[grep("_2_", injections$mzML), ]
```


# Output

```{r, warning=FALSE}
cwp <- CentWaveParam(
  ppm = 50,
  peakwidth = c(2, 20),
  snthresh = 10,
  mzdiff = 0.001,
  prefilter = c(3, 500),
  noise = 100,
  integrate = 2)

for(k in 1#:(length(mix.levels))
){
  
  # Select STDs used in mix "k"
  std_info.i <- std_info[std_info$mix == mix.levels[k], ]
  if(mix.levels[k] == 5){
    std_info.i <- std_info.i[std_info.i$name != 
                               "Glycerophospho-inositol", ]
  } else if(mix.levels[k] == 9){
    std_info.i <- std_info.i[std_info.i$name != 
                               "Eplerone", ]
  } else if(mix.levels[k] == 17){
    std_info.i <- std_info.i[std_info.i$name != 
                               "3-Hydroxy-DL-kynurenine", ]
  }
  
  for(z in seq(length(polarity.all))){
    if(polarity.all[z] == "NEG"){ 
      pol <- -1
      ion <- -1.007276 
    } else if(polarity.all[z] == "POS"){ 
      pol <- 1
      ion <- 1.007276 
    }
    idx <- injections$type == paste0(
      "Mix",
      sprintf(paste0("%0", ceiling(log10(10 + 1L)), "d"), 1:20)[k])
    filename <- paste0(injections$folder[idx], "/", injections$mzML[idx])
    #filename <- injections$mzML[idx]
    filename <- filename[grep(polarity.all[z], filename)]
    data_raw <- readMSData(paste0(MZML_PATH, filename),
                           mode = "onDisk")
    std_info.i$mz <- std_info.i[, which(colnames(std_info.i) == 
                                          paste0("mz_",polarity.all[z]))]
    std_info.i <- std_info.i[!is.na(std_info.i$mz),]
    for(i in 1:(nrow(std_info.i))){
      eic <- chromatogram(data_raw, aggregationFun = "max", 
                          mz = std_info.i$mz[i] + 0.01 * c(-1, 1))
      pks <- data.frame(chromPeaks(findChromPeaks(eic, param = cwp)))
      if(any(which((pks$rtmin - 10) < std_info.i$RT[i] & 
                   (pks$rtmax + 10) > std_info.i$RT[i]) ==  
             which.max(pks$maxo))){
        if(which((pks$rtmin - 10) < std_info.i$RT[i] & 
                 (pks$rtmax + 10) > std_info.i$RT[i]) ==  
           which.max(pks$maxo)){
          pks <- pks[which((pks$rtmin - 10) < std_info.i$RT[i] & 
                             (pks$rtmax + 10) > std_info.i$RT[i]), ]
          if(nrow(pks) > 1){
            pks <- pks[which.max(pks$maxo), ]
          }
          sps <- data_raw %>%
            filterRt(rt = pks$rt + 2.5 * c(-1, 1)) %>%
            extractSpectraData
          sps2 <- Spectra::Spectra(sps)
          sps2 <- Spectra::combineSpectra(sps2, 
                                          peaks = "intersect", 
                                          minProp = 0.75,  
                                          tolerance = 0.001, ppm = 10)
          # "minProp" allows to define which peaks to keep 
          # (if peaks = "intersect"), keeps only peaks present in 
          # >= minProp of input spectra
          
          # "tolerance" and "ppm" allow to define how peaks from 
          # different spectra are matched. They are considered to be the 
          # same peak if the difference in their m/z is smaller 
          # tolerance + ppm(mz)
          
          # Note that by default the mean m/z and intensity are reported. 
          # To get the maximum intensity use "intensityFun = max".
          # The trick is actually the peaks = "intersect" because the 
          # default peaks = "union" keeps all peaks from all spectra.
          
          sps2 <- data.frame(mz = sps2$mz[[1]],
                             i = sps2$intensity[[1]])
          sps2$i100 <- (sps2$i / max(sps2$i))*100
          
          sps2 <- sps2[order(-sps2$i), ]
          idx <- c(which(round(sps2$mz) == round(std_info.i$mz[i]))[1],
                   which(round(sps2$mz) == round(std_info.i$mz[i]+1))[1],
                   which(round(sps2$mz) == round(std_info.i$mz[i]+2))[1])
          sps2 <- sps2[idx, ]
          masses <- sps2$mz
          intensities <- sps2$i
          molecules <- decomposeIsotopes(masses, intensities, ppm = 10)
          formulas <- data.frame(formula = getFormula(molecules))
          tmp <- makeup(as.character(formulas$formula))
          formulas$C <- NA
          formulas$H <- NA
          formulas$O <- NA
          formulas$N <- NA
          formulas$S <- NA
          formulas$P <- NA
          for(j in 1:nrow(formulas)){
            if(length(tmp[[j]][names(tmp[[j]]) == "C"]) == 1){
              formulas$C[j] <- tmp[[j]][names(tmp[[j]]) == "C"]  
            } else {formulas$C[j] <- 0}
            if(length(tmp[[j]][names(tmp[[j]]) == "H"]) == 1){
              formulas$H[j] <- tmp[[j]][names(tmp[[j]]) == "H"]
            } else {formulas$H[j] <- 0}
            if(length(tmp[[j]][names(tmp[[j]]) == "O"]) == 1){
              formulas$O[j] <- tmp[[j]][names(tmp[[j]]) == "O"]
            } else {formulas$O[j] <- 0}
            if(length(tmp[[j]][names(tmp[[j]]) == "N"]) == 1){
              formulas$N[j] <- tmp[[j]][names(tmp[[j]]) == "N"]
            } else {formulas$N[j] <- 0}
            if(length(tmp[[j]][names(tmp[[j]]) == "S"]) == 1){
              formulas$S[j] <- tmp[[j]][names(tmp[[j]]) == "S"]
            } else {formulas$S[j] <- 0}
            if(length(tmp[[j]][names(tmp[[j]]) == "P"]) == 1){
              formulas$P[j] <- tmp[[j]][names(tmp[[j]]) == "P"]
            } else {formulas$P[j] <- 0}
          } # close formula "j"
          
          formulas$C_rule <- FALSE
          tmp <- C_rule((intensities[2] / intensities[1])*100)
          formulas$C_rule[formulas$C >= tmp[1] & formulas$C <= tmp[2]] <- TRUE
          formulas$H_rule <- FALSE
          formulas$H_rule[formulas$H <= H_rule(formulas$C, formulas$N)] <- TRUE
          formulas$N_rule <- formulas$N %% 2 == round(masses[1] - ion) %% 2
          formulas$RPU_rule <- FALSE
          formulas$RPU_rule[RPU_rule(C = formulas$C, H = formulas$H - pol, 
                                     N = formulas$N, P = formulas$P) >= 0] <- TRUE
          formulas$RPU_rule[RPU_rule(C = formulas$C, H = formulas$H - pol, 
                                     N = formulas$N, P = formulas$P) %% 1 > 0] <- FALSE
          
          formulas.ok <- formulas[formulas$C_rule & formulas$H_rule & 
                                    formulas$N_rule & formulas$RPU_rule, ]
          if(nrow(formulas.ok) > 0){
            formulas.ok$isodev <- NA
            for(j in 1:nrow(formulas.ok)){
              formulas.ok$isodev[j] <- mean(abs(
                (getIsotope(getMolecule(as.character(formulas.ok$formula[j])), 
                            seq(1,3))[2,] / 
                   max(getIsotope(getMolecule(as.character(formulas.ok$formula[j])), 
                                  seq(1,3))[2,]))*100 - 
                  (intensities / max(intensities)*100)))
            } # close formula "j"
            formulas.ok <- formulas.ok[order(formulas.ok$isodev), ]
            
            std_info[
              which(std_info$name == std_info.i$name[i]),
              which(colnames(std_info) == 
                      paste0("form_pred_", polarity.all[z]))] <- 
              paste(formulas.ok$formula, collapse = "; ") 
          } # close: if(nrow(formulas.ok) > 0)
        } else {
          print(paste0("Mix-",k,", ", std_info.i$name[i], 
                       ": Check which chromatographic peak is used!!" ))
        }
        rm(pks)
      } # close if any pks
    } # close std "i"
  } # close polarity "z"
} # close mix "k"
```


# Table

```{r, warning=FALSE}
kable(std_info[,c(1:3,5:8,13,14,17,18)]) %>%
  kable_styling(bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive"))
```


# Session information

```{r session}
Sys.time()-startpoint

devtools::session_info()
```
