---
title: "Standards: Matrix effect in RT"
author: "Mar Garcia-Aloy"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false
    toc_float: true
---

In this document I'm going to check the matrix effect in retention times 
(RT) of (internal) standards (IS / STD).  
The data used is from the experiment done by Vinicius Veri 
on November-2019 for ISs and on January-2020 for STDs (MIXs). 
For IS, sample names are stored in the file `data/is_serum_files.txt` and experiment 
description in the file `2019_11_18_IS_Experiment Summary.docx`; 
whereas for STDs, sample names are in the file `data/std_serum_files.txt` 
and experiment description in the file `2019_01_13_Experiment Plan_All Mixes.xlsx`.    

# Experiment description
Three replicates of the following samples were injected in the same 
order as they are named in the list below:  

- **`Low_IS`**: IS solution used for the untargeted protocol.     
- **`High_IS`**: IS solution x10 more concentrated.  
- **`QC`**: QC samples (serum) without IS.  
- **`QC_Low_IS`**: QC samples (serum) with IS concentrations of untargeted protocol.  
- **`QC_High_IS`**: QC samples (serum) with IS x10 more concentrated.  

This scheme is for the IS. The same was done for each of the MIXs.  

# Preliminaries
## Parameters
```{r}
polarity.all <- c("POS", "NEG") 
da <- 0.01
rt.d <- 20
i.thr <- 10
MZML_PATH <- "/data/massspec/mzML/" 
#MZML_PATH <- "C:/Users/mgarciaaloy/Documents/mzML/"
dir.create("images/", 
           recursive = TRUE, showWarnings = FALSE)
startpoint <- Sys.time()
```

## Libraries
```{r, message=FALSE, warning=FALSE}
library(doParallel)
library(Rdisop)
library(CompoundDb)
library(RColorBrewer)
library(kableExtra)
library(xcms)
cwp1 <- CentWaveParam(
  peakwidth = c(2, 20), 
  ppm = 50, 
  snthresh = 5, ###############
  mzdiff = 0.001,
  prefilter = c(3, 1000),
  noise = 100,
  integrate = 2)
cwp2 <- CentWaveParam(
  peakwidth = c(2, 20), 
  ppm = 50, 
  snthresh = 1, ###############
  mzdiff = 0.001,
  prefilter = c(3, 1000),
  noise = 100,
  integrate = 2)
mnp <- MergeNeighboringPeaksParam(
  expandRt = 3, 
  expandMz = 0.001, 
  ppm = 10,
  minProp = 0.66)

which_within <- function(x, y, mzd = 0.01){
  lapply(x, function(z){
    which(y >= z - mzd & y <= z + mzd)
  }
  )
}
```

## Parallel processing
```{r, eval=TRUE}
ncores <- Sys.getenv("SLURM_JOB_CPUS_PER_NODE", 3)
register(bpstart(MulticoreParam(ncores)))
```

## Annotations
```{r}
ih_cmp <- read.table(
  "../data/inhouse_cmps.txt",
  sep = "\t", header = TRUE, as.is = TRUE)

ih_ion <- read.table(
  "../data/inhouse_ions.txt", 
  sep = "\t", header = TRUE, as.is = TRUE)

ih_feat <- read.table(
  "../data/inhouse_feat.txt",
  sep = "\t", header = TRUE, as.is = TRUE)

ih_ip <- read.table(
  "../data/is_isotopic_pattern.txt",
  sep = "\t", header = TRUE, as.is = TRUE)

ih_feat$rt <- NA
for(i in seq(nrow(ih_feat))){
  if(any(which(ih_cmp$cmp_ID == ih_feat$ID[i]))){
    ih_feat$rt[i] <- ih_cmp$RT[which(ih_cmp$cmp_ID == ih_feat$ID[i])]
  } else {
    if(any(which(ih_cmp$comments == ih_feat$ID[i]))){
      ih_feat$rt[i] <- ih_cmp$RT[which(ih_cmp$comments == ih_feat$ID[i])]
    }
  }
}
ih_feat <- subset(ih_feat, 
                  select = c("ID", "rt", "mode", "mz", "assignation" ))

colnames(ih_ip) <- c("ID", "mode", "mz", "assignation", "abundance")
ih_ip$rt <- NA
for(i in seq(nrow(ih_ip))){
  if(any(which(ih_cmp$cmp_ID == ih_ip$ID[i]))){
    ih_ip$rt[i] <- ih_cmp$RT[which(ih_cmp$cmp_ID == ih_ip$ID[i])]
  } else {
    if(any(which(ih_cmp$comments == ih_ip$ID[i]))){
      ih_ip$rt[i] <- ih_cmp$RT[which(ih_cmp$comments == ih_ip$ID[i])]
    }
  }
}
ih_ip <- subset(ih_ip, 
                select = c("ID", "rt", "mode", "mz", "assignation" ))

# Calculate the mz values of the molecular ion:
ions.long.list <- list()
for(k in seq(length(polarity.all))){
  polarity <- polarity.all[k]
  ih_cmp$mzneut = NA
  ih_cmp$mz <- NA
  if(polarity == "NEG"){ ion = -1.007276 
  } else if(polarity == "POS"){ ion = 1.007276 }
  for(i in seq(nrow(ih_cmp))){
    
    if(grepl("C", ih_cmp$formula[i])){
      ih_cmp$mzneut[i] = 
        getMolecule(as.character(ih_cmp$formula[i]))$exactmass
    } else {
      ih_cmp$mzneut[i] = as.numeric(ih_cmp$formula[i])
    }
    
    ih_cmp$mz[i] <- ih_cmp$mzneut[i] + ion
  }
  
  # Get all ions associated with annotated compounds:
  ions.long = data.frame(matrix(ncol=5,nrow=0))
  colnames(ions.long) = c("ID", "rt", "mode", "mz", "assignation")
  
  ions.loop <- data.frame(matrix(ncol=5, nrow=nrow(ih_cmp)))
  colnames(ions.loop) <- c("ID", "rt", "mode", "mz", "assignation")
  
  ions.loop$ID <- ih_cmp$cmp_ID
  ions.loop$rt <- ih_cmp$RT
  ions.loop$mode <- polarity
  ions.loop$mz <- ih_cmp$mz
  if(polarity == "NEG"){
    ions.loop$assignation <- "[M-H]-"
  }else if(polarity == "POS"){
    ions.loop$assignation = "[M+H]+"}
  ions.long = rbind(ions.long, ions.loop)
  
  for(i in 4:ncol(ih_ion)){
    for(j in seq(nrow(ih_ion))){
      if(ih_ion[j,i]!=0){
        ions.loop = data.frame(matrix(ncol=5,nrow=1))
        colnames(ions.loop) = c("ID", "rt", "mode", "mz", "assignation")
        ions.loop$ID = colnames(ih_ion)[i]
        ions.loop$rt = 
          ih_cmp$RT[which(ih_cmp$cmp_ID == colnames(ih_ion)[i])]
        ions.loop$mode <- polarity
        
        if(ih_ion$type[j] == "dimer"){
          ions.loop$mz = (ih_cmp[ih_cmp$cmp_ID == colnames(ih_ion)[i], 
                                 "mzneut"])*2 + ion + ih_ion$delta[j]
          # close statment "if ion "j" is a dimer"
        } else if(ih_ion$type[j] == "trimer"){
          ions.loop$mz = (ih_cmp[ih_cmp$cmp_ID == colnames(ih_ion)[i], 
                                 "mzneut"])*3 + ion + ih_ion$delta[j]
          # close statment "if ion "j" is a trimer"
        } else if(ih_ion$type[j] == "charged"){
          ions.loop$mz = (ih_cmp[ih_cmp$cmp_ID == colnames(ih_ion)[i], 
                                 "mz"] + ion)/2
          ions.loop$assignation <- ih_ion$ion_id[j]
          # close statment "if ion `j` is a charged"
        }else {
          ions.loop$mz = (ih_cmp[ih_cmp$cmp_ID == colnames(ih_ion)[i], 
                                 "mz"]) + ih_ion$delta[j]
        } # close statment "if ion "j" is NOT  dimer / trimer / charged"
        ions.loop$assignation <- ih_ion$ion_id[j]
        ions.long = rbind(ions.long, ions.loop)
      } # close statment "if ion "j" in cmp "i" is detected"...
    } # close loop ion "j"
  } # close loop in-house compound "i"
  
  if(polarity == "POS"){
    ions.long$assignation = gsub("\\*", "+", ions.long$assignation)
  } else if(polarity == "NEG"){
    ions.long$assignation = gsub("\\*", "-", ions.long$assignation)
  }
  
  ions.long <- ions.long[ions.long$mode == polarity, ]
  ions.long.list[[k]] <- ions.long
}
ions.long <- do.call("rbind", ions.long.list)
ions.long <- rbind(ions.long, ih_feat)
ions.long <- rbind(ions.long, ih_ip)
for(i in seq(nrow(ih_cmp))){
  if(ih_cmp$identified[i] == "YES"){
    ions.long$cmp_ID[ions.long$ID == ih_cmp$cmp_ID[i]] <- 
      ih_cmp$comments[i]
  }
}
rm(ions.long.list, ih_cmp, ih_feat, ih_ion, i, j, k, polarity, ions.loop, ion)
```


# Data import
## Standards
```{r}
is_info <- read.table(
  #"https://raw.githubusercontent.com/EuracBiomedicalResearch/lcms-standards/master/data/internal_standards.txt",
  "../data/internal_standards.txt",
  sep = "\t", header = TRUE, as.is = TRUE)
is_info$mix <- 0
is_info$"HMDB.code" <- NA
is_info <- 
  subset(is_info, 
         select = c("mix", "name", "abbreviation", "HMDB.code", 
                    "formula", "POS", "NEG", "RT", "data_set", "sample", 
                    "operator", "version"))
std_info <- read.table(
  #"https://raw.githubusercontent.com/EuracBiomedicalResearch/lcms-standards/master/data/standards_dilution.txt",
  "../data/standards_dilution.txt",
  sep = "\t", header = TRUE, as.is = TRUE)

std_info <- rbind(is_info, std_info)
rm(is_info)
std_info$POS[is.na(std_info$POS)] <- "[M+H]+"
std_info$NEG[is.na(std_info$NEG)] <- "[M-H]-"
std_info <- std_info[order(std_info$mix, std_info$RT),]

mix.levels <- levels(factor(std_info$mix))
```


# Output
```{r}
for(k in 1:(length(mix.levels))
){
  if(mix.levels[k] == 0){ # start import table "injections"
    injections <- read.table("../data/is_serum_files.txt", 
                             sep = "\t", header = TRUE, as.is = TRUE)
    injections <- injections[-grep("Blank", injections$mzML),]
  } else {
    injections <- read.table("../data/std_serum_files.txt", 
                             sep = "\t", header = TRUE, as.is = TRUE)
    injections <- injections[injections$mode == "FS", ]
    inj_blankQC <- injections[injections$type == "Blank_QC", ]
    inj_blankQC <- inj_blankQC[grep("QC_5", inj_blankQC$mzML), ]
    inj_blankQC$class <- "QC"
    injections <- injections[grep("Mix", injections$mzML), ]
    injections <- injections[
      as.numeric(gsub("Mix", "", injections$type)) == mix.levels[k], ]
    injections <- rbind(inj_blankQC, injections)
  } # close import table "injections"
  
  
  # Select STDs used in mix "k"
  std_info.i <- std_info[std_info$mix == mix.levels[k], ]
  if(mix.levels[k] == 5){
    std_info.i <- std_info.i[std_info.i$name != "Glycerophospho-inositol", ]
  } else if(mix.levels[k] == 9){
    std_info.i <- std_info.i[std_info.i$name != "Eplerone", ]
  } else if(mix.levels[k] == 17){
    std_info.i <- std_info.i[std_info.i$name != "3-Hydroxy-DL-kynurenine", ]
  }
  
  
  # Colouring factors
  if(mix.levels[k] == 0){
    injections$class <- 
      factor(injections$class, 
             levels = c("LowIS", "HighIS", "QC", "QC_LowIS", "QC_HighIS"))
  } else {
    injections$class <- gsub(paste0(injections$type[10], "_"), "", 
                             injections$class)
    injections$class <- factor(injections$class, 
                               levels = c("Water_Low", "Water_High", 
                                          "QC", "QC_Low", "QC_High"))
  }
  injections <- injections[order(injections$class),]
  
  col_class <- brewer.pal(6, "Paired")
  col_class <- col_class[-4]
  names(col_class) <- levels(injections$class)
  
  
  for(z in seq(length(polarity.all))
  ){
    polarity <- polarity.all[z]
    injections.j <- injections[injections$polarity == polarity, ]
    myfiles <- paste(injections.j$folder, injections.j$mzML, sep = "/") 
    
    # Read the mzML files
    data_raw <- readMSData(paste0(MZML_PATH, myfiles), 
                           pdata = new("NAnnotatedDataFrame", injections.j),
                           mode = "onDisk")
    
    for(i in 1:(nrow(std_info.i))){
      
      # Calculate the mz for the main ion (mzvalue):
      if(grepl("C", std_info.i$formula[i])){
        mzvalue <- getMolecule(as.character(std_info.i$formula[i]))$exactmass
        mzvalue <- unlist(mass2mz(
          as.numeric(mzvalue), 
          adduct = as.character(
            std_info.i[i, which(colnames(std_info.i) == polarity)]
          )))
      } else{
        mzvalue <- unlist(mass2mz(
          as.numeric(std_info.i$formula[i]), 
          adduct = as.character(
            std_info.i[i, which(colnames(std_info.i) == polarity)]
          )))
      } # close calculate mzvalue
      
      
      cmpname <- std_info.i$abbreviation[i]
      filename <- paste0("images/", mix.levels[k], "_", cmpname, "_", 
                         polarity, "_EIC.png")
      
      png(file = filename, width = 1000, height = 1000)
      par(mfrow = c(2, 2))
      
      # EIC full RT range -------------------------------------------------
      chr <- chromatogram(data_raw, 
                          mz = c(mzvalue - da,  mzvalue + da),
                          #rt = c(std_info.i$RT[i] + 5 * c(-1,1)),
                          aggregationFun = "max")
      plot(chr, main = std_info.i$name[i],
           col = col_class[chr$class],
           bg = "white")
      legend("topright", legend = names(col_class), 
             pch = 16, col = gsub("60", "", col_class))
      abline(v = std_info.i$RT[i], lty = 2)
      
      if(!is.na(std_info.i$RT[i])){
        # EIC RT zoomed -----------------------------------------------------
        chr2 <- findChromPeaks(chr, param = cwp1)
        delta_rt <- abs(chromPeaks(chr2)[, "rt"] - std_info.i$RT[i])
        pks <- data.frame(chromPeaks(chr2)[delta_rt < 60, , drop=FALSE])
        if (nrow(pks)) {
          pks2 <- pks[0, ]
          for(j in levels(factor(pks$column))){
            pks3 <- pks[pks$column == j, ]
            pks2 <- rbind(pks2, pks3[which.max(pks3$maxo),])
          }
          #pks2 <- pks2[pks2$column < 7 | pks2$column > 9, ]
          rtmin <- min(pks2$rtmin)
          rtmax <- max(pks2$rtmax)
          ylim <- c(0, max(pks$maxo, na.rm = TRUE))
        } else {
          rtmin <- std_info.i$RT[i] - 10
          rtmax <- std_info.i$RT[i] + 10
          ylim <- c(0, max(sapply(chr, intensity), na.rm = TRUE))
        }
        
        plot(chr2, peakType = "none", 
             col = col_class[chr2$class],
             xlim = c(rtmin, rtmax))
        legend("topright", legend = names(col_class), 
               pch = 16, col = gsub("60", "", col_class))
        abline(v = std_info.i$RT[i], lty = 2)
        
        # EIC RT zoomed only for serum samples -------------------------------
        if(nrow(pks)){
          xdata.i <- filterFile(data_raw, grep("QC", myfiles))
          chr.i <- chromatogram(xdata.i, 
                                mz = c(mzvalue - da,  mzvalue + da),
                                aggregationFun = "max")
          if(mix.levels[k] == 0){
            pks3 <- pks2[(pks2$column %in% grep("18_QC_LowIS", myfiles)) |
                           (pks2$column %in% grep("18_QC_HighIS", myfiles)),]
            if(nrow(pks3) >0){
              rtmin <- min(pks3$rtmin)
              rtmax <- max(pks3$rtmax)
            }
          }else {
            pks3 <- pks2[(pks2$column %in% grep("QC_LowIS", myfiles)) |
                           (pks2$column %in% grep("QC_HighIS", myfiles)),]
          }
          plot(chr.i, 
               col = col_class[factor(chr.i$class, levels = names(col_class))],
               bg = "white", main = "QC serum",
               xlim = c(rtmin, rtmax))
          abline(v = std_info.i$RT[i], lty = 2)
          
          rm(pks3)
          
          
          # FS spectrum in HighIS_water ----------------------------------------
          if(mix.levels[k] == 0){
            pks3 <- pks2[pks2$column %in% grep("18_HighIS", myfiles),]
            idx <- pks3$column[which.max(pks3$maxo)]
            rt.i <- pks3$rt[which.max(pks3$maxo)]
          }else{ 
            pks3 <- pks2[pks2$column %in% grep("/HighIS", myfiles),]
            idx <- pks3$column[which.max(pks3$maxo)]
            rt.i <- pks3$rt[which.max(pks3$maxo)]
          }
          if((mix.levels[k] == 0 & any(pks2$column %in% grep("18_HighIS", myfiles))) |
             (mix.levels[k] > 0 & any(pks2$column %in% grep("/HighIS", myfiles)))
          ){
            xdata.i <- filterFile(data_raw, idx)
            sps <- xdata.i %>%
              filterRt(rt = c(round(rt.i) - 0.5, 
                              round(rt.i) + 0.5)) %>%
              spectra
            sps2 <- data.frame(sps[[2]])
            sps2$i100 <- (sps2$i / max(sps2$i))*100
            
            il <- ions.long[ions.long$rt > (sps[[2]]@rt - rt.d) & 
                              ions.long$rt < (sps[[2]]@rt + rt.d),]
            il$cmp_ID[is.na(il$cmp_ID)] <- il$ID[is.na(il$cmp_ID)]
            il <- il[il$mode == my.polarity, ]
            
            sps2$cmp <- NA
            sps2$annotation <- NA
            for(j in seq(nrow(sps2))){
              mz.i <- sps2$mz[j]
              id.i <- il[unlist(which_within(mz.i, il$mz, mzd = 0.015)), ]
              if(nrow(id.i) == 1){
                sps2$cmp[j] <- id.i$cmp_ID
                sps2$annotation[j] <- id.i$assignation
              } else if(nrow(id.i) > 1){
                id.i <- id.i[(abs(sps[[2]]@rt - id.i$rt) < 10),]
                if(nrow(id.i) > 1){
                  id.i <- id.i[abs(mz.i - id.i$mz) < 0.01,]
                } # close "if(nrow(id.i) > 1)"
                if(nrow(id.i) == 1){
                  sps2$cmp[j] <- id.i$cmp_ID
                  sps2$annotation[j] <- id.i$assignation
                } else if(nrow(id.i) == 2){
                  sps2$cmp[j] <- paste(id.i$cmp_ID, collapse=" / ")
                  if(id.i$assignation[1] == id.i$assignation[2]){
                    sps2$annotation[j] <- id.i$assignation[1]
                  }
                }
                #print(paste(j, "There is >1 matching!!"))
              }
            }
            sps2$cmp[is.na(sps2$cmp)] <- "Unknown"
            sps2$cmp <- factor(sps2$cmp)
            
            plot(sps2$mz, sps2$i100, type = "h", 
                 xlab = "m/z", ylab = "relative intensity", 
                 main = paste(gsub(".mzML", "", xdata.i@phenoData@data$mzML), 
                              "\n RT:", round(sps[[2]]@rt)))
            sps3 <- sps2[sps2$i100 > i.thr, ]
            sps3$cmp <- droplevels(sps3$cmp)
            cmps <- levels(sps3$cmp)
            for(j in seq(length(cmps))){
              lines(sps3$mz[sps3$cmp == cmps[j]], 
                    sps3$i100[sps3$cmp == cmps[j]], 
                    type = "h", col = j + 1)
              text(sps3$mz[sps3$cmp == cmps[j]], 
                   sps3$i100[sps3$cmp == cmps[j]],
                   paste(round(sps3$mz[sps3$cmp == cmps[j]], 4), "\n", 
                         sps3$annotation[sps3$cmp == cmps[j]]), 
                   col = j + 1#, cex = 0.5
              )
            }
            legend("topright", cmps, 
                   col = seq(length(cmps)) + 1, pch = 16)
          }
        } # close if(nrow(pks)) (II)
      } # close if(!is.na(std_info$RT))
      
      dev.off()
      
      ####################################################################
      
      # Plot 1 chr / sample ----------------------------------------------
      
      chr <- chromatogram(data_raw, # EIC for the mzvalue
                          aggregationFun = "max",
                          mz = mzvalue + 0.01 * c(-1, 1),
                          rt = std_info.i$RT[i] + 30 * c(-1, 1))
      chr2 <- findChromPeaks(chr, param = cwp2) # Peak detection
      chr3 <- refineChromPeaks(chr2, param = mnp) # Peak refinement
      pks <- data.frame(chromPeaks(chr3)) # Extract info detected peaks
      
      if(nrow(pks)){
        # Indicate filenames in "pks" matrix
        if(mix.levels[k] == 0){
          pks$file <- gsub(".mzML", "", 
                           gsub("2019/2019_11/20191118_", "", 
                                myfiles[pks$column]))
        } else{
          pks$file <- gsub(".mzML", "", 
                           gsub("2020/2020_01/", "", 
                                myfiles[pks$column]))
        }
        
        pks <- pks[order(pks$column, pks$rt), ]
        
        # Select only 1 peak / sample:
        pks2 <- pks[0, ]
        for(l in levels(factor(pks$column))){
          pks3 <- pks[pks$column == l, ]
          pks2 <- rbind(pks2, pks3[which.max(pks3$maxo),])
        } # close sample "l"
        rm(pks3)
        
        filename <- paste0("images/", mix.levels[k], "_", cmpname, "_", 
                           polarity, "_peaks.png")
        
        png(file = filename, width = 1000, height = 1000)
        par(mfrow = c(5, 3), mar = c(2, 2, 2, 0))
        for(y in seq(length(chr3))){
          if(mix.levels[k] == 0){
            mytitle <- gsub(".mzML", "", gsub("2019/2019_11/20191118_", "", 
                                              myfiles[y]))
          } else{
            mytitle <- gsub(".mzML", "", gsub("2020/2020_01/", "", 
                                              myfiles[y]))
          }
          
          plot(chr3[[y]], 
               main = mytitle)
          if(any(length(pks$rt[pks$column == y]))){
            for(x in seq(length(pks$rt[pks$column == y]))){
              text(pks$rt[pks$column == y][x],
                   pks$maxo[pks$column == y][x],
                   x)
            }
            points(pks2$rt[pks2$column == y],
                   pks2$maxo[pks2$column == y],
                   pch = 8, col = "red")
          }
        }
        dev.off()
      } # close "if(nrow(pks))"
      
      rm(chr, chr2, chr3)
    } # close std "i"
  } # close polarity "z"
} # close mix.levels "k"
```

# Session information
```{r}
Sys.time()-startpoint

devtools::session_info()
```
