---
title: "Internal Standards: Matrix effect - XCMS processing"
author: "Mar Garcia-Aloy"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false
    toc_float: true
---


```{r include = FALSE}
knitr::opts_chunk$set(echo=TRUE)

startpoint <- Sys.time()
```


This document belongs to the project designed to check the matrix effect 
in retention times (RT) of internal standards (IS). After checking this 
information ([output here](RT_matrix_effect.Rmd)), the current aim is to 
check the feature extraction of each IS.     
I'll start using only samples that consisted with 2 different 
concentrations of IS diluted in water (i.e., `Low_IS` and `High_IS`). 
Later, I'll go with QC-serum samples. The group of samples that are going 
to be used is specified in the parameter "study" (in section 
"preliminaries" - "parameters").  
In this document I'm going to do the XCMS processing. This output will be 
taken for the upcoming document in order to plot each feature 
(from each IS) separately and have the enough information to check the 
processed data.  
  
# Preliminaries
I define the specific parameters for the current study, I load all 
required libraries, and I set up the parallel processing.

## Parameters
```{r, echo=TRUE}
polarity <- "POS" # specify "POS" or "NEG"
study <- "IS"     # specify "IS" or "QC"
rt_cut <- 340
MZML_PATH <- "C:/Users/mgarciaaloy/Documents/mzML/"
```

## Libraries
```{r, message=FALSE, warning=FALSE}
library(doParallel)
library(xcms)
library(RColorBrewer)
```

## Parallel processing
```{r}
ncores <- detectCores()-1
registerDoParallel(ncores)
register(DoparParam(), default = TRUE)
```


# Study samples
Below I'm going to import the document with the file names and later I'll 
restrict them to the current polarity.  
 
```{r}
injections <- read.table(paste0(gsub("std_serum", "", getwd()), 
                                "data/is_serum_files.txt"), 
                 sep = "\t", as.is = TRUE, header = TRUE)

injections <- injections[injections$polarity == polarity, ]
if (study == "IS") {
  injections <- injections[#injections$class == "Blank_Inj" |
                             #injections$class == "Blank_Ext" |
                             injections$class == "LowIS" | 
                             injections$class == "HighIS", ]
} else if (study == "QC"){
  injections <- injections[injections$class == "QC" | 
                           injections$class == "QC_LowIS" |
                             injections$class == "QC_HighIS", ]
}

myfiles <- paste(injections$folder, injections$mzML, sep = "/") 
```

## Colouring factors
Also, I define colors to be used for the various experimental groups 
throughout the experiment.
```{r}
injections$class <- factor(injections$class)
if(length(levels(injections$class)) < 3){
  col_class <- brewer.pal(3, "Set1")
  col_class <- col_class[1:length(levels(injections$class))]}else{
  col_class <- brewer.pal(length(levels(injections$class)), "Set1")
}
names(col_class) <- levels(injections$class)
plot(NULL ,xaxt='n',yaxt='n',bty='n',ylab='',xlab='', xlim=0:1, ylim=0:1)
legend("topleft", legend = names(col_class),
       col = col_class, pch = 16, pt.cex=3, cex=1.5, bty='n')
mtext("Class", at=0.2, cex=2)
```


# Workflow
## readMSData
```{r}
raw_data <- readMSData(files = paste0(MZML_PATH, myfiles), 
                       pdata = new("NAnnotatedDataFrame", injections),
                       mode = "onDisk")
raw_data <- filterRt(raw_data, rt = c(0, rt_cut))
```

## Peak detection & post-processing
```{r}
cwp <- CentWaveParam(
  peakwidth = c(2, 20), 
  ppm = 50, 
  snthresh = 100,
  mzdiff = 0.001,
  prefilter = c(4, 5000),
  noise = 100,
  integrate = 2)

mnp <- MergeNeighboringPeaksParam(
  expandRt = 8, 
  expandMz = 0.001, 
  ppm = 10,
  minProp = 0.75)

xdata <- findChromPeaks(raw_data, param = cwp)

xdata <- refineChromPeaks(xdata,
                          param = mnp, 
                          msLevel = 1L,
                          BPPARAM = bpparam())
```

## Aligment and Correspondance
```{r}
pdp1 <- PeakDensityParam(
  sampleGroups = xdata$class, 
  minFraction = 0.9, 
  bw = 3,
  binSize = 0.01)

pdp2 <- PeakDensityParam(
  sampleGroups = xdata$class, 
  minFraction = 0.9, 
  bw = 20,
  binSize = 0.01)

if (study == "IS") {
  pgp <- PeakGroupsParam(
    span = 0.3,
    subset = which(xdata$class == "LowIS"),  
    subsetAdjust = "average",
    minFraction = 0.9 
    ) 
} else if (study == "QC"){
  pgp <- PeakGroupsParam(
    span = 0.3,
    subset = which(xdata$class == "QC_LowIS"),  
    subsetAdjust = "average",
    minFraction = 0.9 
    ) 
}

xdata <- groupChromPeaks(xdata, param = pdp1)
xdata <- adjustRtime(xdata, param = pgp)
plotAdjustedRtime(xdata,  col = col_class[xdata$class],
                  peakGroupsCol = "grey", peakGroupsPch = 1)
legend("bottomright",
       legend=names(col_class), col = col_class,
       lty = 1, ncol = 1)
xdata <- groupChromPeaks(xdata, param = pdp2)
```

## Peak filling
```{r}
fcp <- FillChromPeaksParam(
  ppm = 20, 
  fixedRt = median(chromPeaks(xdata)[, "rtmax"] -
                     chromPeaks(xdata)[, "rtmin"]) / 2)

xdata <- fillChromPeaks(xdata, param = fcp)
```

## Data export
```{r}
ls()
save(list = c("raw_data", "xdata", 
              "cwp", "mnp", "pdp1", "pdp2", "pgp", "fcp",
              "MZML_PATH", "myfiles"), 
     file = paste0("XCMS_processing_", study, "_", polarity, ".Rdata"))
```

# Summary
Following I mention the changes in XCMS parameters that I've applied for 
this dataset:  
  
- Peak detection (for most IS there are several "ghost peaks"):  
  - snthresh = 100  
  - prefilter = (4, 5000)  
  
- Peak post-processing:  
  - expandRt = 8 (splitted peaks: Carnitine [4] & Creatinine [8])
     
- Peak grouping (correspondance):   
  - minFraction = 0.9 (sample groups are made up 
  of replicates from the same sample)  
  - bw (II) = 20 (for some IS I got 2 features [10, except for Creatinine {20})  
  - binSize = 0.01 (for Ala sample #1 we got "ghost peaks" interfering in the final intensity value)  
  
# Session information
```{r}
Sys.time()-startpoint

devtools::session_info()
```
