---
title: "XCMS processing"
subtitle: "QC Serum: Matrix effect"
author: "Mar Garcia-Aloy"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false
    toc_float: true
---


```{r include = FALSE}
knitr::opts_chunk$set(echo=TRUE)

startpoint <- Sys.time()
```


Te current aim is to annotate the ions associated with the in-house 
standards. In this document I'm going to do the XCMS processing.   

# Preliminaries
I define the specific parameters for the current study, I load all 
required libraries, and I set up the parallel processing.

## Parameters
```{r, echo=TRUE}
polarity <- "POS" # specify "POS" or "NEG"
study <- "QC"     # specify "IS" or "QC"
rt_cut <- 340
MZML_PATH <- "C:/Users/mgarciaaloy/Documents/mzML/"
```

## Libraries
```{r, message=FALSE, warning=FALSE}
library(doParallel)
library(xcms)
library(RColorBrewer)
```

## Parallel processing
```{r}
ncores <- detectCores()-1
registerDoParallel(ncores)
register(DoparParam(), default = TRUE)
```


# Study samples
Below I'm going to import the document with the file names and later I'll 
restrict them to the current polarity.  

```{r}
injections <- read.table(paste0(gsub("std_serum", "", getwd()), 
                                "data/is_serum_files.txt"), 
                         sep = "\t", as.is = TRUE, header = TRUE)

injections <- injections[injections$polarity == polarity, ]

if (study == "IS") {
  injections <- injections[injections$class == "Blank_Inj" |
                             injections$class == "Blank_Ext" |
                             injections$class == "LowIS" | 
                             injections$class == "HighIS", ]
} else if (study == "QC"){
  injections <- injections[injections$class == "Blank_Inj" |
                             injections$class == "QC" | 
                             injections$class == "QC_LowIS" |
                             injections$class == "QC_HighIS", ]
}

myfiles <- paste(injections$folder, injections$mzML, sep = "/") 
```

## Colouring factors
Also, I define colors to be used for the various experimental groups 
throughout the experiment.
```{r}
injections$class <- factor(injections$class)
if(length(levels(injections$class)) == 2){
  col_class <- brewer.pal(3, "Set1")
  col_class <- col_class[-2]
  } else {
    col_class <- brewer.pal(length(levels(injections$class)), "Set1")
  }
names(col_class) <- levels(injections$class)
plot(NULL ,xaxt='n',yaxt='n',bty='n',ylab='',xlab='', xlim=0:1, ylim=0:1)
legend("topleft", legend = names(col_class),
       col = col_class, pch = 16, pt.cex=3, cex=1.5, bty='n')
mtext("Class", at=0.2, cex=2)
```


# Workflow
## readMSData
```{r}
raw_data <- readMSData(files = paste0(MZML_PATH, myfiles), 
                       pdata = new("NAnnotatedDataFrame", injections),
                       mode = "onDisk")
raw_data <- filterRt(raw_data, rt = c(0, rt_cut))
```

## Peak detection & post-processing
```{r}
cwp <- CentWaveParam(
  peakwidth = c(2, 20), 
  ppm = 50, 
  snthresh = 5,
  mzdiff = 0.001,
  prefilter = c(4, 5000),
  noise = 100,
  integrate = 2)

mnp <- MergeNeighboringPeaksParam(
  expandRt = 2, 
  expandMz = 0.001, 
  ppm = 10,
  minProp = 0.66)

xdata <- findChromPeaks(raw_data, param = cwp)

xdata <- refineChromPeaks(xdata,
                          param = mnp, 
                          msLevel = 1L,
                          BPPARAM = bpparam())
save.image("tmp.RData")
```

## Aligment and Correspondance
```{r}
pdp1 <- PeakDensityParam(
  sampleGroups = xdata$class, 
  minFraction = 0.9, 
  bw = 3,
  binSize = 0.01)

pdp2 <- PeakDensityParam(
  sampleGroups = xdata$class, 
  minFraction = 0.9, 
  bw = 2,
  binSize = 0.01)

if (study == "IS") {
  pgp <- PeakGroupsParam(
    span = 0.3,
    subset = which(xdata$class == "LowIS"),  
    subsetAdjust = "average",
    minFraction = 0.9 
  ) 
} else if (study == "QC"){
  pgp <- PeakGroupsParam(
    span = 0.3,
    subset = which(xdata$class == "QC_LowIS"),  
    subsetAdjust = "average",
    minFraction = 0.9 
  ) 
}

xdata <- groupChromPeaks(xdata, param = pdp1)
xdata <- adjustRtime(xdata, param = pgp)
plotAdjustedRtime(xdata,  col = col_class[xdata$class],
                  peakGroupsCol = "grey", peakGroupsPch = 1)
legend("topleft",
       legend=names(col_class), col = col_class,
       lty = 1, ncol = 1)
xdata <- groupChromPeaks(xdata, param = pdp2)
```

## Peak filling
```{r}
fcp <- FillChromPeaksParam(
  ppm = 20, 
  fixedRt = median(chromPeaks(xdata)[, "rtmax"] -
                     chromPeaks(xdata)[, "rtmin"]) / 2)

xdata <- fillChromPeaks(xdata, param = fcp)
```

## Data export
```{r}
ls()
save(list = c("raw_data", "xdata", 
              "cwp", "mnp", "pdp1", "pdp2", "pgp", "fcp",
              "MZML_PATH", "myfiles"), 
     file = paste0("XCMS_processing_", study, "_", polarity, ".RData"))
```

# Summary
Following I mention the changes in XCMS parameters that I’ve applied for this dataset:  
  
- Peak grouping (correspondance):  
  - minFraction = 0.9 (sample groups are made up of replicates from the same sample)  
  

# Session information
```{r}
Sys.time()-startpoint

devtools::session_info()
```
