---
title: "Sample files"
subtitle: "LCMS Standards"
author: "Mar Garcia-Aloy"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false
    toc_float: true
---


This code belows to the experiment of the injections of all mixes in triplicate and randomly.   
Mixes were injected in 2 different concentrations (low- and high-) and in water and in serum (QC).  
Vinicius Veri did the experiments on January 2020.  
Here I'm going to generate the table file with the names and metadata of injected files.  

## Preliminaries
# Libraries
```{r, message=FALSE, warning=FALSE}
startpoint <- Sys.time()
library(xcms)
library(doParallel)
library(CompMetaboTools)
```

## Parallel processing
```{r, echo=TRUE}
ncores <- detectCores()-1
registerDoParallel(ncores)
register(DoparParam(), default = TRUE)
```

# Table
Generate a data frame indicating the folder of the files (`2020/2020_01`) and their names:
```{r}
std_serum_files <- data.frame(
  folder <- "2020/2020_01",
  mzML   <- list.files("C:/Users/mgarciaaloy/Documents/mzML/2020/2020_01"))
colnames(std_serum_files) <- c("folder", "mzML")

tmp <- data.frame(
  do.call("rbind", strsplit(as.character(std_serum_files$mzML), "_")))
```

Indicate the polarity according to file name:
```{r}
std_serum_files$polarity <- gsub(".*_", "", std_serum_files$mzML)
std_serum_files$polarity <- gsub(".mzML", "", std_serum_files$polarity)
```

Add the time stamp from the individual files:
```{r, eval=TRUE}
raw_data <- readMSData(
  files = paste0("C:/Users/mgarciaaloy/Documents/mzML/", 
                 std_serum_files$folder, "/", std_serum_files$mzML), 
  mode = "onDisk")
std_serum_files$time <- as.character(extract_time_stamp(fileNames(raw_data)))
```

Indicate the sample type:
```{r}
std_serum_files$type <- NA
std_serum_files$type[
  which(grepl("Blank", std_serum_files$mzML) & 
          grepl("Water", std_serum_files$mzML))] <- "Blank_Water"
std_serum_files$type[
  which(grepl("Blank", std_serum_files$mzML) & 
          grepl("QC", std_serum_files$mzML))] <- "Blank_QC"
for(i in seq(nrow(std_serum_files))){
  if(is.na(std_serum_files$type[i])){
    if(tmp$X1[i] == "QC"){
      std_serum_files$type[i] <- as.character(tmp$X3[i])
    } else {
      std_serum_files$type[i] <- as.character(tmp$X2[i])
    }
  }
}
```

Indicate the sample class:
```{r}
std_serum_files$class <- NA
std_serum_files$class[
  which(grepl("Blank", std_serum_files$mzML) & 
          grepl("Water", std_serum_files$mzML))] <- "Blank_Water"
std_serum_files$type[
  which(grepl("Blank", std_serum_files$mzML) & 
          grepl("QC", std_serum_files$mzML))] <- "Blank_QC"
for(i in seq(nrow(std_serum_files))){
  if(is.na(std_serum_files$class[i])){
    if(tmp$X1[i] == "QC"){
      std_serum_files$class[i] <- 
        paste(as.character(tmp$X1[i]),
              as.character(tmp$X3[i]), 
              as.character(tmp$X2[i]), sep = "_")
    } else {
      std_serum_files$class[i] <- 
        paste("Water", as.character(tmp$X2[i]), 
              as.character(tmp$X1[i]), sep = "_")
    }
  }
}
std_serum_files$class <- gsub("IS", "", std_serum_files$class)
```

Indicate the scan mode:
```{r}
std_serum_files$mode <- "FS"
for(i in seq(nrow(std_serum_files))){
  if(grepl("CE", std_serum_files$mzML[i])){
    if(tmp$X1[i] == "QC"){
      std_serum_files$mode[i] <- as.character(tmp$X4[i])
    } else {
      std_serum_files$mode[i] <- as.character(tmp$X3[i])
    }
  }
}
```

# Output
```{r}
std_serum_files <- std_serum_files[order(std_serum_files$time),]
write.table(std_serum_files, "std_serum_files.txt", 
            sep = "\t", row.names = FALSE)
```


# Session information
```{r}
Sys.time()-startpoint

devtools::session_info()
```
