---
title: "Feature filtering + grouping"
subtitle: "Standards Library: MIX"
author: "Mar Garcia-Aloy"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false
    toc_float: true
---

```{r include = FALSE}
knitr::opts_chunk$set(echo=TRUE)

startpoint <- Sys.time()
```

# Preliminaries
## Parameters
```{r}
polarity <- "NEG" # specify "POS" or "NEG"
mix <- 17         # specify MIX number
mz.d <- 0.01
rt.d <- 10
myseed <- 589
filter.int <- 50000
cor.thr.i <- 0.7 # correlation intensity
cor.thr.s <- 0.9 # correlation peak shape
```

## Libraries
```{r, message=FALSE}
library(xcms)
```

## Functions
```{r}
stat.cv <- function(x){
  (sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE))*100
}

source("https://raw.githubusercontent.com/EuracBiomedicalResearch/lcms-standards/master/R/which_within.R")
```


## Data import
```{r}
if(mix < 10){
  load(paste0("XCMS_processing_MIX0", mix, "_", polarity, ".Rdata"))
} else {
  load(paste0("XCMS_processing_MIX", mix, "_", polarity, ".Rdata"))
}

# Data extraction
feat <- data.frame(featureDefinitions(xdata))
feat <- feat[,-ncol(feat)]
data <- featureValues(xdata, 
                      method = "sum",
                      value = "into", 
                      filled = FALSE)
data_NONimp <- data.frame(t(data))

# Missing value imputation
set.seed(myseed)
data_imp <- data.frame(t(imputeRowMinRand(data, method = "from_to")))
```



# Filtering
```{r}
dim(data_NONimp)
```

## Blanks
Features occurring in solvents and blanks: Keep features that were at least 2-fold higher in QCs or biological samples than in solvents.  

```{r}
xdata$blank <- 0
xdata$blank[xdata$class == "Blank"] <- 1
data_NONimp_0 <- data_NONimp
data_NONimp_0[is.na(data_NONimp_0)] <- 0
tmp <- data.frame(t(aggregate(data_NONimp_0, list(xdata$blank), mean)))
colnames(tmp) <- c("samples", "blank")
tmp <- tmp[-1,]
tmp$FC <- tmp$samples / tmp$blank
data_NONimp_filter <- 
  data_NONimp[, colnames(data_NONimp) %in% 
                       rownames(tmp)[tmp$FC > 2]]
dim(data_NONimp_filter)
```

## Intensity
Restrict to high intensity features
```{r}
feat$int <- apply(data_NONimp, 2, function(x){
  median(x, na.rm = TRUE)
})
data_NONimp_filter <- 
  data_NONimp_filter[, colnames(data_NONimp_filter) %in% 
                rownames(feat)[which(feat$int > filter.int)]]
dim(data_NONimp_filter)
```

## Variation
Keep features where between-group variation > within-group variation
```{r}
tmp <- aggregate(data_NONimp, by = list(gsub("_.*", "", xdata$sample_group)), stat.cv)
tmp <- setNames(data.frame(t(tmp[,-1])), tmp[,1])
feat$cv_intra <- NA
for(i in seq(nrow(tmp))){
  feat$cv_intra[i] <- max(tmp[i,], na.rm = TRUE)
}
tmp <- apply(data_NONimp, 2, stat.cv)
feat$cv_all <- tmp
data_NONimp_filter <- 
  data_NONimp_filter[, colnames(data_NONimp_filter) %in% 
                       rownames(feat)[which(feat$cv_all > 
                                              feat$cv_intra)]]
dim(data_NONimp_filter)

feat_filter <- feat[rownames(feat) %in% colnames(data_NONimp_filter), ]
```

# Grouping
```{r}
feat_filter$cmp <- NA
feat_filter$cor_int <- NA
feat_filter$cor_ps <- NA
cn <- 0
```

## General rule
```{r, message=FALSE}
# Order features by its intensity
feat_filter <- feat_filter[order(-feat_filter$int), ]

# start grouping
for(z in seq(nrow(feat_filter))){ # go feature by feature ("z")
  if(is.na(feat_filter$cmp[z])){ # go only in case that feature "z" 
                                 # is still not located in any group
    cn <- cn + 1 # increase by 1 the code to give the current group
    if(cn < 10){
      cmpname <- paste0("C000", cn)
    } else if (cn > 9 & cn < 100){
      cmpname <- paste0("C00", cn)
    } else if (cn > 99 & cn < 1000){
      cmpname <- paste0("C0", cn)
    } else{
      cmpname <- paste0("C", cn)
    }
    
    # get data of the current feature
    ft <- rownames(feat_filter)[z]
    mz.i <- feat_filter$mzmed[z]
    rt.i <- feat_filter$rtmed[z]
    
    # restrict grouping to still not located features
    feat.i <- feat_filter[is.na(feat_filter$cmp), ]
    feat.i <- feat.i[order(rownames(feat.i)), ]
    
    # get features co-eluting with feature "z"
    if(length(unlist(which_within(165.13130, mz.i)))==1 &
       feat.i$rtmin[rownames(feat.i)==ft]<90){ 
      # carnitine D3 (IS)
      feat.i <- feat.i[(feat.i$rtmin > 45) & (feat.i$rtmax < 80), ]
    } else if (length(unlist(which_within(154.0742434, mz.i)))==1) { 
      # glutamic acid (IS)
      feat.i <- feat.i[(feat.i$rtmin > 125) & (feat.i$rtmax < 210), ]
    } else {
      # all others
      feat.i <- feat.i[
        ((feat.i$rtmin > (feat.i$rtmin[rownames(feat.i)==ft] - rt.d)) & 
           (feat.i$rtmax) < (feat.i$rtmax[rownames(feat.i)==ft]+ rt.d)),
        ]
    } # close selection co-eluting features
    
    # correlate intensities
    if(nrow(feat.i) > 1){
      tmp <- data_NONimp[, colnames(data_NONimp) %in% rownames(feat.i)]
      cor.i <- c()
      for(i in seq(ncol(tmp))){
        tmp2 <- cbind(tmp[, ft], tmp[,i])
        if(sum(complete.cases(tmp2))>1){
          cor.i <- c(cor.i, cor(log2(tmp2[,1]), log2(tmp2[,2]), 
                              use = "complete.obs"))
        } else {
            cor.i <- c(cor.i, 0)
          }
      }
      feat.i$cor_int <- cor.i
      feat.i <- feat.i[feat.i$cor_int > cor.thr.i,]
      
      # correlate peak shapes
      xdata_n <- 
        filterFile(xdata, order(data_imp[, ft], decreasing=TRUE)[1:3][1])
      rtr <- c(min(feat.i$rtmin), max(feat.i$rtmax)) + rt.d * c(-1, 1)
      feat.i$cor_ps <- NA
      chr1 <- chromatogram(xdata_n, 
                           mz = mz.i + mz.d * c(-1, 1), 
                           rt = rtr, 
                           aggregationFun = "max")
      for(i in seq(nrow(feat.i))){
        chr2 <- chromatogram(xdata_n, 
                             mz = feat.i$mzmed[i] + mz.d * c(-1, 1), 
                             rt = rtr, 
                             aggregationFun = "max")
        feat.i$cor_ps[i] <- correlate(chr1[[1]], chr2[[1]])
      } # close EIC
      feat.i <- feat.i[feat.i$cor_ps > cor.thr.s,]
    } else if(nrow(feat.i) == 1){
      feat.i$cor_int <- 1
      feat.i$cor_ps <- 1
    }
    
    # add the info of grouped peaks in feat matrix
    idx <- which(rownames(feat_filter) %in% rownames(feat.i))
    feat_filter$cmp[idx] <- cmpname
    for(i in seq(nrow(feat.i))){
      idx <- which(rownames(feat_filter)==rownames(feat.i)[i])
      feat_filter$cor_int[idx] <- feat.i$cor_int[i]
      feat_filter$cor_ps[idx] <- feat.i$cor_ps[i]
    }
  } # close if "is.na(cmp[z])"
  save.image("tmp.RData")
} # close feat "z"
rm(idx)
length(levels(factor(feat_filter$cmp)))
```

## Particular cases
```{r}
# carnitine D3 (IS) ----------------------------------------------------
rt.i <- 53
mz.i <- 165.1313000
feat.i <- feat_filter[(feat_filter$rtmed > (rt.i - rt.d)) & 
                        (feat_filter$rtmed < (rt.i + rt.d)), ]
# dimer: 329.2553236
if(length(feat.i$cmp[unlist(which_within(329.2553236, feat.i$mzmed))])>0){
  if(feat.i$cmp[unlist(which_within(mz.i, feat.i$mzmed))] != 
     feat.i$cmp[unlist(which_within(329.2553236, feat.i$mzmed))]){
    feat.i$cmp[which(
      feat.i$cmp == 
        feat.i$cmp[unlist(which_within(329.2553236, feat.i$mzmed))])] <- 
      feat.i$cmp[unlist(which_within(mz.i, feat.i$mzmed))]
  }
}

# Na adduct: 187.1132447
if(length(unlist(which_within(mz.i, feat.i$mzmed))) > 0){
  if(feat.i$cmp[unlist(which_within(mz.i, feat.i$mzmed))] != 
   feat.i$cmp[unlist(which_within(187.1132447, feat.i$mzmed))]){
  feat.i$cmp[which(
    feat.i$cmp == 
      feat.i$cmp[unlist(which_within(187.1132447, feat.i$mzmed))])] <- 
    feat.i$cmp[unlist(which_within(mz.i, feat.i$mzmed))]
}
}
idx <- which(rownames(feat_filter) %in% rownames(feat.i))
feat_filter$cmp[idx] <- feat.i$cmp

length(levels(factor(feat_filter$cmp)))
```


# Save output
```{r}
ls()
save(list = c(
  # XCMS data
  "raw_data", "xdata", "MZML_PATH", "myfiles",
  "cwp", "mnp", "pdp1", "pdp2", "pgp", "fcp",
  
  "data_imp", "data_NONimp", "feat",
  
  # Annotations
  "feat_filter"),
  file = paste0("filter_group_",  
                levels(xdata$class)[levels(xdata$class) != "Blank"], 
                "_", polarity, ".Rdata"))
```


# Session information
```{r}
Sys.time()-startpoint

devtools::session_info()
```

