---
title: "Evaluation XCMS"
subtitle: "Standards Library: MIX"
author: "Mar Garcia-Aloy"
output: 
  BiocStyle::html_document:
    toc: true
    number_sections: false
    toc_float: true
---

```{r include = FALSE}
knitr::opts_chunk$set(echo=TRUE)

startpoint <- Sys.time()
```

In this document I'm going to check the quality of the data obtained 
after processing raw files with XCMS. Different things to 
look for are to check if there is only 1 detected peak per compound 
(peak detection step), as well as, 1 feature per compound 
(correspondance step). 

# Preliminaries
I define the specific parameters for the current study, I load all 
required libraries, and I set up the parallel processing.

## Parameters
```{r, echo=TRUE}
polarity <- "POS" # specify "POS" or "NEG"
mix <- 1         # specify MIX number
mz.d <- 0.01
rt.d <- 10
if(mix < 10){
  dir.create(paste0("images/MIX0", mix, "/", polarity, "/XCMS_eval"), 
                  recursive = TRUE, showWarnings = FALSE)
} else {
  dir.create(paste0("images/MIX", mix, "/", polarity, "/XCMS_eval"), 
                  recursive = TRUE, showWarnings = FALSE)
}
```

## Libraries
```{r, message=FALSE, warning=FALSE}
library(doParallel)
library(xcms)
library(RColorBrewer)
library(Rdisop)
library(CompoundDb)
```

## Parallel processing
```{r}
ncores <- detectCores()-1
registerDoParallel(ncores)
register(DoparParam(), default = TRUE)
```

## Data import
```{r}
if(mix < 10){
  load(paste0("XCMS_processing_MIX0", mix, "_", polarity, ".Rdata"))
} else {
  load(paste0("XCMS_processing_MIX", mix, "_", polarity, ".Rdata"))
}
```

## Colouring factors
Also, I define colors to be used for the various experimental groups 
throughout the experiment.
```{r}
getPalette <- colorRampPalette(brewer.pal(9, "YlOrRd"))
col_class <- c("grey", 
               rev(getPalette(length(which(xdata$class != "Blank")))))
names(col_class) <- levels(factor(xdata$sample_group))
plot(NULL ,xaxt='n',yaxt='n',bty='n',ylab='',xlab='', xlim=0:1, ylim=0:1)
legend("top", legend = names(col_class), bty='n', 
       col = col_class, pch = 16, pt.cex = 1.5, cex = 0.9, ncol = 3)
mtext("Class", at=0.2, cex=2)
```

## Standards library
### Import
```{r}
std_info <- read.table(
    "https://raw.githubusercontent.com/EuracBiomedicalResearch/lcms-standards/master/data/standards_dilution.txt",
    sep = "\t", header = TRUE, as.is = TRUE)
std_info <- std_info[std_info$mix == mix, ]
```

### Quality
```{r, eval=FALSE}
std_info$quality <- c(NA,      # succinic
                      "2P",    # creatinine   
                      "2P",    # glucose      
                      "OK",    # carnitine
                      "2P",    # alanine
                      "2P",    # arginine
                      "OK",    # aspartic
                      "OK",    # cysteine
                      "2P",    # glutamic
                      "2P",    # glycine      
                      "OK",    # histidine
                      NA,      # isoleucine
                      "2P",    # leucine
                      "OK",    # lysine
                      "OK",    # methionine    
                      "OK",    # phenylalanine
                      "OK",    # proline
                      "OK",    # serine        
                      "OK",    # threonine
                      "2P",    # tyrosine
                      "OK"     # valine
                      )
```

### Additional formating
```{r}
std_info <- std_info[!is.na(std_info[, grep(polarity, 
                                            colnames(std_info))]),]
rownames(std_info) <- seq(nrow(std_info))

std_info$mzneut = NA
std_info$mz <- NA
for(i in seq(nrow(std_info))){
  if(grepl("C", std_info$formula[i])){std_info$mzneut[i] = 
    getMolecule(as.character(std_info$formula[i]))$exactmass}else{
      std_info$mzneut[i] = as.numeric(std_info$formula[i])}
  std_info$mz[i] <- unlist(
        mass2mz(std_info$mzneut[i],
                adduct = as.character(
                    std_info[i, grep(polarity, colnames(std_info))])))
}
```

# Workflow
```{r}
dim(featureValues(xdata))

for(i in seq(nrow(std_info))){ # for each STD do the following:
  
  mz.i <- std_info$mz[i]
  rt.i <- std_info$RT[i]
  cmpname <- tolower(gsub(" .*", "", std_info$name[i]))
  cmpname <- gsub("l-", "", cmpname)
  
  # PEAK DETECTION
  chr <- chromatogram(xdata, mz = c((mz.i - mz.d), (mz.i + mz.d)),
                      aggregationFun = "max")
  if(nrow(featureDefinitions(chr)) == 1){
    ft <- rownames(featureDefinitions(chr))
  } else if (nrow(featureDefinitions(chr)) > 1){
    ft <- rownames(featureDefinitions(chr))[
      which.max(rowMeans(featureValues(chr, value = "into"), 
                         na.rm = TRUE))]
  }
  if(exists("ft")){
    ft_chr <- featureChromatograms(xdata, features = ft)
  
  filename <-
    paste0("images/", 
           levels(xdata$class)[levels(xdata$class) != "Blank"], 
           "/", polarity, "/XCMS_eval/", cmpname, "_FT.png")
  png(file = filename, width = 1000, height = 500)
  par(mfrow = c(1,2))
  plot(ft_chr, col = "#00000040", 
     peakCol = col_class[xdata$sample_group[chromPeaks(ft_chr)[, "column"]]],
     peakBg = 
       paste0(col_class[xdata$sample_group[chromPeaks(ft_chr)[, "column"]]],
              "10"))
  plot(ft_chr, col = "#00000040", 
       peakCol = col_class[xdata$sample_group[chromPeaks(ft_chr)[, "column"]]],
       peakBg = paste0(col_class[
         xdata$sample_group[chromPeaks(ft_chr)[, "column"]]], "10"))
  xcms:::.add_chromatogram_peaks(
    ft_chr, chromPeaks(ft_chr),
    col = col_class[xdata$sample_group[chromPeaks(ft_chr)[, "column"]]],
    bg = NA,
    type = "rectangle")
  dev.off()
  } else {
    print(paste(std_info$name, "NOT detected"))
  }
  
  # CORRESPONDANCE
  filename <-
    paste0("images/", 
           levels(xdata$class)[levels(xdata$class) != "Blank"], 
           "/", polarity, "/XCMS_eval/", cmpname, "_GR.png")
  png(file = filename, width = 1000, height = 500)
  if(exists("ft")) {
    plotChromPeakDensity(
    chr, pdp2, 
    peakCol = col_class[xdata$sample_group[chromPeaks(chr)[, "column"]]],
    peakBg = paste0(col_class[
      xdata$sample_group[chromPeaks(chr)[, "column"]]], "10"))
  } else {
    plot(chr, col = col_class[xdata$sample_group])
  }
  dev.off()
  
  chr2 <- chromatogram(raw_data, mz = c((mz.i - mz.d), (mz.i + mz.d)),
                      aggregationFun = "max")
  filename <-
    paste0("images/", 
           levels(xdata$class)[levels(xdata$class) != "Blank"], 
           "/", polarity, "/XCMS_eval/", cmpname, "_GR_zoom.png")
  png(file = filename, width = 1000, height = 500)
  if(nrow(chromPeaks(findChromPeaks(chr2, param = cwp))) > 0){
    chr2 <- findChromPeaks(chr2, param = cwp)
  delta_rt <- abs(chromPeaks(chr2)[, "rt"] - rt.i)
  pks <- data.frame(chromPeaks(chr2)[delta_rt < 60, , drop=FALSE])
  if (nrow(pks)) {
    pks2 <- pks[0, ]
    for(j in seq(length(levels(factor(pks$column))))){
      pks3 <- pks[pks$column == j, ]
      pks2 <- rbind(pks2, pks3[which.max(pks3$maxo),])
    }
    #pks2 <- pks2[pks2$column > 3, ]
    rtmin <- min(pks2$rtmin) - 5
    rtmax <- max(pks2$rtmax) + 5
    ylim <- c(0, max(pks$maxo, na.rm = TRUE))
    } else {
      rtmin <- std_info$RT[i] - 5
      rtmax <- std_info$RT[i] + 5
      ylim <- c(0, max(sapply(chr, intensity), na.rm = TRUE))
    }
  chr2 <- chromatogram(xdata, mz = c((mz.i - mz.d), (mz.i + mz.d)),
                       rt = c((rt.i - rt.d), (rt.i + rt.d)),
                      aggregationFun = "max")
  plotChromPeakDensity(
    chr2, pdp2, 
    peakCol = col_class[xdata$sample_group[chromPeaks(chr2)[, "column"]]],
    peakBg = paste0(col_class[
      xdata$sample_group[chromPeaks(chr2)[, "column"]]], "10"))
  } else {
    plot(chr, col = col_class[xdata$sample_group], xlim = rt.i + 30 * c(-1,2))
  }
  dev.off()
  rm(ft)
}
```

# Session information
```{r}
Sys.time()-startpoint

devtools::session_info()
```

